### 2024-11-10 Exam Assignments/Chapters/01/01.tex
week 4

\documentclass[../../main.tex]{subfiles}

\begin{document}

\section{Introduction to Algorithm Engineering}
\subsection{Analyzing slide 15}
Here is slide 15 again:

~\\
\includegraphics[width=\textwidth]{page_15.png}

~\\
Let us proof the satement
\[
    \int_{0}^{1}\frac{1}{1+x^2} dx = \frac{\pi}{4}
\]

\noindent using
\[
    (f^{-1})'(x_0) = \frac{1}{f'(f^-1(x_0))} \ \ \ \ \ :
\]

\begin{proof}
    Let \( \tan : (- \frac{\pi}{2}, \frac{\pi}{2}) \rightarrow \mathbb R \) be the usual trigonometric function.
    Clearly, it is a bijection, and hence its inverse function \( \arctan : \mathbb R \rightarrow (- \frac{\pi}{2}, \frac{\pi}{2}) \) exists.
    Now, let us argue with the above identity:
    \[
        \arctan'(x_0) = \frac{1}{\tan'(\arctan(x_0))} = \cos^2(\arctan(x_0)) \ \ \ \ \ ,
    \]

    \noindent
    where we have used that \(\tan'(x_0) = \frac{1}{\cos^2(x_0)}\).
    (This can easily be derived by the chain and product rule, knowing the derivatives of \(\sin , \cos \) and \(\frac{1}{x}\), as well as the fact that \(\tan(x)=\frac{\sin(x)}{\cos(x)}\).)

    \noindent
    Now, we would like to transform \(\cos^2(\bullet)\) into something like foo\((\tan(\bullet))\), such that $\tan$ and $\arctan$ would cancel.
    Luckily, we can do this using the trigonometric identity

    \begin{align*}
    &\sin^2 + \cos^2 = 1 \\
    \Leftrightarrow \ &\tan^2 \cos^2 + \cos^2 = 1 \\
    \Leftrightarrow \ &(\tan^2 + 1) \cos^2 = 1 \\
    \Leftrightarrow \ &\cos^2 = \frac{1}{\tan^2 + 1}
    \end{align*}

    ~\\
    \noindent
    Plugging this into our equation yields

    \[
        \arctan'(x_0) = \frac{1}{\tan^2(\arctan(x_0)) + 1} \ \ \ \ \ .
    \]

    ~\\
    But by the definition of $\arctan$, we get

    \[
        \arctan'(x_0) = \frac{1}{x_0^2 + 1} \ \ \ \ \ .    
    \]

    ~\\
    Thus, we have established that $\arctan$ is an antiderivative of $\frac{1}{x^2 + 1}$.
    \\
    By the fundamental theorem of calculus we get

    \[
        \int_{0}^{1}\frac{1}{1+x^2} dx = \arctan(1) - \arctan(0) \ \ \ \ \ .
    \]

    ~\\
    Now note that \(0, \frac{\pi}{4} \in (- \frac{\pi}{2}, \frac{\pi}{2})\), and \(\tan(0) = 0\), \(\tan(\frac{\pi}{4}) = 1\).
    In other words, we have \(\arctan(0) = 0\),  \(\arctan(1) = \frac{\pi}{4}\). Thus,

    \[
        \int_{0}^{1}\frac{1}{1+x^2} dx = \frac{\pi}{4} \ \ \ \ \ .
    \]

\end{proof}

~\\
The reason we established this equation in the first place is because it allows for a simple approximation for $\pi$.
We do this by using numerical methods for approximating integrals.

One way to do this is to calculate the Riemann sum, as shown in the lecture.
The Riemann sum tries to approximate little intervals of the function $\frac{1}{1+x^2}$ by constant functions.
A better way to approximate the function on these intervals would be a linear function connecting the end points, resulting in the \em Trapezoidal Rule\em.

Even better is the approximation by parabolas, which pass through the end points of the interval as well as the mid point. This is called the \em Simpson Rule\em.
It reads as follows:
\[
    \int_{t}^{t+h} f(x) dx \ \approx \ \frac{h}{6} \Big[ f(t) + 4f(t+\frac{h}{2}) + f(t+h) \Big] \ \ \ \ \ .
\]

~\\
When approximating an integral by splitting it into smaller intervals, we use the following formula.
To this end, let $P$ be such a partition. Then

\[
    \int_{a}^{b} f(x) dx \ \ \approx \sum_{(t, t+h) \in P} I_{(t, t+h)}(f) \ \ \ \ \ ,
\]

where \(I_{(t, t+h)}(f)\) tries to approximate \(\int_{t}^{t+h} f(x) dx\).

~\\
Combining these results, one might get even faster convergence.
But note that machines might have trouble dealing with very small numbers, thus this process should be used with caution.

~\\
\subsection{Chapter 1 from Computer Systems: A Programmer's Perspective}
The book gave a quick overview about operating systems and computer hardware, and the flow of execution.
I'd like to discuss the memory hierarchy as well as virtual memory.

\subsubsection{Memory Hierarchy}
The inverse correlation between speed and volume of different memory types war presented in the book.
This lead to the memory hierarchy, where we try exploit the benefits of both worlds be having big, slow memory, and a smaller amount of fast memory, which we the computer tries to employ as much as possible.

~\\
Two reasons for this are
\begin{enumerate}
    \item Increased physical memory cells: More logical memory requires more physical memory cells. These cells need to be connected to control circuits, and larger memory arrays result in longer signal paths and higher capacitance. This increases signal propagation delay, making the memory slower.
    \item Heat production: Larger memory volumes consume more power, and the switching of many transistors generates more heat. Excessive heat can degrade performance and force memory to operate at lower speeds to prevent overheating.
\end{enumerate}

~\\
\subsubsection{Virtual Memory}
One of the core principles in mathematics and computer science is abstraction.
The operating system provides such an abstraction in form of I/O communication, running multiple programs concurrently using processes, and translating virtual addresses into physical ones.

This is very interesting, as by doing so, two distinct processes are unable to access the other ones data.
To illustrate this, imagine you write a simple C program, create a pointer, print the pointer location, write an integer there based on a command line argument, and print the content periodically.
If one now starts two instances of the program, one where it writes a 1, another one with 0, they won't interfere with one another, the first process will always read a 1, the second one 0.

Behind the scenes, the operating system works together with the \em memory management unit\em, which translates every address the program uses to its real address before accessing.
And exactly this translation process is program specific.
Hence, the operating system successfully abstracted the memory.

~\\
\newpage
\subsection{Parallelizing "Estimating $\pi$ using Monte Carlo"}
Here is the parallelized version of the code:

\smallbreak
\smallbreak
\begin{samepage}
\begin{lstlisting}
#include <iostream>
#include <omp.h>
#include <random>

using namespace std;

int main() {

    int n = 100000000; // number of points to generate
    int counter = 0; // counter for points lying in the first quadrant of a unit circle
    auto start_time = omp_get_wtime(); // omp_get_wtime() is an OpenMP library routine

    // compute n points and test if they lie within the first quadrant of a unit circle
    #pragma omp parallel
    {
        default_random_engine re{(size_t) omp_get_thread_num()};
        uniform_real_distribution<double> zero_to_one{0.0, 1.0};

        int local_counter = 0;
        int local_n = (n / omp_get_num_threads()) + ((n % omp_get_num_threads() > omp_get_thread_num()) ? 1 : 0);
        for (int i = 0; i < local_n; ++i) {
            auto x = zero_to_one(re); // generate random number between 0.0 and 1.0
            auto y = zero_to_one(re); // generate random number between 0.0 and 1.0
            if (x * x + y * y <= 1.0) { // if the point lies in the first quadrant of a unit circle
                ++local_counter;
            }
        }
        #pragma omp atomic
        counter += local_counter;
    }

    auto run_time = omp_get_wtime() - start_time;
    auto pi = 4 * (double(counter) / n);

    cout << "pi: " << pi << endl;
    cout << "run_time: " << run_time << " s" << endl;
    cout << "n: " << n << endl;
}
\end{lstlisting}
\end{samepage}
\end{document}

### 2024-11-10 Exam Assignments/Chapters/02/02.tex
week 4

\documentclass[../../main.tex]{subfiles}

\begin{document}
    \section{False Sharing, Race Conditions, and Schedules}
    \subsection{False Sharing}
    Consider two teachers who need to count their students. To divide the work, one teacher decides to count the boys, while the other counts the girls. They use a single sheet of paper to keep track, marking a stroke on their respective sides of the page for each student they count.

    ~\\
    However, even though they each have their own pen, they're still working with only one piece of paper. This setup means that every time one teacher wants to make a mark, they must pass the paper to the other, as both are using different parts of the same sheet.

    ~\\
    This constant handoff slows them down significantly. Had they used separate sheets, they could each count freely without interrupting one another.

    ~\\
    In the same way, false sharing occurs in computing when multiple threads modify different parts of the same cache line. Even though each thread may be working on separate variables, they're forced to constantly invalidate and reload the cache line they "share," leading to substantial inefficiencies.

    ~\\
    Here is an illustration of the sceario:

    \begin{figure}[h]
        \centering
        \resizebox{\textwidth}{!}{
            \begin{tikzpicture}
                % Paper
                \draw[thick] (0, 0) rectangle (6, 3);
                \node[above] at (3, 3) {Shared Paper};

                % Divider line
                \draw[dashed] (3, 0) -- (3, 3);

                % Labels for boys' and girls' side
                \node at (1.5, 1.5) {Boys' Count Side};
                \node at (4.5, 1.5) {Girls' Count Side};

                % Teacher 1
                \node[draw, rectangle, minimum width=2cm, minimum height=1cm, fill=blue!20] (teacher1) at (-3, 1.5) {Teacher 1};
                \node[below=0.1cm of teacher1] {Counting Boys};
                \draw[->, thick] (teacher1.east) -- (-0.2, 1.5);

                % Teacher 2
                \node[draw, rectangle, minimum width=2cm, minimum height=1cm, fill=red!20] (teacher2) at (9, 1.5) {Teacher 2};
                \node[below=0.1cm of teacher2] {Counting Girls};
                \draw[->, thick] (teacher2.west) -- (6.2, 1.5);

                % Arrows to represent passing the paper
                % \draw[<->, thick, dashed] (teacher1.east) -- node[above]{Pass Paper} (teacher2.west);
            \end{tikzpicture}
        }
        \caption{False Sharing Analogy: Two teachers repeatedly passing a single paper to count students separately.}
    \end{figure}

    \newpage
    \subsection{There's plenty of room at the Top}
    Let us analyze the following graphic from the paper \em There’s plenty of room at the Top: What
will drive computer performance after Moore’s law? \em:

    ~\\
    \includegraphics[width=\textwidth]{Chapters/02/more_room_at_the_top.png}

    ~\\
    The entire blue area basically represents the foundation of running software:
    We need hardware ("The Bottom") as well as good software ("The Top").
    The width represents the potential of gains we could make in these areas.
    In particular, the width of "The Bottom" shrank over time and is now becoming less significant.

    ~\\
    The paper assumes that the most potential lies in the three categories \em Software, Algorithms, and Hardware Architecture\em, all of which are located at The Top.
    They describe their potentials in the \em Opportunity \em and \em Examples \em fields. Let us analyze them individually:

    ~\\
    \subsubsection{Software}
    The software that is ultimately run on the hardware can be optimized itself.
    In Software Engineering, there is a phenomenon called \emph{bloat}.
    It describes the accumulation of unnecessary functionality and features, which makes programs bigger and hence slower.
    By carefully writing code to avoid these traps, we can improve the running time of our programs.

    ~\\
    Tailoring software to hardware features allows developers to leverage specific capabilities of the underlying hardware. For instance, optimizing software to use SIMD (Single Instruction, Multiple Data) instructions can dramatically increase performance for tasks that involve large data sets, such as graphics processing or numerical simulations. Additionally, understanding cache architectures can help developers design algorithms that reduce cache misses, leading to faster data access and processing. By aligning software architecture with hardware capabilities, significant improvements in execution speed and resource utilization can be achieved.

    ~\\
    \subsubsection{Algorithms}
    The choice of algorithms plays a critical role in determining software performance. Implementing new algorithms that are more efficient in terms of time and space complexity can lead to significant performance enhancements. For example, employing faster sorting algorithms or utilizing data structures that provide quicker access times can reduce the computational overhead.

    ~\\
    New problem domains often introduce unique challenges that require innovative algorithmic solutions. For instance, the advent of big data has prompted the development of algorithms specifically designed for distributed computing environments, which can process large datasets across multiple machines. By addressing the specific needs of these new domains, developers can create algorithms that are optimized for performance and scalability.

    ~\\
    \subsubsection{Hardware Architecture}
    Hardware streamlining involves optimizing hardware components to enhance overall system performance. This can include reducing the complexity of circuits, minimizing power consumption, and improving thermal management. Streamlined hardware can lead to faster processing speeds and better energy efficiency, directly impacting software performance by reducing bottlenecks.

    ~\\
    Processor simplification is another approach to improving performance. By reducing the number of transistors or focusing on a reduced instruction set, processors can operate at higher speeds and with lower power consumption. Simplified processors are often easier to design, implement, and manufacture, which can also reduce costs while enhancing performance metrics.

    ~\\
    Domain specialization allows hardware to be optimized for specific applications or problem domains. For example, graphics processing units (GPUs) are designed to handle complex graphics calculations more efficiently than general-purpose CPUs. Similarly, application-specific integrated circuits (ASICs) can be tailored for particular tasks, such as cryptocurrency mining or machine learning inference, resulting in substantial performance gains over generic hardware solutions.

    \newpage
    \subsection{Parallelizing pi\_numerical\_integration.cpp}
    Here is the parallelized version using the \em \#pragma omp for \em construct:

    \smallbreak
    \smallbreak
    \begin{samepage}
    \begin{lstlisting}
#include <iomanip>
#include <iostream>
#include <omp.h>

using namespace std;

int main() {
    int num_steps = 100000000; // number of rectangles
    double width = 1.0 / double(num_steps); // width of a rectangle
    double sum = 0.0; // for summing up all heights of rectangles

    double start_time = omp_get_wtime(); // wall clock time in seconds
    #pragma omp parallel for reduction(+:sum)
    for (int i = 0; i < num_steps; i++) {
        double x = (i + 0.5) * width; // midpoint
        sum = sum + (1.0 / (1.0 + x * x)); // add new height of a rectangle
    }
    double pi = sum * 4 * width; // compute pi
    double run_time = omp_get_wtime() - start_time;

    cout << "pi with " << num_steps << " steps is " << setprecision(17)
        << pi << " in " << setprecision(6) << run_time << " seconds\n";
}
    \end{lstlisting}
    \end{samepage}
\end{document}

### 2024-11-10 Exam Assignments/Chapters/03/03.tex
week 4

\documentclass[../../main.tex]{subfiles}

\begin{document}
\section{Parallelizing Code with OpenMP}
\subsection{Chapter 3 of Introduction to Parallel Computing}
\subsubsection{Comparing code snippets}
The code in listing 3.1 reads as follows:

\begin{lstlisting}
#include <stdio.h>
#include <omp.h>

int main() {
    printf("Hello, world:");
    #pragma omp parallel
        printf(" %d", omp_get_thread_num());
    printf("\n");
    return 0;
}
\end{lstlisting}

~\\
In this code, a team of thread get created my the \em \#pragma omp parallel \em clause.
Their task is to print their respective thread numbers.
But since no order is defined, the order of their execution are "random", i.e. determined at runtime by the operating system.

~\\
Listing 3.2 reads as follows:

\begin{lstlisting}
#include <stdio.h>
#include <omp.h>

long fib(int n) {
    return (n < 2 ? 1 : fib(n-1) + fib(n-2));
}

int main() {
    int n = 45;
    #pragma omp parallel
    {
        int t = omp_get_thread_num();
        printf("%d: %d\n", t, fib(n+t));
    }
    return 0;
}
\end{lstlisting}

~\\
While still not defining the order of the tasks, they yet seem to execute in order.
Interestingly, this is not quite true, as the main reason the seem to do so is because the \em terminate \em in order.
This is simply due to the fact that the work load for a thread with a higher thread number is substantially higher than for another one with a lower number.
This is because fib(n+t) is invoked, but it takes very long to calculate because of the big n, and even a singe step by t causes fib(n+t) to take nearly twice as long.

\subsubsection{Game of Life}
The code in Listing 3.10 looks well designed.
It collapses the two nested for loops of size $size$ into a single one of size $size^2$.
Since each iteration approximately takes the same amount of work, the implicit static scheduling ensures a good distribution of the workload.
This way, every thread has to compute \( \frac{size^2}{\text{omp\_get\_num\_threads()}} \) iterations.

~\\
The outer most loop cannot be parallelized, as their iterations are all strictly dependent on the previous one.
This is why the implicit barrier in line 13 is neccessary.
It also shouldn't be that big of a bottleneck, as the threads should reach it almost simultaneously.

~\\
\subsubsection{Optimizing Random Shooting}
When trying to approximate $\pi$ by the method of random shooting we use random numbers.
They, however, don't need to be truly random, they just have to simulate it well enough.
One way to do this simulation of randomness is the rnd function like it is defined in the book (or see the code below).
It calculates a random number quite efficiently.
Another benfit is that we do not need multiple instances of random number generators, and we also don't run the risk of misusing those by calling their methods in different threads.
The code is shown in the \hyperref[fig:calculate_pi_v2]{figure} below.
It runs very quickly and calculates $\pi$ very precisely.
The running time on my particular machine was 0.0313137 s, in comparision to the previous version with a running time of 0.117771 s.

\newpage
\begin{figure}
\label{fig:calculate_pi_v2}
\begin{lstlisting}
#include <iostream>
#include <omp.h>
#include <random>

using namespace std;

double rnd(unsigned int *seed)
{
    *seed = (1140671485 * (*seed) + 12820163) % (1 << 24);
    return ((double)(*seed)) / (1 << 24);
}

int main()
{
    int n = 100000000;                 // number of points to generate
    int counter = 0;                   // counter for points lying in the first quadrant of a unit circle
    auto start_time = omp_get_wtime(); // omp_get_wtime() is an OpenMP library routine

    // compute n points and test if they lie within the first quadrant of a unit circle
    #pragma omp parallel reduction(+:counter)
    {
        unsigned int seed = omp_get_thread_num();
        size_t local_n = (n / omp_get_num_threads()) + ((n % omp_get_num_threads() > omp_get_thread_num()) ? 1 : 0);

        for (int i = 0; i < local_n; ++i)
        {
            auto x = rnd(&seed); // generate random number between 0.0 and 1.0
            auto y = rnd(&seed); // generate random number between 0.0 and 1.0
            if (x * x + y * y <= 1.0)
            { // if the point lies in the first quadrant of a unit circle
                ++counter;
            }
        }
    }

    auto run_time = omp_get_wtime() - start_time;
    auto pi = 4 * (double(counter) / n);

    cout << "pi: " << pi << endl;
    cout << "run_time: " << run_time << " s" << endl;
    cout << "n: " << n << endl;
}
\end{lstlisting}
\end{figure}
\end{document}

### 2024-11-10 Exam Assignments/Chapters/04/04.tex
week 4

\documentclass[../../main.tex]{subfiles}

\begin{document}
\section{Quicksort and Concurrency}
\subsection{Comparing quicksort implementations}
\subsubsection{Benchmarking environment}
The benchmarks were performed on the following system:

\begin{description}[style=nextline, labelwidth=4.8cm, labelindent=0cm, leftmargin=5cm, font=\normalfont\bfseries]
    \item[Operating System:] Ubuntu 24.04.1 LTS
    \item[RAM:] 32 GB
    \item[CPU:] AMD Ryzen 5 5600G with Radeon Graphics (6 Cores, 2 Threads per Core)
    \item[Compiler:] gcc version 13.2.0
\end{description}

\newpage
\subsubsection{Running time comparison with different numbers of threads}
Here are the results when running the sorting algorithms with varying numbers of threads:

\includegraphics[width=\linewidth]{./running_times.png}

~\\
And here are the performance gains compared to std::sort():

\includegraphics[width=\linewidth]{./speedup.png}

\newpage
\subsubsection{Running time comparison with different array sizes}
Here are the running times with varying array sizes:

\includegraphics[width=\linewidth]{./running_times_array.png}

~\\
These are the speedups compared to std::sort():

\includegraphics[width=\linewidth]{./speedup_array.png}

~\\
\subsubsection{Explaining the results}
In the first two graphs when measuring the running times and speed ups when varying the threads, we can clearly see that the serial std::sort has a constant running time, whereas both parallel sorting algorithms run quicker with increasing thread number, thus overtaking std::sort already with two threads significantly (almost halving the running time of std::sort).
With further increasing thread number, the running times of the parallel algorithms further decreases.
They are proportional to $\frac{1}{\text{Number of Threads}}$, as one might expect, but the proportionality factor is roughly at 2, meaning there is quite some overhead compared to the optimal running time of $\frac{\text{Serial Running Time}}{\text{Number of Threads}}$.

~\\
Let us now analyze the second two pictures, showing the running times at different array sizes.
The graphs are as expected, the running times of the parallel algorithms is better, even more so for very large array sizes, as in those instances we can fully employ the parallel power of the cpu cores, thus decreasing the running time significantly.
This is why the speedup gets better with increasing array size.
One might think that this speedup is exponential, but it is not (note the logarithmic scale!), it is rather approximately $(\text{Serial Running Time}) \cdot (1 - \frac{1}{12 \cdot E})$, where $E$ is the efficiency $E = \frac{\frac{\text{Serial Running Time}}{\text{Threads}}}{\text{Parallel Running Time}}$, and note that the number of threads on this machine is 12.
Note, that Serial Running Time is expected to grow $\Theta(n \cdot \log(n))$, and so should the speedup.

\end{document}

### 2024-11-10 Exam Assignments/Chapters/01/calculate_pi.cpp

@@ -1,0 +1,38 @@
+#include <iostream>
+#include <omp.h>
+#include <random>
+
+using namespace std;
+
+int main() {
+    
+    int n = 100000000; // number of points to generate
+    int counter = 0; // counter for points lying in the first quadrant of a unit circle
+    auto start_time = omp_get_wtime(); // omp_get_wtime() is an OpenMP library routine
+
+    // compute n points and test if they lie within the first quadrant of a unit circle
+    #pragma omp parallel
+    {
+        default_random_engine re{(size_t) omp_get_thread_num()};
+        uniform_real_distribution<double> zero_to_one{0.0, 1.0};
+
+        int local_counter = 0;
+        int local_n = (n / omp_get_num_threads()) + ((n % omp_get_num_threads() > omp_get_thread_num()) ? 1 : 0);
+        for (int i = 0; i < local_n; ++i) {
+            auto x = zero_to_one(re); // generate random number between 0.0 and 1.0
+            auto y = zero_to_one(re); // generate random number between 0.0 and 1.0
+            if (x * x + y * y <= 1.0) { // if the point lies in the first quadrant of a unit circle
+                ++local_counter;
+            }
+        }
+        #pragma omp atomic
+        counter += local_counter;
+    }
+
+    auto run_time = omp_get_wtime() - start_time;
+    auto pi = 4 * (double(counter) / n);
+
+    cout << "pi: " << pi << endl;
+    cout << "run_time: " << run_time << " s" << endl;
+    cout << "n: " << n << endl;
+}

### 2024-11-10 Exam Assignments/Chapters/03/calculate_pi_v2.cpp

@@ -1,0 +1,42 @@
+#include <iostream>
+#include <omp.h>
+#include <random>
+
+using namespace std;
+
+double rnd(unsigned int *seed)
+{
+    *seed = (1140671485 * (*seed) + 12820163) % (1 << 24);
+    return ((double)(*seed)) / (1 << 24);
+}
+
+int main()
+{
+    int n = 100000000;                 // number of points to generate
+    int counter = 0;                   // counter for points lying in the first quadrant of a unit circle
+    auto start_time = omp_get_wtime(); // omp_get_wtime() is an OpenMP library routine
+
+    // compute n points and test if they lie within the first quadrant of a unit circle
+    #pragma omp parallel reduction(+:counter)
+    {
+        unsigned int seed = omp_get_thread_num();
+        size_t local_n = (n / omp_get_num_threads()) + ((n % omp_get_num_threads() > omp_get_thread_num()) ? 1 : 0);
+
+        for (int i = 0; i < local_n; ++i)
+        {
+            auto x = rnd(&seed); // generate random number between 0.0 and 1.0
+            auto y = rnd(&seed); // generate random number between 0.0 and 1.0
+            if (x * x + y * y <= 1.0)
+            { // if the point lies in the first quadrant of a unit circle
+                ++counter;
+            }
+        }
+    }
+
+    auto run_time = omp_get_wtime() - start_time;
+    auto pi = 4 * (double(counter) / n);
+
+    cout << "pi: " << pi << endl;
+    cout << "run_time: " << run_time << " s" << endl;
+    cout << "n: " << n << endl;
+}

### 2024-11-10 Exam Assignments/Chapters/04/min_max_quicksort.cpp

@@ -1,0 +1,167 @@
+#include <iostream>
+#include <vector>
+#include <limits>
+#include <omp.h>
+#include <parallel/algorithm>
+#include <fstream>
+
+// compute the average of two integers without overflow
+inline int64_t average(int64_t a, int64_t b) {
+    return (a & b) + ((a ^ b) >> 1);
+}
+
+// partitioning function for quicksort
+inline int64_t
+partition(int64_t *arr, int64_t left, int64_t right, int64_t pivot, int64_t &smallest, int64_t &biggest) {
+    int64_t *left_ptr = &arr[left];
+    int64_t *right_ptr = &arr[right];
+    while (left_ptr < right_ptr) {
+        smallest = (*left_ptr < smallest) ? *left_ptr : smallest;
+        biggest = (*left_ptr > biggest) ? *left_ptr : biggest;
+        if (*left_ptr > pivot) {
+            --right_ptr;
+            std::swap(*left_ptr, *right_ptr);
+        } else {
+            ++left_ptr;
+        }
+    }
+    return left_ptr - arr;
+}
+
+inline void insertion_sort(int64_t *arr, int64_t left, int64_t right) {
+    for (int64_t i = left + 1; i <= right; i++) {
+        int64_t key = arr[i];
+        int64_t j = i - 1;
+        while (j >= left && arr[j] > key) {
+            arr[j + 1] = arr[j];
+            j = j - 1;
+        }
+        arr[j + 1] = key;
+    }
+}
+
+// the core recursive quicksort function
+void qs_core(int64_t *arr, int64_t left, int64_t right, const int64_t pivot) {
+    if (right - left < 32) {
+        insertion_sort(arr, left, right);
+        return;
+    }
+
+    int64_t smallest = std::numeric_limits<int64_t>::max();
+    int64_t biggest = std::numeric_limits<int64_t>::min();
+    int64_t bound = partition(arr, left, right + 1, pivot, smallest, biggest);
+
+    if (smallest == biggest)
+        return;
+
+#pragma omp task final(bound - left < 10000)
+    qs_core(arr, left, bound - 1, average(smallest, pivot));
+    qs_core(arr, bound, right, average(pivot, biggest));
+}
+
+// wrapper for the quicksort function
+void min_max_quicksort(int64_t *arr, int64_t n, int num_threads = omp_get_max_threads()) {
+#pragma omp parallel num_threads(num_threads)
+#pragma omp single nowait
[…]

### 2024-11-10 Exam Assignments/Chapters/04/min_max_quicksort_array_sizes.cpp

@@ -1,0 +1,165 @@
+#include <iostream>
+#include <vector>
+#include <limits>
+#include <omp.h>
+#include <parallel/algorithm>
+#include <fstream>
+
+// compute the average of two integers without overflow
+inline int64_t average(int64_t a, int64_t b) {
+    return (a & b) + ((a ^ b) >> 1);
+}
+
+// partitioning function for quicksort
+inline int64_t
+partition(int64_t *arr, int64_t left, int64_t right, int64_t pivot, int64_t &smallest, int64_t &biggest) {
+    int64_t *left_ptr = &arr[left];
+    int64_t *right_ptr = &arr[right];
+    while (left_ptr < right_ptr) {
+        smallest = (*left_ptr < smallest) ? *left_ptr : smallest;
+        biggest = (*left_ptr > biggest) ? *left_ptr : biggest;
+        if (*left_ptr > pivot) {
+            --right_ptr;
+            std::swap(*left_ptr, *right_ptr);
+        } else {
+            ++left_ptr;
+        }
+    }
+    return left_ptr - arr;
+}
+
+inline void insertion_sort(int64_t *arr, int64_t left, int64_t right) {
+    for (int64_t i = left + 1; i <= right; i++) {
+        int64_t key = arr[i];
+        int64_t j = i - 1;
+        while (j >= left && arr[j] > key) {
+            arr[j + 1] = arr[j];
+            j = j - 1;
+        }
+        arr[j + 1] = key;
+    }
+}
+
+// the core recursive quicksort function
+void qs_core(int64_t *arr, int64_t left, int64_t right, const int64_t pivot) {
+    if (right - left < 32) {
+        insertion_sort(arr, left, right);
+        return;
+    }
+
+    int64_t smallest = std::numeric_limits<int64_t>::max();
+    int64_t biggest = std::numeric_limits<int64_t>::min();
+    int64_t bound = partition(arr, left, right + 1, pivot, smallest, biggest);
+
+    if (smallest == biggest)
+        return;
+
+#pragma omp task final(bound - left < 10000)
+    qs_core(arr, left, bound - 1, average(smallest, pivot));
+    qs_core(arr, bound, right, average(pivot, biggest));
+}
+
+// wrapper for the quicksort function
+void min_max_quicksort(int64_t *arr, int64_t n, int num_threads = omp_get_max_threads()) {
+#pragma omp parallel num_threads(num_threads)
+#pragma omp single nowait
[…]

### 2024-11-10 Exam Assignments/Chapters/04/plot.py

@@ -1,0 +1,31 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+
+# Read the CSV file into a pandas DataFrame
+# df = pd.read_csv('/home/jonas/Documents/University/5. Semester/Algorithm Engineering/Exam Assignments/Chapters/04/sorting_algorithms_running_times.csv')
+df = pd.read_csv('/home/jonas/Documents/University/5. Semester/Algorithm Engineering/Exam Assignments/Chapters/04/speedup_array.csv')
+
+print(df.columns)
+# Plotting the data
+plt.figure(figsize=(10, 6))
+
+# Loop through each column (except for the first one 'threads') and plot the data
+for column in df.columns[1:]:
+    plt.plot(df['Array Size'], df[column], label=f' {column}', marker='o')
+
+# Add titles and labels
+plt.xscale('log')
+plt.title('Speedup over std::sort at varying array sizes', fontsize=14)
+# plt.title('Speedup over std::sort (Time vs. Array Size)', fontsize=14)
+# plt.title('Performance of Sorting Algorithms (Time vs. Threads)', fontsize=14)
+plt.xlabel('Array Size (in MB)', fontsize=12)
+# plt.xlabel('Threads', fontsize=12)
+# plt.ylabel('Time (seconds)', fontsize=12)
+plt.ylabel('Speedup', fontsize=12)
+
+# Show grid and legend
+plt.grid(True)
+plt.legend()
+
+# Show the plot
+plt.show()

### 2024-11-18 Exam Assignments/Chapters/04/04.tex
Week 4

@@
 
+\bigskip
+
 ~\\
+% Let us now analyze the second two pictures, showing the running times at different array sizes.
+% The graphs are as expected, the running times of the parallel algorithms is better, even more so for very large array sizes, as in those instances we can fully employ the parallel power of the cpu cores, thus decreasing the running time significantly.
+% This is why the speedup gets better with increasing array size.
+% One might think that this speedup is exponential, but it is not (note the logarithmic scale!), it is rather approximately $(\text{Serial Running Time}) \cdot (1 - \frac{1}{12 \cdot E})$, where $E$ is the efficiency $E = \frac{\frac{\text{Serial Running Time}}{\text{Threads}}}{\text{Parallel Running Time}}$, and note that the number of threads on this machine is 12.
+% Note, that Serial Running Time is expected to grow $\Theta(n \cdot \log(n))$, and so should the speedup.
 Let us now analyze the second two pictures, showing the running times at different array sizes.
-The graphs are as expected, the running times of the parallel algorithms is better, even more so for very large array sizes, as in those instances we can fully employ the parallel power of the cpu cores, thus decreasing the running time significantly.
-This is why the speedup gets better with increasing array size.
-One might think that this speedup is exponential, but it is not (note the logarithmic scale!), it is rather approximately $(\text{Serial Running Time}) \cdot (1 - \frac{1}{12 \cdot E})$, where $E$ is the efficiency $E = \frac{\frac{\text{Serial Running Time}}{\text{Threads}}}{\text{Parallel Running Time}}$, and note that the number of threads on this machine is 12.
-Note, that Serial Running Time is expected to grow $\Theta(n \cdot \log(n))$, and so should the speedup.
+The graphs are as expected, the running times of the parallel algorithms is better, even more so for very large array sizes, as in those instances we can fully employ the parallel power of the cpu cores without much overhead relative to the running time, thus decreasing the running time significantly.
+This is why the speedup gets better with increasing array size initially.
+The speedup, however, start to plateau quickly at an array size of roughly 1 MB, as at this point the overhead of parallelization with respect to the running time stays constant (i.e.
+$\frac{\text{Time of Overhead}}{\text{Parallel Running Time}} = \text{const.}$):
 
+~\\
+Since
+\[
+    \text{speedup} = \frac{\text{Serial Running Time}}{\text{Parallel Running Time}} \ \ \ \ \ ,
+\]
+and
+\begin{align*}
+    \text{Parallel Running Time} &= \frac{\text{Serial Running Time}}{\text{Number of Threads}} + \text{Time of Overhead} \\ \\
+    &= \frac{\text{Serial Running Time}}{\text{Number of Threads}} + \alpha \cdot \text{Parallel Running Time} \\ \\
+    % \frac{\text{Time of Overhead}}{\text{Parallel Running Time}} \\
+    &= \frac{\frac{\text{Serial Running Time}}{Number of Threads}}{1-\alpha} \\ \\
+    &= \alpha' \cdot \text{Serial Running Time} \ \ \ \ \ ,
+\end{align*}
+we have
+\[
+    \text{speedup} = \frac{1}{\alpha'} = \text{const.} \ \ \ \ \ .
+\]
+
+% One might think that this speedup is exponential, but it is not (note the logarithmic scale!), it is rather approximately $(\text{Serial Running Time}) \cdot (1 - \frac{1}{12 \cdot E})$, where $E$ is the efficiency $E = \frac{\frac{\text{Serial Running Time}}{\text{Threads}}}{\text{Parallel Running Time}}$, and note that the number of threads on this machine is 12.
+% Note, that Serial Running Time is expected to grow $\Theta(n \cdot \log(n))$, and so should the speedup.
+
+\newpage
+\subsection{What every systems programmer should know about concurrency}
+\subsubsection{Why it is not enough to declare shared variables as volatile}
+By declaring variables as volatile in C and C++, we make the compiler read and write the variable from and to memory every time we access it.
+Cache coherency protocols then ensure that every local cache of the cpu cores has the same value of the shared variable.
+So everything should be good then, right?
+
+~\\
+Well, there is one problem: Modifying the shared variable isn't an atomic operation.
+Imagine two threads wanting to increment a shared counter.
+But as executing this increment operation takes multiple clock cycles, thread one starts reading the counter from memory, and a few ticks later thread two reads the same value of memory.
+Now, both threads increment this value and write it back to memory, so thread one does it first, which causes the cache of thread two to now have the updated counter variable in its cache, but the value it is operating on, which lies in its registers, is unchanged, so it will get the same result and write it back to memory.
+In the end, the counter increased by one, even though it got incremented two times, which is (in most cases) undesired and incorrect behavior.
+
+~\\
+\subsubsection{But what if my ISA doesn't support atomic access to my custom data structure / code section}
+We now learned that we have to use atomic variables to ensure correct parallel code execution.
+But what if the shared variable has no atomic variable data type like bool has with std::atomic\_bool?
+Or we want to synchronize an entire code section?
+As it turns out, we only need a few of these atomic operations and data types to implement generic synchronization functionality, as those can be used to built so called \em locks\em .
+
+~\\
+One well known example is a \em semaphore\em . It basically consists of an atomic counter, which represents how many threads may still access the shared resource (it is mostly either 0 or 1).
+If two threads want to access a shared resource, they have to first check if the semaphore has a value of one, and only if it does, decrement the value immediately and progress in their execution.
+If the semaphore is zero, the thread has to wait (which is mostly done by making it sleep).
+As you can see, "check if semaphore is one, and if so decrement it" must be an atomic operation for the semaphore to work.
+And this is exactly why modern ISA have instructions like \em compare-and-swap\em , which implement such operations atomically in hardware.
+There are also some other common instructions of this kind. They are combined under the term \em read-modify-write\em .
 \end{document}

### 2024-11-19 Exam Assignments/Code/05/Catch2/catch_tests_fizzbuzz.cpp
Week 5

@@ -1,0 +1,50 @@
+#include "catch.hpp"
+#include "fizzbuzz.h"
+
+TEST_CASE("Test positives", "[classic]")
+{
+   SECTION("Test all up to 10") {
+      REQUIRE(fizzbuzz(1) == "1"); 
+      REQUIRE(fizzbuzz(2) == "2");
+      REQUIRE(fizzbuzz(3) == "fizz");
+      REQUIRE(fizzbuzz(4) == "4");
+      REQUIRE(fizzbuzz(5) == "buzz");
+      REQUIRE(fizzbuzz(6) == "fizz");
+      REQUIRE(fizzbuzz(7) == "7");
+      REQUIRE(fizzbuzz(8) == "8");
+      REQUIRE(fizzbuzz(9) == "fizz");
+      REQUIRE(fizzbuzz(10) == "buzz");
+   }
+
+   SECTION("Test all multiples of 3 only up to 100") {
+      for (int i = 3; i <= 100; i+=3) {
+         if (i % 5) REQUIRE(fizzbuzz(i) == "fizz");
+      }
+   }
+
+   SECTION("Test all multiples of 5 only up to 100") {
+      for (int i = 5; i <= 100; i += 5) {
+         if (i % 3) REQUIRE(fizzbuzz(i) == "buzz");
+      }
+   }
+
+   SECTION("Test all multiples of 3 and 5 up to 100") {
+      for (int i = 15; i <= 100; i += 15) {
+         REQUIRE(fizzbuzz(i) == "fizzbuzz");
+      }
+   }
+}
+
+TEST_CASE("Test negatives", "[classic]")
+{
+   REQUIRE(fizzbuzz(-1) == "-1");
+   REQUIRE(fizzbuzz(-2) == "-2");
+   REQUIRE(fizzbuzz(-3) == "fizz");
+   REQUIRE(fizzbuzz(-4) == "-4");
+   REQUIRE(fizzbuzz(-5) == "buzz");
+   REQUIRE(fizzbuzz(-6) == "fizz");
+   REQUIRE(fizzbuzz(-7) == "-7");
+   REQUIRE(fizzbuzz(-8) == "-8");
+   REQUIRE(fizzbuzz(-9) == "fizz");
+   REQUIRE(fizzbuzz(-10) == "buzz");
+}

### 2024-11-19 Exam Assignments/Code/05/Catch2/fizzbuzz/include/fizzbuzz.h

@@ -1,0 +1,3 @@
+#include <string>
+
+std::string fizzbuzz(int const);

### 2024-11-27 Exam Assignments/Chapters/06/06.tex
Week 7

\documentclass[../../main.tex]{subfiles}

\begin{document}
    \section{Auto Vectorization}
    \subsection{Characteristics of SSE, AVX(2) and AVX-512}

    \begin{description}
        \item[SSE (Streaming SIMD Extensions):] ~\\ Introduced with x86, supports 128-bit vector registers, enabling SIMD operations on 4 single-precision floating-point or 2 double-precision values in parallel. Limited in width and functionality compared to later extensions. \\
        \item[AVX/AVX2 (Advanced Vector Extensions):] ~\\ Extends SIMD width to 256 bits, doubling parallelism compared to SSE. AVX introduced floating-point operations, while AVX2 added integer operations and support for gather instructions. \\
        \item[AVX-512:] ~\\ Widens SIMD to 512 bits, allowing operations on 16 single-precision floats or 8 double-precision values simultaneously. It includes more registers, mask registers for conditional operations, and broader instruction sets but increases power consumption and requires hardware support. \\
    \end{description}

    \bigskip
    \subsection{Impact of Memory Aliasing on Performance}
    Memory aliasing occurs when two or more pointers reference overlapping memory regions, making it difficult for the compiler or CPU to optimize memory accesses. This uncertainty can prevent reordering or parallelizing memory operations, resulting in stalls and reduced performance due to additional dependency checks and cache conflicts.

    \bigskip
    \subsection{Advantages of Unit Stride Memory Access}
    Unit stride (stride-1) memory access sequentially accesses adjacent memory locations, maximizing cache utilization and enabling efficient prefetching. Larger strides, like stride-8, result in less efficient cache usage, increased memory latency, and potential cache line thrashing due to non-contiguous data access patterns.

    \bigskip
    \subsection{When to Prefer Structure of Arrays}
    Structure of Arrays (SoA) is preferred when working with SIMD/vectorized operations or when the workload requires processing individual fields of a dataset independently. SoA provides better memory alignment and avoids unnecessary memory fetches, leading to improved cache efficiency and parallelism compared to Array of Structures (AoS).

\end{document}

### 2024-11-27 Exam Assignments/Chapters/07/07.tex
Week 7

\documentclass[../../main.tex]{subfiles}

\begin{document}
\newcommand{\drawcells}[3]{%
    \begin{tikzpicture}
        \foreach \j in {1,...,#1} {
            % Draw each cell
            \node[draw, fill=#3!20, minimum width = \linewidth/#1, minimum height=0.5cm, align=center] (cell\j) at (\j*\linewidth/#1, 0) {#2};
        }
    \end{tikzpicture}}
\section{Guided Vectorization and Data Types for Vector Intrinsics}
\subsection{Vectorization Clauses}
\begin{description}
    \item[aligned] ~\\ This clause tells the compiler that every element of an array $a$ is $n$-bit aligned if one uses \texttt{\#pragma opm simd aligned(a, n)}. The compiler employs this information to effectively vectorize the code section.
    \bigskip
    \item[safelen] ~\\ Imagine we had the following code:
    \begin{lstlisting}
a[0] = b[0];
for (int i = 1; i < n; i++) {
    a[i] = a[i-1] * (i % 2) + b[i];
}
    \end{lstlisting}
    We have a data dependency, but only every second iteration: iteration $i$, where $i$ is odd, depends on $i-1$. Iteration $i$ and $i+1$ are completely independent, however. Hence, we can vectorize the loop wath a vector length of $2$. To specify this, we use the \texttt{safelen} clause like so:
    \begin{lstlisting}
a[0] = b[0];
#pragma omp simd safelen(2)
for (int i = 1; i < n; i++) {
    a[i] = a[i-1] * (i % 2) + b[i];
}
    \end{lstlisting}   
    \bigskip
    \item[reduction] ~\\ This clause has the same functionality when paired with \texttt{simd} as when paired with \texttt{paraller for}: It will reduce a variable by an associative operation. But instead of reducing it over multiple threads, we use vectorization for reducing. I imagine that the reduced variable itself will be vectorized, so that the in loop we can employ vector operations, and at the end the elements in the resulting vector will get reduced.
\end{description}

\newpage
\subsubsection{AVX512 data types}
\resizebox{\textwidth}{!}{ % Scale to the width of the text
\begin{tabular}{ r c l }
    \texttt{\_\_m512} & \drawcells{16}{F}{blue} & \texttt{16x 32-bit float} \\
    \texttt{\_\_m512d} & \drawcells{8}{D}{green} & \texttt{8x 64-bit double} \\
    \texttt{\_\_m512i} & \drawcells{47}{}{red} & \texttt{64x 8-bit byte} \\
    \texttt{\_\_m512i} & \drawcells{32}{s}{red} & \texttt{32x 16-bit short} \\
    \texttt{\_\_m512i} & \drawcells{16}{i}{red} & \texttt{16x 32-bit integer} \\
    \texttt{\_\_m512i} & \drawcells{8}{ll}{red} & \texttt{8x 64-bit long} \\
    \texttt{\_\_m512i} & \drawcells{4}{dqw}{red} & \texttt{4x 128-bit quad} \\
    \texttt{\_\_m512i} & \drawcells{2}{qqw}{red} & \texttt{2x 256-bit integer} \\
    \texttt{\_\_m512i} & \drawcells{1}{dqqw}{red} & \texttt{1x 512-bit integer} \\
\end{tabular}
}

\bigskip
\subsubsection{Intel MMX}
In the paper \href{https://dl.acm.org/doi/pdf/10.1145/242857.242865}{Intel MMX for Multimedia PCs} the authors introduced the new MMX technology, which they integrated into their new cpus. It allows the cpu to employ vector instructions to increase throuphput and hence optimize running time of common computing tasks, such as rendering pixels for example. They illustrated the use cases in several examples, two of them being:

\begin{description}
    \item[Vectorizing Branches] Say we want overlay sprite onto a scene, i.e. we have a pixel array representing the sprite, and another one representing the scene (so far). The sprite array consists of color values. However, there is one constant value which represents a \em clear color\em , which represents the sprite not being visible at that point, and hence it should not get rendered onto the scene. The overlaying operation can the be performed by:
    \begin{lstlisting}
for i = 1 to Sprite_Line_size
if A[i] = clear_color then
        Out_frame[i] = C[i] else
        Out_frame[i] = A[i]
    \end{lstlisting}

    This is a typical example of branching inside a loop, where assignments have the following form:
    \[
        \texttt{a[i] = expr\_1 if cond else expr\_2}
    \]
    The idea is that we can vectorize this using vector istructions by:
    \[
        \texttt{a[i] = (expr\_1 $\land$ cond) $\lor$ (expr\_2 $\land$ $\lnot$cond)}
    \]
    Where \texttt{cond} is evaluated to \texttt{111\dots111} if true, and to \texttt{000\dots000} if false. The expressions should be easily vectorizable themselves, such that ultimately the entire statement can be vectorized efficiently.

    ~\\
    To see this in action, the authors provided a graphic for the example discussed earlier:

    \includegraphics[width=\textwidth]{Chapters/07/vectorizing_branches.png}

    ~\\
    In \em Phase 1\em , the mask is computed, i.e. the condition evaluated (the condition is that the color value equals the \texttt{clear\_color} constant).

    Next, in \em Phase 2\em , we evaluate the expressions (in this case it is omitted), and mask the results with the appropriate mask (\texttt{mask}, and $\lnot$\texttt{mask}). Finally, the two results get reduced by a logical or, yielding the final output.

    At the bottom of the graphic is a little assembly code section, emphasizing the fact that the operation can entirely be done with vector instructions.

    \bigskip
    \item[Matrix Transposition] Vector instructions can be fully utilized if the data is stored sequentially in memory. Hence, accessing rows of matrices can vectorized. But I we want to work on the columns of the matrix, it might be beneficial to transpose the matrix. This transposing can be vectorized as well by employing the so called \em unpack \em operation, which interlieves two operands word-wise, as can be seen in the picture:

    \includegraphics[width=\textwidth]{Chapters/07/unpack.png}

    ~\\
    To now transpose an entire matrix we go on like this: Imagine for the sake of simplicity that we have the following matrix $M$:
    \[ M = \begin{pmatrix}
        d3 & d2 & d1 & d0 \\
        c3 & c2 & c1 & c0 \\
        b3 & b2 & b1 & b0 \\
        a3 & a2 & a1 & a0
    \end{pmatrix}\]
    We now compute the last two rows of the matrix by firstly unpacking the last two entries of two adjacent rows, in this case $d$ and $c$, and $b$ and $a$, like shown in \em Phase 1 \em in the following graphic:

    \includegraphics[width=\textwidth]{Chapters/07/transpose.png}

    We now created two \em column slices \em of length two in every result of the unpack operations. Thus, we now need to unpack those among themselves, to get longer column slices, or in this case entire columns, as demonstrated in \em Phase 2\em .
    Once again, all the operations can be implemented with vector instructions, as hinted by the small assembly program at the bottom.
\end{description}
\end{document}

### 2024-11-28 Exam Assignments/Code/einspeed/.legacy/func.cpp
einspeed Python API

@@ -1,0 +1,9 @@
+// #include "func.h"
+
+// template <typename FROM, typename TO>
+// TO* cast_all(size_t _n, FROM *_a) {
+//     TO *_r = new TO[_n];
+//     for (size_t i = 0; i < _n; i++)
+//         _r[i] = static_cast<TO>(_a[i]);
+//     return _r;
+// }

### 2024-11-28 Exam Assignments/Code/einspeed/.legacy/parse_tensor.cpp

@@ -1,0 +1,41 @@
+// #include "parse_tensor.h"
+// #include "func.h"
+
+// // Mapping C++ types to NumPy type constants
+// template <typename T>
+// constexpr int _numpy_type();
+// template <>
+// constexpr int _numpy_type<double>() { return NPY_DOUBLE; }
+// template <>
+// constexpr int _numpy_type<float>() { return NPY_FLOAT; }
+// template <>
+// constexpr int _numpy_type<int>() { return NPY_INT; }
+// template <>
+// constexpr int _numpy_type<long>() { return NPY_LONG; }
+
+
+// template <typename T>
+// Tensor<T>* parse_tensor(PyObject *_t) {
+//     // Ensure the input is a NumPy array
+//     if (!PyArray_Check(_t)) {
+//         PyErr_SetString(PyExc_TypeError, "Expected a NumPy array");
+//         return nullptr;
+//     }
+
+//     // Cast to PyArrayObject and check type
+//     PyArrayObject* _a = reinterpret_cast<PyArrayObject*>(_t);
+//     if (PyArray_TYPE(_a) != _numpy_type<T>()) {
+//         // TODO: customize Error message
+//         PyErr_SetString(PyExc_TypeError, "Expected a NumPy array of type double");
+//         return NULL;
+//     }
+
+//     // Get number of dimensions and shape
+//     size_t ndim = (size_t) PyArray_NDIM(_a);
+//     // npy_intp* shape = PyArray_DIMS(array);  // Array of dimension sizes
+
+//     // Get a pointer to the data
+//     // double* data = static_cast<double*>(PyArray_DATA(array));
+
+//     return new Tensor<T>(PyArray_NDIM(_a), cast_all<npy_intp, size_t>(ndim, PyArray_DIMS(_a)), static_cast<T*>(PyArray_DATA(_a)));
+// }

### 2024-11-28 Exam Assignments/Code/einspeed/.legacy/setup.py

@@ -1,0 +1,34 @@
+# from setuptools import setup, Extension
+# import numpy as np
+
+# module = Extension(
+#     "mymodule",
+#     sources=["mymodule.c"],
+#     include_dirs=[np.get_include()]
+# )
+
+# setup(
+#     name="mymodule",
+#     version="1.0",
+#     ext_modules=[module]
+# )
+
+
+from setuptools import setup, Extension
+import numpy as np
+
+module = Extension(
+    "mymodule",
+    sources=["mymodule.cpp"],
+    include_dirs=[np.get_include()],
+    extra_compile_args=["-O3", "-fopenmp"],  # Optimization and OpenMP support
+    extra_link_args=["-fopenmp"],            # Link OpenMP libraries
+)
+
+setup(
+    name="mymodule",
+    version="1.0",
+    ext_modules=[module],
+)
+
+

### 2024-11-28 Exam Assignments/Code/einspeed/.legacy/tensor.cpp

@@ -1,0 +1,11 @@
+// #include "tensor.h"
+
+// template <typename T>
+// Tensor<T>::Tensor(size_t ndim, size_t* shape, T* data)
+//     : ndim(ndim), shape(shape), data(data) {}
+
+// template <typename T>
+// Tensor<T>::~Tensor() {
+//     delete[] shape;
+//     // delete[] data; Don't delete data as it belongs to Python
+// }

### 2024-11-28 Exam Assignments/Code/einspeed/build/src/CMakeFiles/einspeed.dir/compiler_depend.ts

@@ -1,0 +1,2 @@
+# CMAKE generated file: DO NOT EDIT!
+# Timestamp file for compiler generated dependencies management for einspeed.

### 2024-11-28 Exam Assignments/Code/einspeed/einspeed.py

@@ -1,0 +1,12 @@
+import einspeed
+import numpy as np
+
+# Pass and process a large matrix
+matrix = np.array([[1.0, 2.0, 3.0], [4.0, 5.0, 6.0], [7.0, 8.0, 9.0]], dtype=np.float64)
+einspeed.einsumd("", matrix, matrix)
+
+matrix = np.array([[1.0, 2.0, 3.0], [4.0, 5.0, 6.0], [7.0, 8.0, 9.0]], dtype=np.float32)
+einspeed.einsum("", matrix, matrix)
+
+print("Updated matrix:")
+print(matrix)

### 2024-11-28 Exam Assignments/Code/einspeed/include/func.h

@@ -1,0 +1,14 @@
+#ifndef FUNC_H
+#define FUNC_H
+
+#include <cstdlib>
+
+template <typename FROM, typename TO>
+TO* cast_all(size_t _n, FROM *_a) {
+    TO *_r = new TO[_n];
+    for (size_t i = 0; i < _n; i++)
+        _r[i] = static_cast<TO>(_a[i]);
+    return _r;
+}
+
+#endif

### 2024-11-28 Exam Assignments/Code/einspeed/include/parse_tensor.h

@@ -1,0 +1,44 @@
+#ifndef PARSE_TENSOR_H
+#define PARSE_TENSOR_H
+
+#include <Python.h>
+#include <numpy/arrayobject.h>
+#include "tensor.h"
+#include "func.h"
+
+// Mapping C++ types to NumPy type constants
+template <typename T>
+constexpr int _numpy_type();
+template <>
+constexpr int _numpy_type<double>() {return NPY_DOUBLE;}
+template <>
+constexpr int _numpy_type<float>() {return NPY_FLOAT;}
+template <>
+constexpr int _numpy_type<int>() {return NPY_INT;}
+template <>
+constexpr int _numpy_type<long>() {return NPY_LONG;}
+
+
+template <typename T>
+Tensor<T>* parse_tensor(PyObject *_t) {
+    // Ensure the input is a NumPy array
+    if (!PyArray_Check(_t)) {
+        PyErr_SetString(PyExc_TypeError, "Expected a NumPy array");
+        return nullptr;
+    }
+
+    // Cast to PyArrayObject and check type
+    PyArrayObject* _a = reinterpret_cast<PyArrayObject*>(_t);
+    if (PyArray_TYPE(_a) != _numpy_type<T>()) {
+        // TODO: customize Error message
+        PyErr_SetString(PyExc_TypeError, "Expected a NumPy array of type double");
+        return NULL;
+    }
+
+    // Get number of dimensions and shape
+    size_t ndim = PyArray_NDIM(_a);
+
+    return new Tensor<T>(PyArray_NDIM(_a), cast_all<npy_intp, size_t>(ndim, PyArray_DIMS(_a)), static_cast<T*>(PyArray_DATA(_a)));
+};
+
+#endif

### 2024-11-28 Exam Assignments/Code/einspeed/include/tensor.h

@@ -1,0 +1,22 @@
+#ifndef TENSOR_H
+#define TENSOR_H
+
+#include <cstddef> // For size_t
+
+template <typename T>
+class Tensor {
+public:
+    size_t ndim;    // Number of dimensions
+    size_t* shape;  // Pointer to array defining the shape
+    T* data;        // Pointer to the actual data
+
+    Tensor(size_t ndim, size_t* shape, T* data)
+        : ndim(ndim), shape(shape), data(data) {}
+
+    ~Tensor() {
+        delete[] shape;
+        // delete[] data; Don't delete data as it belongs to Python
+    }
+};
+
+#endif

### 2024-12-10 Exam Assignments/Chapters/08/08.tex
week 08

\documentclass[../../main.tex]{subfiles}

\begin{document}
    \section{Vector Intrinsics and Instruction-Level Parallelism}
    \subsection{Analyzing slide 14}
    Slide 14 explains how we can utilize vector instrinsics to transpose 8x8 matrices. Here is the code:
    \begin{lstlisting}
// transpose 8 x 8 matrix with intrinsics
inline void transpose_8x8_int_intrinsics(__m256i *vecs) {
    __m256 *v = reinterpret_cast<__m256 *>(vecs);
    __m256 a = _mm256_unpacklo_ps(v[0], v[1]);
    __m256 b = _mm256_unpackhi_ps(v[0], v[1]);
    __m256 c = _mm256_unpacklo_ps(v[2], v[3]);
    __m256 d = _mm256_unpackhi_ps(v[2], v[3]);
    __m256 e = _mm256_unpacklo_ps(v[4], v[5]);
    __m256 f = _mm256_unpackhi_ps(v[4], v[5]);
    __m256 g = _mm256_unpacklo_ps(v[6], v[7]);
    __m256 h = _mm256_unpackhi_ps(v[6], v[7]);
    auto tmp = _mm256_shuffle_ps(a, c, 0x4E);
    a = _mm256_blend_ps(a, tmp, 0xCC);
    c = _mm256_blend_ps(c, tmp, 0x33);
    tmp = _mm256_shuffle_ps(b, d, 0x4E);
    b = _mm256_blend_ps(b, tmp, 0xCC);
    d = _mm256_blend_ps(d, tmp, 0x33);
    tmp = _mm256_shuffle_ps(e, g, 0x4E);
    e = _mm256_blend_ps(e, tmp, 0xCC);
    g = _mm256_blend_ps(g, tmp, 0x33);
    tmp = _mm256_shuffle_ps(f, h, 0x4E);
    f = _mm256_blend_ps(f, tmp, 0xCC);
    h = _mm256_blend_ps(h, tmp, 0x33);
    v[0] = _mm256_permute2f128_ps(a, e, 0x20);
    v[1] = _mm256_permute2f128_ps(c, g, 0x20);
    v[2] = _mm256_permute2f128_ps(b, f, 0x20);
    v[3] = _mm256_permute2f128_ps(d, h, 0x20);
    v[4] = _mm256_permute2f128_ps(a, e, 0x31);
    v[5] = _mm256_permute2f128_ps(c, g, 0x31);
    v[6] = _mm256_permute2f128_ps(b, f, 0x31);
    v[7] = _mm256_permute2f128_ps(d, h, 0x31);
}
    \end{lstlisting}
Let us analyze how we can calculate a specific column, say $v[0]$. It depends on the variables $a$ and $e$, so let us track them:

~\\
We use
\begin{lstlisting}
    __m256 a = _mm256_unpacklo_ps(v[0], v[1]);
\end{lstlisting}
in the first step (similar for $e$). Here, we unpack the first two rows, which means we interleave them (compare to last week). \em lo \em specifies that we focus on the \em lower order bits \em ($a[\bullet][0], a[\bullet][1], a[\bullet][4], a[\bullet][5]$). Hence, after this operation, $a$ and $e$ contain the following matrix elements:
\begin{lstlisting}
    a = {a00, a10, a01, a11, a04, a14, a05, a15};
    e = {a40, a50, a41, a51, a44, a54, a45, a55};
\end{lstlisting}

~\\
For the next operation, we calculate $tmp$, for which we need variable $c$ in case of $a$, so let us quickly do that:
\begin{lstlisting}
    c = {a20, a30, a21, a31, a24, a34, a25, a35};
\end{lstlisting}
Now we can calculate $tmp$ with:
\begin{lstlisting}
    auto tmp = _mm256_shuffle_ps(a, c, 0x4E);
\end{lstlisting}

~\\
We have
\begin{verbatim}
    0x4E = 0b01001110 (= 1 0 3 2)     ,
\end{verbatim}
and hence $tmp$ becomes:
\begin{lstlisting}
    tmp = {a01, a11, a20, a30, a05, a15, a24, a34};
\end{lstlisting}

~\\
Let us visualize $tmp$:

\begin{table}[h]
    \begin{tabular}{c|c|c|c|c|c}
        &1&&&&5\\
        \hline
        &2&&&&6\\
        \hline
        3&&&&7&\\
        \hline
        4&&&&8&
    \end{tabular}
\end{table}

~\\
Next, we calculate $a$ using:
\begin{lstlisting}
     a = _mm256_blend_ps(a, tmp, 0xCC);
\end{lstlisting}
Note that
\begin{verbatim}
    0xCC = 0b11001100     .
\end{verbatim}
Hence, we get after this blending operation, which selects elements based on the mask:
\begin{lstlisting}
    a = {a00, a10, a20, a30, a04, a14, a24, a34};
    e = {a40, a50, a60, a70, a44, a54, a64, a74};
\end{lstlisting}

~\\
From here, we can see that we are almost done, $a$ and $e$ look like they almost contain one column, in fact the contain two columns "cut in the middle". Hence, after a final permutation we should be done calculating a column. To this end, we use \texttt{\_mm256\_permute2f128\_ps} and specify that we want to use the lower bits of both $a$ and $e$ by setting the mask to \texttt{CC}:
\begin{lstlisting}
    v[0] = _mm256_permute2f128_ps(a, e, 0x20);
    v[0] = {a00, a10, a20, a30, a40, a50, a60, a70};
\end{lstlisting}

~\\
This explanation was not very detailed, but it gives a guideline to get a \em feel \em for the algorithm, which might improve one's understanding of this algorithm.

\bigskip
\subsubsection{Latency and Throughput}
\begin{description}
    \item[Latency] measures the time it takes for an intrinsic function to produce its result once it starts executing, typically expressed in clock cycles.
    \item[Throughput] indicates the rate at which the function can be executed, showing how many operations can be completed per unit of time (e.g., cycles per instruction) when pipelined.
\end{description}

~\\
Obviously, the lower the latency, and the the higher the throughput, the better. In certain situations, however, one metric is much more decisive than the other:

~\\
If one writes dependent code, i.e. the next line of code may only be executed after the current one because of a dependency, the latency of the operation will be the main factor for performance, since we want our result as quickly as possible to procede with the execution.

~\\
On the other hand, if we write independent code which also uses many repeated operations, throughput will be key, as it determines how many clock cycles we approximately need to finish all calculations.
%  \\ ($no\_clock\_cycles \approx \frac{no\_operations}{throughput\_of\_operation}$ (Note that the number of operations does not directly correlate with the number of instructions in code, as certain code instructions need more cpu instructions)).
In this case, the latency has a neglectable influence on the perfoncmance, especially if the number of operations grows bigger.

\bigskip
\subsubsection{Analyzing Vectorized Hash Tables Across CPU Architectures}
In this paper the authors explained how to employ SIMD instructions to efficiently implement hash tables. Hash tables are quite diverse, as different implementation approaches can be used. One common implementation employs linear probing to search through the key array to find the specified key, so that the desired value can be returned for reading. Here is how we might implement this using vector instructions:

\begin{description}
    \item[Vectorized Linear Probing] The keys should be integer values, so that they can fit comfortably into the vector registers. Linear probing starts by checking the index specified by the hash value of the current key first, and then keeps going one step at a time until we found our key. The "key" idea is to simply load multiple conscutive keys into one vector register, and to compare them all against the search key. If one comparision is successful, we can calculate the index and return tha value at that index. Otherwise, we keep probing with the next set of keys.

    The authors also mentioned that there are subtle challenges, as vector instructions require the data in memory to be correctly aligned (by their register size). When we start probing at the hash index, however, we are almost certainly not at an aligned address. Hence, we should ensure to "align ourselves", for example by loading the aligned memory region containing the first search index into a vector register.

    \item[Vectorized Fingerprinting] As mentioned, vectorized linear probing require integer keys. So what could we do if wanted to use strings instead? The "key" idea is to calculate "keys" for the keys themself (quite many keys around here). One can think of these keys as hash values (though not necessarily the same as the real hashes) of the keys. To avoid confusion, they are called \em fingerprints \em instead, and are quite short, in the paper the authors used fingerprints of size 8 or 16 bits. Using this technique, we can use vectorized linear probing on these fingerprints as described earlier, with the only difference being that we might run into fingerprint-conflicts. Thus, after a successful match of fingerprints, we have to then compare the actual keys to determine an actual hit.
\end{description}
\end{document}

### 2024-12-10 Exam Assignments/Chapters/03/03.tex
Week 09

@@
 The outer most loop cannot be parallelized, as their iterations are all strictly dependent on the previous one.
-This is why the implicit barrier in line 13 is neccessary.
+This is why the implicit barrier in line 13 is necessary.
 It also shouldn't be that big of a bottleneck, as the threads should reach it almost simultaneously.
@@
 It runs very quickly and calculates $\pi$ very precisely.
-The running time on my particular machine was 0.0313137 s, in comparision to the previous version with a running time of 0.117771 s.
+The running time on my particular machine was 0.0313137 s, in comparison to the previous version with a running time of 0.117771 s.
 

### 2024-12-10 Exam Assignments/Chapters/07/07.tex
Week 09

@@
 Iteration $i$ and $i+1$ are completely independent, however.
-Hence, we can vectorize the loop wath a vector length of $2$.
+Hence, we can vectorize the loop with a vector length of $2$.
 To specify this, we use the \texttt{safelen} clause like so:
@@
     \]
-    The idea is that we can vectorize this using vector istructions by:
+    The idea is that we can vectorize this using vector instructions by:
     \[
@@
 But I we want to work on the columns of the matrix, it might be beneficial to transpose the matrix.
-This transposing can be vectorized as well by employing the so called \em unpack \em operation, which interlieves two operands word-wise, as can be seen in the picture:
+This transposing can be vectorized as well by employing the so called \em unpack \em operation, which interleaves two operands word-wise, as can be seen in the picture:
 

### 2024-12-10 Exam Assignments/Chapters/08/08.tex
Week 09

@@
 %  \\ ($no\_clock\_cycles \approx \frac{no\_operations}{throughput\_of\_operation}$ (Note that the number of operations does not directly correlate with the number of instructions in code, as certain code instructions need more cpu instructions)).
-In this case, the latency has a neglectable influence on the perfoncmance, especially if the number of operations grows bigger.
+In this case, the latency has a neglectable influence on the performance, especially if the number of operations grows bigger.
 
@@
 Linear probing starts by checking the index specified by the hash value of the current key first, and then keeps going one step at a time until we found our key.
-The "key" idea is to simply load multiple conscutive keys into one vector register, and to compare them all against the search key.
-If one comparision is successful, we can calculate the index and return tha value at that index.
+The "key" idea is to simply load multiple consecutive keys into one vector register, and to compare them all against the search key.
+If one comparison is successful, we can calculate the index and return tha value at that index.
 Otherwise, we keep probing with the next set of keys.

### 2024-12-10 Exam Assignments/Chapters/09/09.tex
Week 09

\documentclass[../../main.tex]{subfiles}

\begin{document}
\section{Cache and Main Memory}
\subsection{How do bandwidth-bound computations differ from
compute-bound computations?}
The difference is quite obvious. When running a program, it will run in a finite amount of time (hopefully). But why doesn't it terminate instantly? What a silly question, because a computer has to execute one instruction at a time, and each instruction takes a little bit of time to complete.

Now, we differentiate between two kinds of operations: Usual cpu instructions and memory operations. The reason for this is that memory operations are relatively much more time expensive, and the recent hardware trends make this gap grow bigger. Hence, we would like to know whether our program is bottlenecked by these expensive memory operations, which we very much would like to avoid, as we want to utilize our cpu with useful instructions instead.

To this end, we call computations \em memory-bound \em iff they are bottlenecked by the latency of memory operations. This, of course, is dependent on many factors, such as hardware and so on.

If our program is not memory-bound, it still executes not instantly, as it is bound by the execution speed of the cpu. We have a \em compute-bound\em . In a sense, every program or operation experiences a compute-bound. Hence, it is more useful when we quantify this bound, by analyzing the cpu frequency and the execution speed of individual operations, so that we can estimate the program's execution time.

\bigskip
\subsection{Influence of temporal locality and spatial locality on program performance}
If our program uses a lot of data, and thus faces a memory-bound, we have to intentionally engineer a program that employs good temporal as well spatial locality.

This is, because a lot of data will read from and written to memory, which, as explained, is very time consuming. Luckily, computer engineers introduced caches to this end, which store some data of main memory, and provide it to the program with faster access times. The only problem will be if the caches or, more precisely, their hash bucket are full. Because caches are usually not \em fully-associative\em , which means that if we have a cache line in memory, we can not write it to any cache line in the cache as we like. In fact, we only have $N$ cache lines available in an \em N-way associative \em cache.

As we can see, caches are very useful, but we have to understand their behavior in order to specifically tailor a program to effectively utilize them. In fact, a \em cache-aware \em algorithm employs good temporal as well as spatial locality. But what does that mean?

\begin{description}
    \item[Temporal locality] describes the concept that recently accessed memory has a higher probability to be accessed again.
    \item[Spatial locality] on the other hand is used to refer to the concept that memory, which is spatially close to recently accessed memory, has a higher probability of being accessed themself.
\end{description}

But why is all that useful? Well, it is due to the \em replacement policy \em of caches. As explained, caches will run full, and at that instance, they will have to evict a cache line. But how do they decide? Well, one common replacement policy is \em LRU (Least Recently Used)\em , which means that they choose the cache line whose access is the most far away in time. This is why temporal locality comes in handy, as it increases the chances that our data lies in cache, and thus reduces the risk of a cache miss.

Finally, spatial locality is useful, because, as described, caches hold entire cache lines, which commonly hold 64 bytes of memory. Hence, if the data we access is closely clustered, we reduce the potential number of cache lines required, reducing the space requirement in the cache, and we will have more accesses to our cache lines. All of this contributes to a lower risk of a cache miss, which ultimately reduces the running time of our program.

\bigskip
\subsection{Cache Associativity}
On the slides 15 and 17 we hear about caches being \em N-way associative\em . Let's investigate on this topic further. Let's analyze the following picture (from Wikipedia):

\begin{figure}[ht]
    \centering
    \includegraphics[width=\linewidth]{Chapters/09/Set-Associative_Cache_Snehal_Img.png}
    \caption~{2-way associative cache \footnotemark}
    \label{fig:example}
\end{figure}
\footnotetext{Image source: \href{https://en.wikipedia.org/wiki/Cache\_placement\_policies\#/media/File:Set-Associative\_Cache\_Snehal\_Img.png}{https://en.wikipedia.org/wiki/Cache\_placement\_policies\#/media/File:Set-Associative\_Cache\_Snehal\_Img.png}}

~\\
Since this cache is 2-way associative, two consecutive cache lines form a group. For example, cache line 0 and 1 both belong to group 0. If we both know the $cache\_line\_size$ and the $cache\_size$, we can calculate $no\_cache\_lines = \frac{cache\_size}{cache\_line\_size}$. Now, we can calculate the number of sets these cache lines form because of associativity: $no\_sets = \frac{no\_cache\_lines}{N}$.

~\\
Now, here comes the clever part. When determining the memory address of the stored data in the cache, we can save 5 bits in this example, as this information is stored in the index of the block. More precisely, we have $no\_index\_bits = \log_2(no\_sets)$. The offset determines the exact location inside the cache line, and hence $no\_offset\_bits = \log_2(cache\_line\_size)$.

~\\
Finally, the remaining bits that are needed to uniquely identify the memory address are grouped into the \em Tag\em . How many bits do we need exactly? Well, here is one explanation: In total, there are $no\_address\_bits = \log_2(memory\_size)$ bits needed. But we already have \em index \em and \em offset \em bits, so there are $no\_tag\_bits = no\_address\_bits - (no\_index\_bits + no\_offset\_bits)$ bits remaining.

\bigskip
\subsection{An Overview of Cache Optimization Techniques and Cache-Aware Numerical Algorithms}
In this paper the author explains the importance of cache-aware programs, as compilers won't help us much in this regard. Hence, let's investigate on some practical examples where we can improve temporal and / or spatial locality.

\begin{description}
    \item[Loop Blocking] This is best understood by an example. Say we have the following code:

    \includegraphics[width=\linewidth]{Chapters/09/loop_blocking.png}

    Both matrices $a$ and $b$ are stored in row-order fashion. Hence, when running the left algorithm, $a$ has good spatial locality, whereas $b$'s is horrendous, as it leads to a cache miss on every access. The idea is now to find a tradeoff between these two. Let us analyze the access pattern of the two algorithms:

    \includegraphics[width=\linewidth]{Chapters/09/access_pattern.png}

    Note that walking along the rows is good for spatial locality, and also note that for b the graphic should be flipped along $y=x$.

    So why is the right access pattern better? Note the following: for $a$, the spatial locality does get a bit worse, but only in $O(1)$ fashion, as now we have to maybe load two more cache lines into the cache, not really a big deal. But what about $b$? Well, it basically divides its cache misses by a factor of 2 (in case of a block size of 2)! Thus, the overall cache misses will be reduced.

    \item[Array Padding] This one is rather specific, but it is quite useful to keep it in the back of one's mind. Imagine for the sake of argument that we have cache which can hold 1024 doubles in total, and has a 1-way associative cache (we will now see why this is maybe a bit too aggressive). Next, consider this algorithm:

    \includegraphics[width=\linewidth]{Chapters/09/array_padding.png}

    It just so happens that in this instance both array $a$ and $b$ will map to the same cache lines (compare to my previous explanation). But, as there is only one slot in the cache for a specific cache line with a certain index, we will load a cache line of array $a$, right after that we will evict it for $b$, and then again for $a$ and so on. In the end, we will have a cache miss on every access, yikes. If only the cache line indices of $a$ and $b$ wouldn't align. Well, exactly this is the idea behind the algorithm on the right: We simply occupy some padding space in-between the two arrays, which the cache will be grateful for.

    This example was obviously very specific, but we can see that we might have to be a bit more careful if we had large arrays, maybe multiple of them, which in memory are separated by a large power of 2.
\end{description}
\end{document}

### 2024-12-14 Exam Assignments/Code/einspeed/.legacy/tensor.h
batch_matmul, reshaping and transposing target

@@ -1,0 +1,288 @@
+#ifndef TENSOR_H
+#define TENSOR_H
+
+#include <Python.h>
+#include <numpy/arrayobject.h>
+// #include <cstddef>
+#include <vector>
+#include <algorithm>
+#include "numpy_types.h"
+// #include <stdexcept>
+#include <cstring> // For memcpy
+
+#include <iostream>
+
+template <typename T>
+class Tensor {
+public:
+    size_t ndim = 0;          // Number of dimensions
+    std::vector<size_t> shape = nullptr;  // Pointer to array defining the shape
+    T* data = nullptr;        // Pointer to the actual data
+
+    Tensor(size_t ndim, std::vector<size_t> shape, T* data)
+        : ndim(ndim), shape(shape), data(data) {};
+    Tensor(size_t ndim, size_t *shape, T *data) 
+        : ndim(ndim), shape(shape, shape + ndim), data(data) {}
+    Tensor() {}
+    
+    Tensor(Tensor<T> &&other)
+        : ndim(other.ndim), shape(other.shape), data(other.data) {
+        // Nullify the moved-from object to avoid double deletion
+        other.ndim = 0;
+        other.shape = nullptr;
+        other.data = nullptr;
+    }
+
+    Tensor<T>& operator=(Tensor<T> &&other) {
+        if (this != &other) {  // Prevent self-assignment
+            // Clean up any existing resources (if necessary)
+            delete[] shape;
+            // Don't delete data as it belongs to Python or is managed elsewhere.
+
+            // Move the resources
+            ndim = other.ndim;
+            shape = other.shape;
+            data = other.data;
+
+            // Nullify the moved-from object
+            other.ndim = 0;
+            other.shape = nullptr;
+            other.data = nullptr;
+        }
+        return *this;
+    }
+
+
+    ~Tensor() {
+        std::cout << "DESTROYING TENSOR:\n";
+        // this->print();
+        if (this->shape) delete[] this->shape; // It is okay to delete this, since when casting to PyObject, we will create a copy (which we will later also delete though)
+        // if (this->data) delete[] this->data; // Don't delete data as it belongs to Python
+    }
+
+    PyObject* to_PyObject() const {
+        // Initialize NumPy (if not already done)
+        if (!PyArray_API) {
+            import_array(); // Necessary to initialize NumPy C API
+        }
+
+        // Create a NumPy array
+        npy_intp* np_shape = new npy_intp[ndim];
+        for (size_t i = 0; i < ndim; ++i) {
[…]

### 2024-12-14 Exam Assignments/Code/einspeed/.legacy/tests/catch_tests_blas.cpp

@@ -1,0 +1,63 @@
+#define CATCH_CONFIG_MAIN
+#include "catch.hpp"
+#include "tensor.h"
+#include "blas.h"
+
+TEST_CASE("Batch Matrix Multiplication", "[batch_matmul]") {
+    REQUIRE(true);
+    // SECTION("Basic batch multiplication with 2x2 matrices") {
+    //     // Create two tensors with batch size 2 and matrices of size 2x2
+    //     size_t shape[3] = {2, 2, 2}; // Batch size 2, 2x2 matrices
+    //     double data_lhs[] = {1, 2, 3, 4,   // First matrix in batch
+    //                          5, 6, 7, 8};  // Second matrix in batch
+
+    //     double data_rhs[] = {9, 10, 11, 12,  // First matrix in batch
+    //                          13, 14, 15, 16}; // Second matrix in batch
+
+    //     Tensor<double> lhs(3, shape, data_lhs);
+    //     Tensor<double> rhs(3, shape, data_rhs);
+
+    //     // Expected output: Batch of multiplied matrices
+    //     double expected_data[] = {
+    //         31, 34,  71, 78,  // First batch result
+    //         149, 162, 221, 240 // Second batch result
+    //     };
+    //     Tensor<double> expected_result(3, shape, expected_data);
+
+    //     // Perform batch matrix multiplication
+    //     auto result = batch_matmul(lhs, rhs);
+
+    //     // Compare the resulting tensor with the expected tensor
+    //     REQUIRE(result.ndim == expected_result.ndim);
+    //     REQUIRE(std::equal(result.shape, result.shape + 3, expected_result.shape));
+    //     REQUIRE(std::equal(result.data, result.data + 8, expected_result.data));
+    // }
+
+    // SECTION("Invalid batch dimensions should throw an exception") {
+    //     // Batch size mismatch between tensors
+    //     size_t shape_lhs[3] = {2, 2, 2};
+    //     size_t shape_rhs[3] = {3, 2, 2}; // Batch size mismatch
+    //     double data[] = {1, 2, 3, 4, 5, 6, 7, 8};
+
+    //     Tensor<double> lhs(3, shape_lhs, data);
+    //     Tensor<double> rhs(3, shape_rhs, data);
+
+    //     // Expect an exception due to mismatched batch dimensions
[…]

### 2024-12-14 Exam Assignments/Code/einspeed/einspeed.py

@@ -5,11 +5,13 @@ import numpy as np
 matrix = np.array([[1.0, 2.0, 3.0], [4.0, 5.0, 6.0], [7.0, 8.0, 9.0]], dtype=np.float64)
 
 # matrix = np.array([[1.0, 2.0, 3.0], [4.0, 5.0, 6.0], [7.0, 8.0, 9.0]], dtype=np.float32)
-result = einspeed.einsum("ij,kj->ik", matrix, matrix)
+result = einspeed.einsum("ai,jb->ij", matrix, matrix)
+
+print(np.einsum("ai,jb->ij", matrix, matrix))
 
 print("result:")
 print(result)
-print(result.shape)
-print("Updated matrix:")
-print(matrix)
-print(matrix.shape)
+# print(result.shape)
+# print("Updated matrix:")
+# print(matrix)
+# print(matrix.shape)

### 2024-12-14 Exam Assignments/Code/einspeed/include/blas.h

@@ -1,0 +1,55 @@
+#ifndef BLAS_H
+#define BLAS_H
+
+#include "tensor.h"
+
+template <typename T>
+Tensor<T> batch_matmul(const Tensor<T>& lhs, const Tensor<T>& rhs) {
+    // Check dimensions
+    if (lhs.ndim != 3 || rhs.ndim != 3) {
+        throw std::invalid_argument("Both tensors must have 3 dimensions for batch matrix multiplication.");
+    }
+    if (lhs.shape[0] != rhs.shape[0]) {
+        throw std::invalid_argument("Batch sizes must match for batch matrix multiplication.");
+    }
+    if (lhs.shape[2] != rhs.shape[1]) {
+        throw std::invalid_argument("Inner matrix dimensions must match for batch matrix multiplication.");
+    }
+
+    // Shapes
+    size_t batch_size = lhs.shape[0];
+    size_t rows = lhs.shape[1];
+    size_t cols = rhs.shape[2];
+    size_t inner_dim = lhs.shape[2]; // Same as rhs.shape[1]
+
+    // Allocate the result tensor
+    size_t result_shape[] = {batch_size, rows, cols};
+    T* result_data = new T[batch_size * rows * cols];
+
+    // Perform batch matrix multiplication
+    for (size_t b = 0; b < batch_size; ++b) {
+        // Get pointers to the current batch matrices
+        const T* lhs_batch = lhs.data + b * rows * inner_dim;
+        const T* rhs_batch = rhs.data + b * inner_dim * cols;
+        T* result_batch = result_data + b * rows * cols;
+
+        // Multiply the two matrices for the current batch
+        for (size_t i = 0; i < rows; ++i) {
+            for (size_t j = 0; j < cols; ++j) {
+                for (size_t k = 0; k < inner_dim; ++k) {
+                    result_batch[i * cols + j] += lhs_batch[i * inner_dim + k] * rhs_batch[k * cols + j];
+                }
+            }
+        }
+    }
+
+    std::cout << "HERE";
+    // Return the result tensor
+    auto t = Tensor<T>(3, result_shape, result_data);
+    std::cout << t.shape[0] << t.shape[1] << t.shape[2];
+    // t.print();
+    return t;
+    // return Tensor<T>(3, result_shape, result_data);
+}
+
+#endif

### 2024-12-14 Exam Assignments/Code/einspeed/include/compute_einsum.h

@@ -1,27 +1,30 @@
 #ifndef COMPUTE_EINSUM_H
 #define COMPUTE_EINSUM_H
+#define DEBUG
 
 #include <string>
 #include <unordered_set>
 #include <stdexcept>
-#include "tensor.h"
+// #include "tensor.h"
+#include "blas.h"
 
 // ******************************************
 // The core functionality is implemented HERE
 // ******************************************
 
-template <typename T>
-PyObject* compute_einsum(const char * const s, const Tensor<T> &&lhs_tensor, const Tensor<T> &&rhs_tensor) {
-    if (!is_valid_einsum_expression(s, lhs_tensor, rhs_tensor)) Py_RETURN_NONE;
-    auto target_current_column_identifiers = transform_matrices(s, lhs_tensor, rhs_tensor);
+// template <typename T>
+// PyObject* _compute_einsum(const char * const s, const Tensor<T> &&lhs_tensor, const Tensor<T> &&rhs_tensor) {
+//     if (!is_valid_einsum_expression(s, lhs_tensor, rhs_tensor)) Py_RETURN_NONE;
+//     // auto target_current_column_identifiers = transform_matrices(s, lhs_tensor, rhs_tensor);
+//     return transform_matrices<T>(s, lhs_tensor, rhs_tensor).to_PyObject();
 
-    if constexpr (std::is_same_v<T, double>) {
-        lhs_tensor.data[0] = 42;
-    }
-    auto perm = std::vector<size_t>({1, 0});
-    auto shape = std::vector<size_t>({9});
-    return lhs_tensor.transpose(perm).reshape(shape).to_PyObject();
-}
+//     if constexpr (std::is_same_v<T, double>) {
+//         // lhs_tensor.data[0] = 42;
+//     }
+//     auto perm = std::vector<size_t>({1, 0});
+//     auto shape = std::vector<size_t>({9});
+//     return lhs_tensor.transpose(perm).reshape(shape).to_PyObject();
+// }
 
 
 std::vector<size_t> column_identifiers_like(const std::vector<size_t> &target_column_identifiers, const std::string &target_identifier_string, const std::string &origin_identifier_string) {
@@ -36,7 +39,7 @@ std::vector<size_t> column_identifiers_like(const std::vector<size_t> &target_co
 
 // TODO: reread the entire comments to check for errors (since I changed all occurences of target to target_string, conflicting maybe with target_tensor)
 template <typename T>
[…]

### 2024-12-14 Exam Assignments/Code/einspeed/include/func.h

@@ -5,6 +5,7 @@
 #include <algorithm> 
 #include <cctype>
 #include <locale>
+#include <unordered_set>
 
 template <typename FROM, typename TO>
 TO* cast_all(size_t _n, FROM *_a) {
@@ -50,12 +51,13 @@ void print_vector(const std::vector<T>& vec, const std::string& label = "Vector"
 
 
 template <typename T>
-std::vector<size_t> multi_index(std::vector<size_t> indices, T *reference) {
-    std::vector<size_t> result(indices.size());
+std::vector<T> multi_index(std::vector<size_t> indices, T *reference) {
+    std::vector<T> result(indices.size());
     for (size_t i = 0; i < indices.size(); i++)
         result[i] = reference[indices[i]];
     return result;
 }
+
 
 // Source: https://stackoverflow.com/questions/216823/how-to-trim-a-stdstring#217605
 // trim from start (in place)
@@ -94,6 +96,47 @@ inline std::string rtrim_copy(std::string s) {
 inline std::string trim_copy(std::string s) {
     trim(s);
     return s;
+}
+
+
+template <typename T>
+std::vector<size_t> get_permutation(const std::vector<T>& source, const std::vector<T>& target) {
+    // Check that both vectors have the same size
+    if (source.size() != target.size()) {
+        throw std::invalid_argument("Source and target vectors must have the same size.");
+    }
+
+    // Check that both vectors have unique elements
+    auto has_duplicates = [](const std::vector<T>& vec) {
+        std::unordered_set<T> elements(vec.begin(), vec.end());
+        return elements.size() != vec.size();
+    };
+
+    if (has_duplicates(source)) {
+        throw std::invalid_argument("Source vector contains duplicate elements.");
+    }
+    if (has_duplicates(target)) {
+        throw std::invalid_argument("Target vector contains duplicate elements.");
+    }
+
+    // Check that both vectors have the same elements
+    auto sorted_source = source;
+    auto sorted_target = target;
+    std::sort(sorted_source.begin(), sorted_source.end());
+    std::sort(sorted_target.begin(), sorted_target.end());
+
+    if (sorted_source != sorted_target) {
+        throw std::invalid_argument("Source and target vectors must contain the same elements.");
+    }
+
+    // Create the permutation vector
+    std::vector<size_t> permutation(source.size());
+    for (size_t i = 0; i < source.size(); ++i) {
[…]

### 2024-12-14 Exam Assignments/Code/einspeed/include/parse_tensor.h

@@ -20,7 +20,7 @@ PyArrayObject* PyObject_to_PyArrayObject(PyObject * const _t) {
 template <typename T>
 const Tensor<T> parse_tensor(PyArrayObject * const _a) {
     size_t ndim = PyArray_NDIM(_a);
-    return Tensor<T>(PyArray_NDIM(_a), cast_all<npy_intp, size_t>(ndim, PyArray_DIMS(_a)), static_cast<T*>(PyArray_DATA(_a)));
+    return Tensor<T>(ndim, cast_all<npy_intp, size_t>(ndim, PyArray_DIMS(_a)), static_cast<T*>(PyArray_DATA(_a)));
 };
 
 #endif

### 2024-12-14 Exam Assignments/Code/einspeed/include/tensor.h

@@ -15,12 +15,15 @@
 template <typename T>
 class Tensor {
 public:
-    size_t ndim = 0;    // Number of dimensions
+    size_t ndim = 0;          // Number of dimensions
     size_t* shape = nullptr;  // Pointer to array defining the shape
     T* data = nullptr;        // Pointer to the actual data
 
     Tensor(size_t ndim, size_t* shape, T* data)
-        : ndim(ndim), shape(shape), data(data) {}
+        : ndim(ndim), data(data) {
+            this->shape = new size_t[ndim];
+            std::memcpy(this->shape, shape, ndim * sizeof(size_t));
+        }
     Tensor() {}
     
     Tensor(Tensor<T> &&other)

### 2024-12-14 Exam Assignments/Code/einspeed/src/einspeed.cpp

@@ -39,7 +39,7 @@ static PyObject* einsum(PyObject * const self, PyObject * const args) {
     for_<NO_SUPPORTED_NUMPY_TYPES>([&] (auto i) {
         if (SUPPORTED_NUMPY_TYPES[i.value] == a_type) {
             using cpp_type = typename numpy_to_cpp_type<SUPPORTED_NUMPY_TYPES[i.value]>::type;
-            result = compute_einsum<cpp_type>(s, parse_tensor<cpp_type>(a), parse_tensor<cpp_type>(b));
+            result = compute_einsum<cpp_type>(s, parse_tensor<cpp_type>(a), parse_tensor<cpp_type>(b)).to_PyObject();
         }
     });
 

### 2024-12-15 Exam Assignments/Code/einspeed/einspeed.py
Some improvements; heuristic bmm

@@ -1,17 +1,19 @@
 import einspeed
 import numpy as np
 
-# Pass and process a large matrix
-matrix = np.array([[1.0, 2.0, 3.0], [4.0, 5.0, 6.0], [7.0, 8.0, 9.0]], dtype=np.float64)
+dtype = np.complex128
+matrix = np.array([[1.0, 2.0, 3.0+4j], [4.0, 5.0, 6.0], [7.0, 8.0, 9.0]], dtype)
+copy = np.copy(matrix)
+vector = np.array([1, 2, 3], dtype)
+string = "ab,cd->dcba"
 
-# matrix = np.array([[1.0, 2.0, 3.0], [4.0, 5.0, 6.0], [7.0, 8.0, 9.0]], dtype=np.float32)
-result = einspeed.einsum("ai,jb->ij", matrix, matrix)
+result = einspeed.einsum(string, matrix, matrix)
+correct = np.einsum(string, matrix, matrix)
 
-print(np.einsum("ai,jb->ij", matrix, matrix))
+print("correct result:")
+print(correct)
 
 print("result:")
 print(result)
-# print(result.shape)
-# print("Updated matrix:")
-# print(matrix)
-# print(matrix.shape)
+
+print("Correct: ", (result==correct).all() and (matrix==copy).all())

### 2024-12-15 Exam Assignments/Code/einspeed/include/blas.h

@@ -17,10 +17,10 @@ Tensor<T> batch_matmul(const Tensor<T>& lhs, const Tensor<T>& rhs) {
     }
 
     // Shapes
-    size_t batch_size = lhs.shape[0];
-    size_t rows = lhs.shape[1];
-    size_t cols = rhs.shape[2];
-    size_t inner_dim = lhs.shape[2]; // Same as rhs.shape[1]
+    const size_t batch_size = lhs.shape[0];
+    const size_t rows = lhs.shape[1];
+    const size_t cols = rhs.shape[2];
+    const size_t inner_dim = lhs.shape[2]; // Same as rhs.shape[1]
 
     // Allocate the result tensor
     size_t result_shape[] = {batch_size, rows, cols};
@@ -36,6 +36,7 @@ Tensor<T> batch_matmul(const Tensor<T>& lhs, const Tensor<T>& rhs) {
         // Multiply the two matrices for the current batch
         for (size_t i = 0; i < rows; ++i) {
             for (size_t j = 0; j < cols; ++j) {
+                result_batch[i * cols + j] = 0;
                 for (size_t k = 0; k < inner_dim; ++k) {
                     result_batch[i * cols + j] += lhs_batch[i * inner_dim + k] * rhs_batch[k * cols + j];
                 }
@@ -43,13 +44,7 @@ Tensor<T> batch_matmul(const Tensor<T>& lhs, const Tensor<T>& rhs) {
         }
     }
 
-    std::cout << "HERE";
-    // Return the result tensor
-    auto t = Tensor<T>(3, result_shape, result_data);
-    std::cout << t.shape[0] << t.shape[1] << t.shape[2];
-    // t.print();
-    return t;
-    // return Tensor<T>(3, result_shape, result_data);
+    return Tensor<T>(3, result_shape, result_data);
 }
 
 #endif

### 2024-12-15 Exam Assignments/Code/einspeed/include/compute_einsum.h

@@ -1,45 +1,36 @@
 #ifndef COMPUTE_EINSUM_H
 #define COMPUTE_EINSUM_H
-#define DEBUG
 
-#include <string>
+// uncomment the following line to show debug information
+// #define DEBUG
+
+#include <unordered_map>
 #include <unordered_set>
 #include <stdexcept>
-// #include "tensor.h"
+#include <string>
+
+#include "tensor.h"
 #include "blas.h"
-
-// ******************************************
-// The core functionality is implemented HERE
-// ******************************************
-
-// template <typename T>
-// PyObject* _compute_einsum(const char * const s, const Tensor<T> &&lhs_tensor, const Tensor<T> &&rhs_tensor) {
-//     if (!is_valid_einsum_expression(s, lhs_tensor, rhs_tensor)) Py_RETURN_NONE;
-//     // auto target_current_column_identifiers = transform_matrices(s, lhs_tensor, rhs_tensor);
-//     return transform_matrices<T>(s, lhs_tensor, rhs_tensor).to_PyObject();
-
-//     if constexpr (std::is_same_v<T, double>) {
-//         // lhs_tensor.data[0] = 42;
-//     }
-//     auto perm = std::vector<size_t>({1, 0});
-//     auto shape = std::vector<size_t>({9});
-//     return lhs_tensor.transpose(perm).reshape(shape).to_PyObject();
-// }
+#include "func.h"
 
 
-std::vector<size_t> column_identifiers_like(const std::vector<size_t> &target_column_identifiers, const std::string &target_identifier_string, const std::string &origin_identifier_string) {
-    // it is assumed that every character in target_identifier_string accessed by target_column_identifiers is also present in origin_identifier_string
-    std::vector<size_t> origin_column_identifiers;
-    for (size_t i = 0; i < target_column_identifiers.size(); i++) {
-        origin_column_identifiers.push_back(origin_identifier_string.find(target_identifier_string[target_column_identifiers[i]]));
-    }
-    return origin_column_identifiers;
-}
+// **************************************************
+// *** The core functionality is implemented HERE ***
+// **************************************************
 
 
-// TODO: reread the entire comments to check for errors (since I changed all occurences of target to target_string, conflicting maybe with target_tensor)
+// Some declarations (implementation can be found below)
[…]

### 2024-12-15 Exam Assignments/Code/einspeed/include/func.h

@@ -1,21 +1,21 @@
 #ifndef FUNC_H
 #define FUNC_H
 
-#include <cstddef>
+#include <unordered_set>
 #include <algorithm> 
+#include <cstddef>
 #include <cctype>
 #include <locale>
-#include <unordered_set>
 
 template <typename FROM, typename TO>
-TO* cast_all(size_t _n, FROM *_a) {
+TO* cast_all(size_t _n, FROM* _a) {
     TO *_r = new TO[_n];
     for (size_t i = 0; i < _n; i++)
         _r[i] = static_cast<TO>(_a[i]);
     return _r;
 }
 
-inline bool contains(std::string &s, char c) {
+inline bool contains(const std::string &s, char c) {
     return s.find(c) != std::string::npos;
 }
 
@@ -51,7 +51,7 @@ void print_vector(const std::vector<T>& vec, const std::string& label = "Vector"
 
 
 template <typename T>
-std::vector<T> multi_index(std::vector<size_t> indices, T *reference) {
+std::vector<T> multi_index(const std::vector<size_t>& indices, T* reference) {
     std::vector<T> result(indices.size());
     for (size_t i = 0; i < indices.size(); i++)
         result[i] = reference[indices[i]];

### 2024-12-15 Exam Assignments/Code/einspeed/include/parse_tensor.h

@@ -6,21 +6,20 @@
 #include "tensor.h"
 #include "func.h"
 
-PyArrayObject* PyObject_to_PyArrayObject(PyObject * const _t) {
+const PyArrayObject* PyObject_to_PyArrayObject(const PyObject * const _t) {
     // Ensure the input is a NumPy array
     if (!PyArray_Check(_t)) {
-        PyErr_SetString(PyExc_TypeError, "Expected a NumPy array");
+        PyErr_SetString(PyExc_TypeError, "Expected a NumPy array.");
         return nullptr;
     }
-    // Cast to PyArrayObject and check type
-    return reinterpret_cast<PyArrayObject*>(_t);
+    return reinterpret_cast<const PyArrayObject*>(_t);
 }
 
 
 template <typename T>
-const Tensor<T> parse_tensor(PyArrayObject * const _a) {
+const Tensor<T> parse_tensor(const PyArrayObject * const _a) {
     size_t ndim = PyArray_NDIM(_a);
-    return Tensor<T>(ndim, cast_all<npy_intp, size_t>(ndim, PyArray_DIMS(_a)), static_cast<T*>(PyArray_DATA(_a)));
+    return Tensor<T>(ndim, cast_all<npy_intp, size_t>(ndim, PyArray_DIMS(const_cast<PyArrayObject*>(_a))), static_cast<T*>(PyArray_DATA(const_cast<PyArrayObject*>(_a))));
 };
 
 #endif

### 2024-12-15 Exam Assignments/Code/einspeed/include/tensor.h

@@ -3,14 +3,13 @@
 
 #include <Python.h>
 #include <numpy/arrayobject.h>
-// #include <cstddef>
-#include <vector>
-#include <algorithm>
-#include "numpy_types.h"
-// #include <stdexcept>
-#include <cstring> // For memcpy
 
+#include <algorithm>
 #include <iostream>
+#include <cstring>
+#include <vector>
+
+#include "numpy_types.h"
 
 template <typename T>
 class Tensor {
@@ -18,6 +17,7 @@ public:
     size_t ndim = 0;          // Number of dimensions
     size_t* shape = nullptr;  // Pointer to array defining the shape
     T* data = nullptr;        // Pointer to the actual data
+                              // IMPORTANT: Tensor class does NOT free the data!
 
     Tensor(size_t ndim, size_t* shape, T* data)
         : ndim(ndim), data(data) {
@@ -55,8 +55,8 @@ public:
 
 
     ~Tensor() {
-        if (this->shape) delete[] this->shape; // It is okay to delete this, since when casting to PyObject, we will create a copy (which we will later also delete though)
-        // if (this->data) delete[] this->data; // Don't delete data as it belongs to Python
+        if (this->shape) delete[] this->shape; // It is okay to delete this, since when casting to PyObject, we will create a copy (which we will later also delete though, Python seems to make a copy)
+        // Don't delete data as it might belong to Python
     }
 
     PyObject* to_PyObject() const {
@@ -65,14 +65,20 @@ public:
             import_array(); // Necessary to initialize NumPy C API
         }
 
+        int numpy_type = numpy_type_of<T>();
+
+        // If ndim == 0, return a scalar
+        if (ndim == 0) {
+            if (data) return PyArray_Scalar(data, PyArray_DescrFromType(numpy_type), nullptr);
+            else Py_RETURN_NONE;
+        }
+
         // Create a NumPy array
         npy_intp* np_shape = new npy_intp[ndim];
         for (size_t i = 0; i < ndim; ++i) {
             np_shape[i] = static_cast<npy_intp>(shape[i]);
         }
 
-        int numpy_type = numpy_type_of<T>();
-        
         PyObject* numpy_array = PyArray_SimpleNewFromData(
             static_cast<int>(ndim), np_shape, numpy_type, data
         );
@@ -80,7 +86,7 @@ public:
         // Set ownership of the data to NumPy
         PyArray_ENABLEFLAGS(reinterpret_cast<PyArrayObject*>(numpy_array), NPY_ARRAY_OWNDATA);
 
[…]

### 2024-12-15 Exam Assignments/Code/einspeed/src/einspeed.cpp

@@ -1,33 +1,27 @@
 #define NPY_NO_DEPRECATED_API NPY_1_7_API_VERSION
 #include <Python.h>
 #include <numpy/arrayobject.h>
-#include <complex>
 
-#include "parse_tensor.h"
 #include "compute_einsum.h"
+#include "parse_tensor.h"
 #include "numpy_types.h"
 #include "unroll_loop.h"
 
-#include <omp.h>
-#include <iostream>
-
-
-static PyObject* einsum(PyObject * const self, PyObject * const args) {
+static PyObject* einsum(PyObject* const self, PyObject* const args) {
     PyObject *lhs_o, *rhs_o;
     const char *s;
 
     // Parse the input argument as a generic PyObject
     if (!PyArg_ParseTuple(args, "sOO", &s, &lhs_o, &rhs_o)) {
-        return NULL;
+        PyErr_SetString(PyExc_TypeError, "Invalid arguments are passed to einsum.");
+        Py_RETURN_NONE;   
     }
 
-    PyArrayObject *a, *b;
-    a = PyObject_to_PyArrayObject(lhs_o);
-    b = PyObject_to_PyArrayObject(rhs_o);
+    const PyArrayObject* const a = PyObject_to_PyArrayObject(lhs_o);
+    const PyArrayObject* const b = PyObject_to_PyArrayObject(rhs_o);
 
-    int a_type, b_type;
-    a_type = PyArray_TYPE(a);
-    b_type = PyArray_TYPE(b);
+    const int a_type = PyArray_TYPE(a);
+    const int b_type = PyArray_TYPE(b);
 
     if (a_type != b_type) {
         PyErr_SetString(PyExc_TypeError, "The numpy arrays are expected to have the same data type.");

### 2024-12-15 Exam Assignments/Code/einspeed/einspeed.py
measuring time

@@ -1,19 +1,21 @@
 import einspeed
 import numpy as np
+import time
 
 dtype = np.complex128
-matrix = np.array([[1.0, 2.0, 3.0+4j], [4.0, 5.0, 6.0], [7.0, 8.0, 9.0]], dtype)
-copy = np.copy(matrix)
+matrix_lhs = np.random.rand(2, 2, 2).astype(dtype)
+# matrix_rhs = np.random.rand(100, 100, 100).astype(dtype)
+matrix_rhs = np.array([1], dtype)
+copy = np.copy(matrix_lhs)
 vector = np.array([1, 2, 3], dtype)
-string = "ab,cd->dcba"
+string = "bik,a->kib"
 
-result = einspeed.einsum(string, matrix, matrix)
-correct = np.einsum(string, matrix, matrix)
+start_time = time.time()
+correct = np.einsum(string, matrix_lhs, matrix_rhs)
+print(f"correct result in {time.time() - start_time} seconds")
 
-print("correct result:")
-print(correct)
+start_time = time.time()
+result = einspeed.einsum(string, matrix_lhs, matrix_rhs)
+print(f"result in {time.time() - start_time} seconds")
 
-print("result:")
-print(result)
-
-print("Correct: ", (result==correct).all() and (matrix==copy).all())
+print("Correct: ", (result==correct).all() and (matrix_lhs==copy).all())

### 2024-12-15 Exam Assignments/Code/einspeed/include/compute_einsum.h

@@ -149,18 +149,27 @@ Tensor<T> compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
 
     // Finally, perform the batch matrix multiplication.
     Tensor<T> bmm_result;
-    if (!(reduced_rhs.size() == 1 && reduced_rhs.data[0] == T(1))) // little heuristic to improve performance
-        bmm_result = std::move(batch_matmul(reduced_lhs, reduced_rhs));
-    else {
+    {
+    if (reduced_rhs.size() == 1 && reduced_rhs.data[0] == T(1)) { // little heuristic to improve performance
         bmm_result = std::move(reduced_lhs);
+        delete[] reduced_rhs.data;
+#ifdef DEBUG
+        std::cout << "skipped batch_matmul!\n";
+#endif
+    } else if (reduced_lhs.size() == 1 && reduced_lhs.data[0] == T(1)) {
+        bmm_result = std::move(reduced_rhs);
+        delete[] reduced_rhs.data;
 #ifdef DEBUG
         std::cout << "skipped batch_matmul!\n";
 #endif
     }
-
-    // Now we free the data of reduced_lhs and reduced_rhs, as it is no longer needed
-    // delete[] reduced_lhs.data;
-    // delete[] reduced_rhs.data;
+    else {
+        bmm_result = std::move(batch_matmul(reduced_lhs, reduced_rhs));
+        // Now we free the data of reduced_lhs and reduced_rhs, as it is no longer needed
+        delete[] reduced_lhs.data;
+        delete[] reduced_rhs.data;
+    }
+    }
 
 #ifdef DEBUG
     std::cout << "Result of Batch Matrix Multiplication:\n";
@@ -180,7 +189,12 @@ Tensor<T> compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
     target_shape.reserve(target_string.size());
     for (auto& i : batch_dims_lhs) target_shape.push_back(lhs_tensor.shape[i]);
     for (auto& i : kept_left_dims_lhs) target_shape.push_back(lhs_tensor.shape[i]);
-    for (auto& i : kept_right_dims_rhs) target_shape.push_back(lhs_tensor.shape[i]);
+    for (auto& i : kept_right_dims_rhs) target_shape.push_back(rhs_tensor.shape[i]);
+
+#ifdef DEBUG
+    std::cout << "New shape of target matrix:\n";
+    print_vector(target_shape);
+#endif
 
     bmm_result.reshape(target_shape);
 

### 2024-12-15 Exam Assignments/Code/einspeed/include/tensor.h

@@ -201,6 +201,10 @@ public:
 
     // Utility Methods
 
+    bool inline is_scalar() const {
+        return ndim == 0 && data;
+    }
+    
     static size_t calculate_size(const size_t* shape, size_t ndim) {
         size_t size = 1;
         for (size_t i = 0; i < ndim; ++i) {

### 2024-12-16 Exam Assignments/Code/einspeed/build/external/hptt/CMakeFiles/hptt.dir/compiler_depend.ts
hptt library

@@ -1,0 +1,2 @@
+# CMAKE generated file: DO NOT EDIT!
+# Timestamp file for compiler generated dependencies management for hptt.

### 2024-12-16 Exam Assignments/Code/einspeed/einspeed.py

@@ -2,20 +2,23 @@ import einspeed
 import numpy as np
 import time
 
-dtype = np.complex128
-matrix_lhs = np.random.rand(2, 2, 2).astype(dtype)
-# matrix_rhs = np.random.rand(100, 100, 100).astype(dtype)
-matrix_rhs = np.array([1], dtype)
+dtype = np.float64
+matrix_lhs = np.random.rand(100, 100, 100).astype(dtype)
+matrix_rhs = np.random.rand(100, 100, 100).astype(dtype)
+# matrix_lhs = np.array([[[1,2],[3,4]],[[5,6], [7,8]]], dtype)
+# matrix_rhs = matrix_lhs
+# matrix_rhs = np.array([1], dtype)
 copy = np.copy(matrix_lhs)
 vector = np.array([1, 2, 3], dtype)
-string = "bik,a->kib"
+string = "bik,bkj->bij"
 
 start_time = time.time()
 correct = np.einsum(string, matrix_lhs, matrix_rhs)
 print(f"correct result in {time.time() - start_time} seconds")
+print(correct.strides)
 
 start_time = time.time()
 result = einspeed.einsum(string, matrix_lhs, matrix_rhs)
 print(f"result in {time.time() - start_time} seconds")
-
+print(result.strides)
 print("Correct: ", (result==correct).all() and (matrix_lhs==copy).all())

### 2024-12-16 Exam Assignments/Code/einspeed/include/blas.h

@@ -47,4 +47,142 @@ Tensor<T> batch_matmul(const Tensor<T>& lhs, const Tensor<T>& rhs) {
     return Tensor<T>(3, result_shape, result_data);
 }
 
+
+template <typename T>
+Tensor<T> _batch_matrix_times_transpose_matrix(const Tensor<T>& lhs, const Tensor<T>& rhs) {
+    // Check dimensions
+    if (lhs.ndim != 3 || rhs.ndim != 3) {
+        throw std::invalid_argument("Both tensors must have 3 dimensions for batch matrix multiplication.");
+    }
+    if (lhs.shape[0] != rhs.shape[0]) {
+        throw std::invalid_argument("Batch sizes must match for batch matrix multiplication.");
+    }
+    if (lhs.shape[2] != rhs.shape[2]) {
+        throw std::invalid_argument("Inner matrix dimensions must match for batch multiplication with transpose.");
+    }
+
+    // Shapes
+    const size_t batch_size = lhs.shape[0];
+    const size_t rows = lhs.shape[1];
+    const size_t cols = rhs.shape[1]; // Since rhs is transposed
+    const size_t inner_dim = lhs.shape[2];
+
+    // Allocate the result tensor
+    size_t result_shape[] = {batch_size, rows, cols};
+    T* result_data = new T[batch_size * rows * cols];
+
+    // Perform batch matrix multiplication with transpose
+    #pragma omp parallel for
+    for (size_t b = 0; b < batch_size; ++b) {
+        // Get pointers to the current batch matrices
+        const T* lhs_batch = lhs.data + b * rows * inner_dim;
+        const T* rhs_batch = rhs.data + b * cols * inner_dim; // Use rhs.shape[1] because it's transposed
+        T* result_batch = result_data + b * rows * cols;
+
+        // Multiply lhs with transposed rhs for the current batch
+        for (size_t i = 0; i < rows; ++i) {
+            for (size_t j = 0; j < cols; ++j) {
+                result_batch[i * cols + j] = 0;
+                for (size_t k = 0; k < inner_dim; ++k) {
+                    result_batch[i * cols + j] += lhs_batch[i * inner_dim + k] * rhs_batch[j * inner_dim + k];
+                }
+            }
+        }
+    }
+
+    return Tensor<T>(3, result_shape, result_data);
+}
+
+
+template <typename T>
+Tensor<T> batch_matrix_times_transpose_matrix(const Tensor<T>& lhs, const Tensor<T>& rhs) {
+    // Check dimensions
[…]

### 2024-12-16 Exam Assignments/Code/einspeed/include/compute_einsum.h

@@ -96,8 +96,9 @@ Tensor<T> compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
     // lhs_tensor[batch, kept left, contracted, summed left]
     // rhs_tensor[batch, contracted, kept right, summed right]
     Tensor<T> transposed_lhs = lhs_tensor.transpose(merge_vectors({batch_dims_lhs, kept_left_dims_lhs, contracted_dims_lhs, summed_left_dims_lhs}));
-    Tensor<T> transposed_rhs = rhs_tensor.transpose(merge_vectors({batch_dims_rhs, contracted_dims_rhs, kept_right_dims_rhs, summed_right_dims_rhs}));
-
+    // Tensor<T> transposed_rhs = rhs_tensor.transpose(merge_vectors({batch_dims_rhs, contracted_dims_rhs, kept_right_dims_rhs, summed_right_dims_rhs}));
+    Tensor<T> transposed_rhs = rhs_tensor.transpose(merge_vectors({batch_dims_rhs, kept_right_dims_rhs, contracted_dims_rhs, summed_right_dims_rhs}));
+    print_vector(merge_vectors({batch_dims_rhs, kept_right_dims_rhs, contracted_dims_rhs, summed_right_dims_rhs}), "B permutation");
 #ifdef DEBUG
     std::cout << "Transposed Tensors:" << '\n';
     transposed_lhs.print();
@@ -113,10 +114,15 @@ Tensor<T> compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
                                                 Tensor<T>::calculate_size(multi_index(kept_left_dims_lhs, lhs_tensor.shape)),
                                                 Tensor<T>::calculate_size(multi_index(contracted_dims_lhs, lhs_tensor.shape)),
                                                 Tensor<T>::calculate_size(multi_index(summed_left_dims_lhs, lhs_tensor.shape))}));
+    // transposed_rhs.reshape(std::vector<size_t>({Tensor<T>::calculate_size(multi_index(batch_dims_rhs, rhs_tensor.shape)),
+    //                                             Tensor<T>::calculate_size(multi_index(contracted_dims_rhs, rhs_tensor.shape)),
+    //                                             Tensor<T>::calculate_size(multi_index(kept_right_dims_rhs, rhs_tensor.shape)),
+    //                                             Tensor<T>::calculate_size(multi_index(summed_right_dims_rhs, rhs_tensor.shape))}));
[…]

### 2024-12-16 Exam Assignments/Code/einspeed/include/func.h

@@ -8,7 +8,7 @@
 #include <locale>
 
 template <typename FROM, typename TO>
-TO* cast_all(size_t _n, FROM* _a) {
+TO* cast_all(size_t _n, const FROM* _a) {
     TO *_r = new TO[_n];
     for (size_t i = 0; i < _n; i++)
         _r[i] = static_cast<TO>(_a[i]);

### 2024-12-16 Exam Assignments/Code/einspeed/include/tensor.h

@@ -4,12 +4,16 @@
 #include <Python.h>
 #include <numpy/arrayobject.h>
 
+#include <omp.h>
+
 #include <algorithm>
 #include <iostream>
 #include <cstring>
 #include <vector>
 
 #include "numpy_types.h"
+#include "func.h"
+#include "hptt.h"
 
 template <typename T>
 class Tensor {
@@ -157,6 +161,9 @@ public:
         // Create and return the new transposed tensor
         return Tensor<T>(ndim, new_shape, new_data);
     }
+    Tensor<T> fast_transpose(const std::vector<size_t>& perm) const;
+
+    
 
 
     // Reshape method
@@ -270,5 +277,62 @@ private:
         return strides;
     }
 };
+
+
+
+
+template <typename T>
+Tensor<T> Tensor<T>::fast_transpose(const std::vector<size_t>& perm) const {
+#ifdef DEBUG
+    std::cout << "Using fast_transpose\n";
+#endif
+    if (perm.size() != ndim) {
+        throw std::invalid_argument("Permutation size must match the number of dimensions.");
+    }
+    if (ndim == 0) {
+        // needs special treatment
+        if (data) {
+            T* new_data =  new T[1];
+            *new_data = this->data[0];
+            return Tensor<T>(ndim, shape, new_data);
+        } else return Tensor<T>();
+    }
+
+    // Check if perm is a valid permutation
+    std::vector<bool> seen(ndim, false);
+    for (size_t i : perm) {
+        if (i >= ndim || seen[i]) {
+            throw std::invalid_argument("Invalid permutation array.");
+        }
+        seen[i] = true;
+    }
+
+    // Compute the new shape
+    size_t* new_shape = new size_t[ndim];
+    for (size_t i = 0; i < ndim; ++i) {
+        new_shape[i] = shape[perm[i]];
+    }
+
+
+    T* transposed = new T[Tensor<T>::calculate_size(shape, ndim)];
+    int* tmp_size = cast_all<size_t, int>(ndim, shape);
+    int* tmp_perm = cast_all<size_t, int>(ndim, perm.data());
+    // create a plan (shared_ptr)
+    auto plan = hptt::create_plan( tmp_perm, (int) ndim, 
+                                T(1), data, tmp_size, NULL, 
+                                T(0),  transposed, NULL, 
+                                hptt::ESTIMATE, omp_get_num_procs(), nullptr, true);
+    
[…]

### 2024-12-16 Exam Assignments/Code/einspeed/einspeed.py
working fast version

@@ -3,14 +3,16 @@ import numpy as np
 import time
 
 dtype = np.float64
-matrix_lhs = np.random.rand(100, 100, 100).astype(dtype)
-matrix_rhs = np.random.rand(100, 100, 100).astype(dtype)
+epsilon = 1e-2 # we need to be generous here, as there is a big computing error on large tensors
+size = 500
+matrix_lhs = np.random.rand(size, size, size).astype(dtype)
+matrix_rhs = np.random.rand(size, size, size).astype(dtype)
 # matrix_lhs = np.array([[[1,2],[3,4]],[[5,6], [7,8]]], dtype)
 # matrix_rhs = matrix_lhs
-# matrix_rhs = np.array([1], dtype)
+matrix_rhs = np.array([1], dtype)
 copy = np.copy(matrix_lhs)
 vector = np.array([1, 2, 3], dtype)
-string = "bik,bkj->bij"
+string = "bik,a->kib"
 
 start_time = time.time()
 correct = np.einsum(string, matrix_lhs, matrix_rhs)
@@ -20,5 +22,5 @@ print(correct.strides)
 start_time = time.time()
 result = einspeed.einsum(string, matrix_lhs, matrix_rhs)
 print(f"result in {time.time() - start_time} seconds")
+print("Correct: ", (np.abs((result-correct)) < epsilon).all() and (matrix_lhs==copy).all())
 print(result.strides)
-print("Correct: ", (result==correct).all() and (matrix_lhs==copy).all())

### 2024-12-16 Exam Assignments/Code/einspeed/include/compute_einsum.h

@@ -25,7 +25,7 @@ template <typename T> bool is_valid_einsum_expression(const char* const s, const
 
 
 template <typename T>
-Tensor<T> compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const Tensor<T>& rhs_tensor) {
+PyObject* compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const Tensor<T>& rhs_tensor) {
 
     // *************************
     // *** Function Overview ***
@@ -98,7 +98,6 @@ Tensor<T> compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
     Tensor<T> transposed_lhs = lhs_tensor.transpose(merge_vectors({batch_dims_lhs, kept_left_dims_lhs, contracted_dims_lhs, summed_left_dims_lhs}));
     // Tensor<T> transposed_rhs = rhs_tensor.transpose(merge_vectors({batch_dims_rhs, contracted_dims_rhs, kept_right_dims_rhs, summed_right_dims_rhs}));
     Tensor<T> transposed_rhs = rhs_tensor.transpose(merge_vectors({batch_dims_rhs, kept_right_dims_rhs, contracted_dims_rhs, summed_right_dims_rhs}));
-    print_vector(merge_vectors({batch_dims_rhs, kept_right_dims_rhs, contracted_dims_rhs, summed_right_dims_rhs}), "B permutation");
 #ifdef DEBUG
     std::cout << "Transposed Tensors:" << '\n';
     transposed_lhs.print();
@@ -170,7 +169,6 @@ Tensor<T> compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
 #endif
     }
     else {
-        // bmm_result = std::move(batch_matmul(reduced_lhs, reduced_rhs));
         bmm_result = std::move(batch_matrix_times_transpose_matrix(reduced_lhs, reduced_rhs));
         // Now we free the data of reduced_lhs and reduced_rhs, as it is no longer needed
         delete[] reduced_lhs.data;
@@ -219,7 +217,7 @@ Tensor<T> compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
         multi_index(kept_right_dims_rhs, rhs_string.data()),
     }), std::vector<char>(target_string.begin(), target_string.end()));
 
-    return bmm_result.transpose(target_permutation);
+    return bmm_result.lazy_transpose_and_return_PyObjet(target_permutation);
 }
 
 

### 2024-12-16 Exam Assignments/Code/einspeed/include/tensor.h

@@ -5,6 +5,7 @@
 #include <numpy/arrayobject.h>
 
 #include <omp.h>
+#include <immintrin.h>  // For SIMD intrinsics (optional)
 
 #include <algorithm>
 #include <iostream>
@@ -104,6 +105,7 @@ public:
 
     // Transpose method
     Tensor<T> transpose(const std::vector<size_t>& perm) const {
+        // return fast_transpose(perm);
         if (perm.size() != ndim) {
             throw std::invalid_argument("Permutation size must match the number of dimensions.");
         }
@@ -162,6 +164,58 @@ public:
         return Tensor<T>(ndim, new_shape, new_data);
     }
     Tensor<T> fast_transpose(const std::vector<size_t>& perm) const;
+    PyObject* lazy_transpose_and_return_PyObjet(const std::vector<size_t> &perm) const {
+        // Initialize NumPy (if not already done)
+        if (!PyArray_API) {
+            import_array(); // Necessary to initialize NumPy C API
+        }
+
+        int numpy_type = numpy_type_of<T>();
+
+        // If ndim == 0, return a scalar
+        if (ndim == 0) {
+            if (data) return PyArray_Scalar(data, PyArray_DescrFromType(numpy_type), nullptr);
+            else Py_RETURN_NONE;
+        }
+
+        // Create the shape and strides arrays
+        npy_intp* np_shape = new npy_intp[ndim];
+        npy_intp* np_strides = new npy_intp[ndim];
+
+        std::vector<size_t> strides = compute_strides(shape, ndim);
+        for (size_t i = 0; i < ndim; ++i) {
+            np_shape[i] = static_cast<npy_intp>(shape[perm[i]]);
+            np_strides[i] = static_cast<npy_intp>(strides[perm[i]] * sizeof(T)); // Convert strides to byte offsets
+        }
+
+        // Create a NumPy array
+        PyObject* numpy_array = PyArray_New(
+            &PyArray_Type,                         // Type of array
+            static_cast<int>(ndim),                // Number of dimensions
+            np_shape,                              // Shape
+            numpy_type,                            // NumPy data type
+            np_strides,                            // Strides
+            data,                                  // Data pointer
+            0,                                     // Item size (0 = use the default for the dtype)
+            NPY_ARRAY_CARRAY,                      // Flags
[…]

### 2024-12-16 Exam Assignments/Code/einspeed/src/einspeed.cpp

@@ -33,7 +33,7 @@ static PyObject* einsum(PyObject* const self, PyObject* const args) {
     for_<NO_SUPPORTED_NUMPY_TYPES>([&] (auto i) {
         if (SUPPORTED_NUMPY_TYPES[i.value] == a_type) {
             using cpp_type = typename numpy_to_cpp_type<SUPPORTED_NUMPY_TYPES[i.value]>::type;
-            result = compute_einsum<cpp_type>(s, parse_tensor<cpp_type>(a), parse_tensor<cpp_type>(b)).to_PyObject();
+            result = compute_einsum<cpp_type>(s, parse_tensor<cpp_type>(a), parse_tensor<cpp_type>(b));
         }
     });
 

### 2024-12-16 Exam Assignments/Code/einspeed/einspeed.py
added reusing available memory

@@ -2,17 +2,18 @@ import einspeed
 import numpy as np
 import time
 
-dtype = np.float64
+dtype = np.complex128
 epsilon = 1e-2 # we need to be generous here, as there is a big computing error on large tensors
-size = 500
+size = 50
 matrix_lhs = np.random.rand(size, size, size).astype(dtype)
 matrix_rhs = np.random.rand(size, size, size).astype(dtype)
 # matrix_lhs = np.array([[[1,2],[3,4]],[[5,6], [7,8]]], dtype)
 # matrix_rhs = matrix_lhs
-matrix_rhs = np.array([1], dtype)
+# matrix_rhs = np.array([1], dtype)
 copy = np.copy(matrix_lhs)
 vector = np.array([1, 2, 3], dtype)
-string = "bik,a->kib"
+# string = "abc,def->" # einsum implementation is HORRENDOUS at reducing
+string = "abc,def->"
 
 start_time = time.time()
 correct = np.einsum(string, matrix_lhs, matrix_rhs)
@@ -24,3 +25,4 @@ result = einspeed.einsum(string, matrix_lhs, matrix_rhs)
 print(f"result in {time.time() - start_time} seconds")
 print("Correct: ", (np.abs((result-correct)) < epsilon).all() and (matrix_lhs==copy).all())
 print(result.strides)
+print(correct, result)

### 2024-12-16 Exam Assignments/Code/einspeed/include/blas.h

@@ -3,99 +3,9 @@
 
 #include "tensor.h"
 
-template <typename T>
-Tensor<T> batch_matmul(const Tensor<T>& lhs, const Tensor<T>& rhs) {
-    // Check dimensions
-    if (lhs.ndim != 3 || rhs.ndim != 3) {
-        throw std::invalid_argument("Both tensors must have 3 dimensions for batch matrix multiplication.");
-    }
-    if (lhs.shape[0] != rhs.shape[0]) {
-        throw std::invalid_argument("Batch sizes must match for batch matrix multiplication.");
-    }
-    if (lhs.shape[2] != rhs.shape[1]) {
-        throw std::invalid_argument("Inner matrix dimensions must match for batch matrix multiplication.");
-    }
-
-    // Shapes
-    const size_t batch_size = lhs.shape[0];
-    const size_t rows = lhs.shape[1];
-    const size_t cols = rhs.shape[2];
-    const size_t inner_dim = lhs.shape[2]; // Same as rhs.shape[1]
-
-    // Allocate the result tensor
-    size_t result_shape[] = {batch_size, rows, cols};
-    T* result_data = new T[batch_size * rows * cols];
-
-    // Perform batch matrix multiplication
-    for (size_t b = 0; b < batch_size; ++b) {
-        // Get pointers to the current batch matrices
-        const T* lhs_batch = lhs.data + b * rows * inner_dim;
-        const T* rhs_batch = rhs.data + b * inner_dim * cols;
-        T* result_batch = result_data + b * rows * cols;
-
-        // Multiply the two matrices for the current batch
-        for (size_t i = 0; i < rows; ++i) {
-            for (size_t j = 0; j < cols; ++j) {
-                result_batch[i * cols + j] = 0;
-                for (size_t k = 0; k < inner_dim; ++k) {
-                    result_batch[i * cols + j] += lhs_batch[i * inner_dim + k] * rhs_batch[k * cols + j];
-                }
-            }
-        }
-    }
-
-    return Tensor<T>(3, result_shape, result_data);
-}
-
 
 template <typename T>
-Tensor<T> _batch_matrix_times_transpose_matrix(const Tensor<T>& lhs, const Tensor<T>& rhs) {
-    // Check dimensions
-    if (lhs.ndim != 3 || rhs.ndim != 3) {
-        throw std::invalid_argument("Both tensors must have 3 dimensions for batch matrix multiplication.");
-    }
-    if (lhs.shape[0] != rhs.shape[0]) {
[…]

### 2024-12-16 Exam Assignments/Code/einspeed/include/compute_einsum.h

@@ -43,7 +43,7 @@ PyObject* compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
     // We do the same for rhs_string. Then, we transpose and reshape lhs_tensor and rhs_tensor in the following manner:
     //
     // lhs_tensor[batch, kept left, contracted, summed left]
-    // rhs_tensor[batch, contracted, kept right, summed right]
+    // rhs_tensor[batch, kept right, contracted, summed right]
     //
     // IMPORTANT: the column identifiers of batch and contracted of both tensors are obviously the same.
     //            We have to ensure that they also have the same dimensions and ordering!
@@ -94,10 +94,10 @@ PyObject* compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
 
     // Reshaping the lhs and rhs tensors as described:
     // lhs_tensor[batch, kept left, contracted, summed left]
-    // rhs_tensor[batch, contracted, kept right, summed right]
+    // rhs_tensor[batch, kept right, contracted, summed right]
     Tensor<T> transposed_lhs = lhs_tensor.transpose(merge_vectors({batch_dims_lhs, kept_left_dims_lhs, contracted_dims_lhs, summed_left_dims_lhs}));
-    // Tensor<T> transposed_rhs = rhs_tensor.transpose(merge_vectors({batch_dims_rhs, contracted_dims_rhs, kept_right_dims_rhs, summed_right_dims_rhs}));
     Tensor<T> transposed_rhs = rhs_tensor.transpose(merge_vectors({batch_dims_rhs, kept_right_dims_rhs, contracted_dims_rhs, summed_right_dims_rhs}));
+
 #ifdef DEBUG
     std::cout << "Transposed Tensors:" << '\n';
     transposed_lhs.print();
@@ -113,15 +113,10 @@ PyObject* compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
                                                 Tensor<T>::calculate_size(multi_index(kept_left_dims_lhs, lhs_tensor.shape)),
                                                 Tensor<T>::calculate_size(multi_index(contracted_dims_lhs, lhs_tensor.shape)),
                                                 Tensor<T>::calculate_size(multi_index(summed_left_dims_lhs, lhs_tensor.shape))}));
-    // transposed_rhs.reshape(std::vector<size_t>({Tensor<T>::calculate_size(multi_index(batch_dims_rhs, rhs_tensor.shape)),
[…]

### 2024-12-16 Exam Assignments/Code/einspeed/include/numpy_types.h

@@ -50,7 +50,7 @@ const int NUMPY_TYPES[] = {
 };
 
 constexpr const int SUPPORTED_NUMPY_TYPES[] = {
-    NPY_INT64,
+    // NPY_INT64, // no int64, as hptt does not support it
     NPY_FLOAT64,
     NPY_COMPLEX128
 };

### 2024-12-16 Exam Assignments/Code/einspeed/include/tensor.h

@@ -5,7 +5,7 @@
 #include <numpy/arrayobject.h>
 
 #include <omp.h>
-#include <immintrin.h>  // For SIMD intrinsics (optional)
+#include <immintrin.h>
 
 #include <algorithm>
 #include <iostream>
@@ -104,67 +104,8 @@ public:
     // **********
 
     // Transpose method
-    Tensor<T> transpose(const std::vector<size_t>& perm) const {
-        // return fast_transpose(perm);
-        if (perm.size() != ndim) {
-            throw std::invalid_argument("Permutation size must match the number of dimensions.");
-        }
-        if (ndim == 0) {
-            // needs special treatment
-            if (data) {
-                T* new_data =  new T[1];
-                *new_data = this->data[0];
-                return Tensor<T>(ndim, shape, new_data);
-            } else return Tensor<T>();
-        }
-
-        // Check if perm is a valid permutation
-        std::vector<bool> seen(ndim, false);
-        for (size_t i : perm) {
-            if (i >= ndim || seen[i]) {
-                throw std::invalid_argument("Invalid permutation array.");
-            }
-            seen[i] = true;
-        }
-
-        // Compute the new shape
-        size_t* new_shape = new size_t[ndim];
-        for (size_t i = 0; i < ndim; ++i) {
-            new_shape[i] = shape[perm[i]];
-        }
-
-        // Allocate memory for the transposed data
-        T* new_data = new T[calculate_size(new_shape, ndim)];
-
-        // Compute strides for both old and new tensor layouts
-        std::vector<size_t> old_strides = compute_strides(shape, ndim);
-        std::vector<size_t> new_strides = compute_strides(new_shape, ndim);
-
-        // Fill the new_data array
-        std::vector<size_t> old_indices(ndim, 0);
-        for (size_t new_index = 0; new_index < calculate_size(new_shape, ndim); ++new_index) {
-            // Map the flat new_index to multi-dimensional indices in the new layout
-            size_t remaining = new_index;
-            for (size_t i = 0; i < ndim; ++i) {
-                old_indices[perm[i]] = remaining / new_strides[i];
-                remaining %= new_strides[i];
-            }
-
-            // Compute the flat index in the original tensor
-            size_t old_index = 0;
[…]

### 2024-12-17 Exam Assignments/paper/paper.tex
including unit tests

\documentclass[sigconf]{acmart}
\usepackage{booktabs}

%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% These commands are for a PROCEEDINGS abstract or paper.
\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in

\acmConference[AEPRO 2025]{AEPRO 2025: Algorithm Engineering Projects}{March 1}{Jena, Germany}

% convert text to title case
% http://individed.com/code/to-title-case/

% that helps you to formulate your sentences
% https://www.deepl.com/translator

\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title[Fast Entity Resolution With Mock Labels and Sorted Integer Sets]{Fast Entity Resolution With Mock Labels and Sorted Integer Sets\\\large Algorithm Engineering 2025 Project Paper}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.

\author{Max Mustermann}
\affiliation{%
  \institution{Friedrich Schiller University Jena}
  \country{Germany}}
\email{max.mustermann@uni-jena.de}

\author{Erika Mustermann}
\affiliation{%
  \institution{Friedrich Schiller University Jena}
  \country{Germany}}
\email{erika.mustermann@uni-jena.de}

%% The abstract is a short summary of the work to be presented in the article.
\begin{abstract}

The five-finger pattern \cite{macgilchrist2014}:
\begin{enumerate}
\item \textbf{Topic and background:} What topic does the paper deal with? What is the point of departure for your research? Why are you studying this now?
\item \textbf{Focus:} What is your research question? What are you studying precisely?
\item \textbf{Method:} What did you do?
\item \textbf{Key findings:} What did you discover?
\item \textbf{Conclusions or implications:} What do these findings mean? What broader issues do they speak to?
\end{enumerate}

\end{abstract}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{entity resolution, data cleansing, programming contest}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\let\thefootnote\relax\footnotetext{AEPRO 2025, March 1, Jena, Germany. Copyright \copyright 2025 for this paper by its authors. Use permitted under Creative Commons License Attribution 4.0 International (CC BY 4.0).}

\section{Introduction}

\subsection{Background}

\subsection{Related Work}

\subsection{Our Contributions}

\subsection{Outline}

\section{The Algorithm}

\subsection{Internal Representation of Mock Labels}
\label{sub:sec:internal}

In Figure~\ref{fig:integer:sets} we convert the mock labels to sorted integer sets.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\linewidth]{./graphics/integer_sets.pdf}
  \caption{Conversion of mock camera labels to sorted integer sets. 
We map each unique token (key) in camera labels to a unique value. 
Based on these key-value-mappings, we convert camera labels to sorted integer sets. 
A camera can have different names in different countries. Therefore, repeating IDs reference the same cameras (see, for example, ID=3).} 
  \label{fig:integer:sets}
\end{figure}

\subsection{Efficient Preprocessing of Input Data}
\label{sub:sec:preprocessing}

The following findings are important to speed up preprocessing of the input data:

\begin{itemize}
\item Reading many small files concurrently, with multiple threads (compared to a single thread), takes advantage of the internal parallelism of SSDs and thus leads to higher throughput \cite{Zhuang2016}.

\item C-string manipulation functions are often significantly faster than their C++ pendants. For example, locating substrings with \texttt{strstr} is around five times faster than using the C++ \texttt{std::string} function \texttt{find}.

\item Hardcoding regular expressions with \emph{while, for, switch} or \emph{if-else} statements results in faster execution times than using standard RegEx libraries, where regular expressions are compiled at runtime into state machines.

\item Changing strings in place, instead of treating them as immutable objects, eliminates allocation and copying overhead.

\end{itemize}

\section{Experiments}

Table~\ref{tab:results} shows the running times of the resolution step of the five best placed teams.

\begin{table}[htbp]
  \caption{Comparison of the F-measure and the running times of the resolution step of the five best placed teams. The input data for the resolution step consisted of 29{,}787 in JSON formatted e-commerce websites. Measurements were taken on a
laptop running Ubuntu 19.04 with 16 GB of RAM and two Intel Core i5-4310U CPUs. The underlying SSD was a 500\,GB 860 EVO mSATA. We cleared the page cache, dentries, and inodes before each run to avoid reading the input data from RAM instead of the SSD.}
  \label{tab:results}
\resizebox{\columnwidth}{!}{
  \begin{tabular}{lcrr}
    \toprule
    Team& Language & F-measure & Running time (s)\\
    \midrule
	PictureMe (\textbf{this paper}) &C++& 0.99 & \textbf{0.61}\\
    DBGroup@UniMoRe &Python& 0.99 & 10.65\\
    DBGroup@SUSTech &C++& 0.99 & 22.13\\
    eats\_shoots\_and\_leaves &Python& 0.99 & 28.66\\
    DBTHU &Python& 0.99& 63.21\\
  \bottomrule
\end{tabular}
}
\end{table}

\section{Conclusions}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{literature}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.

### 2024-12-17 Exam Assignments/Code/einspeed/einspeed.py

@@ -5,8 +5,8 @@ import time
 dtype = np.complex128
 epsilon = 1e-2 # we need to be generous here, as there is a big computing error on large tensors
 size = 50
-matrix_lhs = np.random.rand(size, size, size).astype(dtype)
-matrix_rhs = np.random.rand(size, size, size).astype(dtype)
+matrix_lhs = (1+np.random.rand(size, size, size)).astype(dtype)
+matrix_rhs = (1+np.random.rand(size, size, size)).astype(dtype)
 # matrix_lhs = np.array([[[1,2],[3,4]],[[5,6], [7,8]]], dtype)
 # matrix_rhs = matrix_lhs
 # matrix_rhs = np.array([1], dtype)

### 2024-12-17 Exam Assignments/Code/einspeed/tests/catch_tests_fizzbuzz.cpp

@@ -1,0 +1,55 @@
+#include "catch.hpp"
+#include "tensor.h"
+
+TEST_CASE("moin", "[classic]") {
+   SECTION("LOL") {
+      REQUIRE(true);
+   }
+}
+// TEST_CASE("Test positives", "[classic]")
+// {
+//    SECTION("Test all up to 10") {
+//       REQUIRE(fizzbuzz(1) == "1"); 
+//       REQUIRE(fizzbuzz(2) == "2");
+//       REQUIRE(fizzbuzz(3) == "fizz");
+//       REQUIRE(fizzbuzz(4) == "4");
+//       REQUIRE(fizzbuzz(5) == "buzz");
+//       REQUIRE(fizzbuzz(6) == "fizz");
+//       REQUIRE(fizzbuzz(7) == "7");
+//       REQUIRE(fizzbuzz(8) == "8");
+//       REQUIRE(fizzbuzz(9) == "fizz");
+//       REQUIRE(fizzbuzz(10) == "buzz");
+//    }
+
+//    SECTION("Test all multiples of 3 only up to 100") {
+//       for (int i = 3; i <= 100; i+=3) {
+//          if (i % 5) REQUIRE(fizzbuzz(i) == "fizz");
+//       }
+//    }
+
+//    SECTION("Test all multiples of 5 only up to 100") {
+//       for (int i = 5; i <= 100; i += 5) {
+//          if (i % 3) REQUIRE(fizzbuzz(i) == "buzz");
+//       }
+//    }
+
+//    SECTION("Test all multiples of 3 and 5 up to 100") {
+//       for (int i = 15; i <= 100; i += 15) {
+//          REQUIRE(fizzbuzz(i) == "fizzbuzz");
+//       }
+//    }
+// }
+
+// TEST_CASE("Test negatives", "[classic]")
+// {
+//    REQUIRE(fizzbuzz(-1) == "-1");
+//    REQUIRE(fizzbuzz(-2) == "-2");
+//    REQUIRE(fizzbuzz(-3) == "fizz");
+//    REQUIRE(fizzbuzz(-4) == "-4");
+//    REQUIRE(fizzbuzz(-5) == "buzz");
+//    REQUIRE(fizzbuzz(-6) == "fizz");
+//    REQUIRE(fizzbuzz(-7) == "-7");
+//    REQUIRE(fizzbuzz(-8) == "-8");
[…]

### 2024-12-17 Exam Assignments/paper/paper.tex
paper introduction

@@
 
-\acmConference[AEPRO 2025]{AEPRO 2025: Algorithm Engineering Projects}{March 1}{Jena, Germany}
+% \acmConference[AEPRO 2025]{AEPRO 2025: Algorithm Engineering Projects}{March 1}{Jena, Germany}
 
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title[Fast Entity Resolution With Mock Labels and Sorted Integer Sets]{Fast Entity Resolution With Mock Labels and Sorted Integer Sets\\\large Algorithm Engineering 2025 Project Paper}
+\title[Fast Einsum Implementation]{Fast Einsum Implementation Using Parallelization, Vectorization, and More\\\large Algorithm Engineering 2025 Project Paper}
 
@@
 
-\author{Max Mustermann}
+\author{Jonas Peters}
 \affiliation{%
@@
   \country{Germany}}
-\email{max.mustermann@uni-jena.de}
+\email{jonas.peters@uni-jena.de}
 
-\author{Erika Mustermann}
-\affiliation{%
-  \institution{Friedrich Schiller University Jena}
-  \country{Germany}}
-\email{erika.mustermann@uni-jena.de}
+% \author{Erika Mustermann}
+% \affiliation{%
+%   \institution{Friedrich Schiller University Jena}
+%   \country{Germany}}
+% \email{erika.mustermann@uni-jena.de}
 
@@
 
-The five-finger pattern \cite{macgilchrist2014}:
+% The five-finger pattern \cite{macgilchrist2014}:
 \begin{enumerate}
-\item \textbf{Topic and background:} What topic does the paper deal with? What is the point of departure for your research? Why are you studying this now?
+% \item \textbf{Topic and background:} What topic does the paper deal with? What is the point of departure for your research? Why are you studying this now?
+\item \textbf{Topic and background:} The topic of this paper is tensor transformations, and how to perform them efficiently.
+Tensors consist of many dimensions.
+Take matrices as specialized tensors with two dimensions (rows and columns) for example.
+When multiplying two matrices, we calculate "row times column" for each position in the resulting matrix.
+In tensor language, this is equivalent to \em contracting \em the the second dimension of the first matrix with the first dimension of the second matrix.
+We can also just rearrange the dimensions in a new order, which is called  \em transposing\em , similar to matrices.
+Lastly, we can \em reduce \em entire dimensions by summing over all entries, for example reducing the second dimension of a matrix $M$ is equivalent to
+\[ M_{reduced} = M \cdot \vec{1} \quad . \]
+
+In order to not always hardcode these operations, \em einsum \em can be employed by passing in the two operand matrices, together with a format string, which defines how every dimension will be treated.
+For example \texttt{ik,kj->ij} specifies a normal matrix-matrix-multiplication.
+Here is how these \em column identifiers \em of \texttt{lhs\_string} in the eisnum expression \texttt{<lhs\_string>,<rhs\_string>-><target\_string>} are interpreted:
+
+\begin{table}[h!]
+  \centering
+  \caption{Column Identifiers of Einsum Expression}
+  \label{tab:column_identifiers}
+  \begin{tabular}{lccc}
+    \toprule
+    \textbf{Category} & \textbf{rhs\_string} & \textbf{target\_string} \\
+    \midrule
+    Batch              & Yes    & Yes    \\
+    Kept left          & No     & Yes    \\
+    Contracted         & Yes    & No     \\
+    Summed left        & No     & No     \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+You have to read the table in the following manner.
+Take a column identifier of \texttt{lhs\_string}.
+Then you check whether it is present in \texttt{rhs\_string} and \texttt{target\_string}, and based on these results you assign it to one of the four specified categories, where "Summed left" describes a dimension to reduce.
+
+A similar table can be produced for column identifiers in \texttt{rhs\_string}.
+The column identifiers in \texttt{target\_string} also specify the resulting shape of the operation.
+All of the necessary computations to achieve the desired transformation is performed by the einsum function.
+
 \item \textbf{Focus:} What is your research question? What are you studying precisely?
@@
 \subsection{Related Work}
+I used this~\cite{hptt2017} C++ library
 
@@
 \subsection{Internal Representation of Mock Labels}
-\label{sub:sec:internal}
-
-In Figure~\ref{fig:integer:sets} we convert the mock labels to sorted integer sets.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=\linewidth]{./graphics/integer_sets.pdf}
-  \caption{Conversion of mock camera labels to sorted integer sets. 
-We map each unique token (key) in camera labels to a unique value. 
-Based on these key-value-mappings, we convert camera labels to sorted integer sets. 
-A camera can have different names in different countries. Therefore, repeating IDs reference the same cameras (see, for example, ID=3).} 
-  \label{fig:integer:sets}
-\end{figure}
 

### 2024-12-22 Exam Assignments/paper/paper.tex
paper algorithm

@@
 % \item \textbf{Topic and background:} What topic does the paper deal with? What is the point of departure for your research? Why are you studying this now?
-\item \textbf{Topic and background:} The topic of this paper is tensor transformations, and how to perform them efficiently.
+\item \textbf{Topic and background:} This paper concerns tensor operations, specifically those implemented in the Python library numpy with its \texttt{numpy.einsum} method.
+Tenor operations are very common in todays data driven world, especially with the rise of \em Deep Learning\em .
+Hence, we should ensure that those important methods are implemented efficiently, as \em time \em and \em space complexity \em are a big concern when operating on increasingly bigger data.
+
+% \item \textbf{Focus:} What is your research question? What are you studying precisely?
+\item \textbf{Focus:} The einsum operation implements three fundamental tensor operations: \em Transposition\em , \em Reduction \em and \em Contraction\em .
+This paper tests all of these operations, as well as their combinations, and compares it to a hardcoded implementation written in C++.
+\item \textbf{Method:} To this end, I implemented a bare einsum implementation which implements these three operatios in C++, where one external library for transposing tensors is employed.
+Then, both implementations are compared, and the findings are presented in this paper.
+\item \textbf{Key findings:} The library einsum implementation does a really good job at transposing tensors.
+However, when contraction and especially reduction take up much of the workload, numpy's einsum implementation is remarkably ineffecient, as my C++ implementation executes faster by magnitudes.
+% \item \textbf{Conclusions or implications:} What do these findings mean? What broader issues do they speak to?
+\item \textbf{Conclusions or implications:} These findings emphasize that we cannot trust every implementation of a library to be implemented efficiently, even when speaking of popular libraries like numpy.
+In this specfic case of einsum, I advise the reader not to use it for big data operations which rely heavily on contraction or reduction.
+\end{enumerate}
+
+\end{abstract}
+
+%%
+%% Keywords. The author(s) should pick words that accurately describe
+%% the work being presented. Separate the keywords with commas.
+% \keywords{entity resolution, data cleansing, programming contest}
+\keywords{Tensor Operations, High-Performance Computing, NumPy, Transposition, Reduction, Contraction}
+
+%%
+%% This command processes the author and affiliation and title
+%% information and builds the first part of the formatted document.
+\maketitle
+
+\let\thefootnote\relax\footnotetext{AEPRO 2025, March 1, Jena, Germany.
+Copyright \copyright 2025 for this paper by its authors.
+Use permitted under Creative Commons License Attribution 4.0 International (CC BY 4.0).}
+
+\section{Introduction}
+\subsection{Background}
+The topic of this paper is tensor transformations, and how to perform them efficiently.
 Tensors consist of many dimensions.
@@
 Lastly, we can \em reduce \em entire dimensions by summing over all entries, for example reducing the second dimension of a matrix $M$ is equivalent to
-\[ M_{reduced} = M \cdot \vec{1} \quad . \]
+\[ M_{reduced} = M \cdot \begin{pmatrix}
+  1 \\
+  \vdots \\
+  1
+\end{pmatrix} \quad . \]
 
-In order to not always hardcode these operations, \em einsum \em can be employed by passing in the two operand matrices, together with a format string, which defines how every dimension will be treated.
+In order to not always hardcode these operations, \em einsum \em can be employed by passing in the two operand tensors, together with a format string, which defines how every dimension will be treated.
 For example \texttt{ik,kj->ij} specifies a normal matrix-matrix-multiplication.
@@
 
-\item \textbf{Focus:} What is your research question? What are you studying precisely?
-\item \textbf{Method:} What did you do?
-\item \textbf{Key findings:} What did you discover?
-\item \textbf{Conclusions or implications:} What do these findings mean? What broader issues do they speak to?
+\subsection{Related Work}
+In order to implement einsum, it is necessary to transpose tensors. To this end, I used {this external C++ library}~\cite{hptt2017}.
+
+\subsection{My Contributions}
+In order to have a complete einsum implementation, many steps are necessary, like parsing the input string, transposing and reshaping the operand tensors, reducing them, batch matrix multiply them, and reshape and retranspose the resulting matrix again.
+The details of all of these steps can be found in section \ref{sec:algorithm}.
+Implementing these individual operations, as well as orchestrating all of them, was part of this project.
+
+\subsection{Outline}
+Next, I present the implemented algorithm in more detail, and explain the design choices (section \ref{sec:algorithm}).
+After that, we use this implementation to do tensor operations on NumPy arrays, and compare the running times with NumPy's einsum implementation (section \ref{sec:experiments}).
+To this end, we test different einsum calls, some with much emphasize on individual tensor operations like transposing, reducing or contracting, and other calls which require a combination of these operations.
+Finally, we analyze the findings and draw conclusions (section \ref{sec:conclusions}).
+
+\section{The Algorithm}
+\label{sec:algorithm}
+
+\subsection{Prerequisites}
+At this point, the reader should be familiar how einsum categorizes the dimensions of tensors (based on the format string), and how to treat the tensors accordingly.
+We only focus on two operand and one result tensor, and we only consider the three fundamental tensor operations described above.
+We store all our tensors in \em row-major order\em .
+
+\subsection{Parsing the Input}
+Firstly, we need to ensure the inputs are valid.
+Since we get a generic \texttt{PyObject*} as input, this includes checking if we have three objects: a string, and two NumPy arrays of the same data type.
+This basic parsing is done in \texttt{src/einspeed.cpp}.
+
+~\\
+Next, we have to ensure that the tuple \texttt{(std::string, Tensor<T>, Tensor<T>)} is a valid einsum expression.
+To this end, I implemented \texttt{is\_valid\_einsum\_expression} function found in \texttt{include/compute\_einsum.h}.
+It checks the following five criteria:
+
+\begin{enumerate}
+  \item Every character in target\_string must be present in either lhs\_string or rhs\_string.
+  \item There are no duplicate characters in lhs\_string, rhs\_string or target\_string.
+  \item The size of lhs\_string must match the number of dimensions of lhs\_tensor. The same for rhs\_string and rhs\_tensor.
+  \item Common column identifiers in lhs\_string and rhs\_string must correspond to matching dimension sizes in lhs\_tensor and rhs\_tensor.
 \end{enumerate}
 
-\end{abstract}
+In case the tuple passed all these checks, the function won't throw any error, and hence we can now proceed with the function \texttt{compute\_einsum} knowing we have a valid einsum expression.
 
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{entity resolution, data cleansing, programming contest}
+\subsection{Computing Einsum}
+\subsubsection{Categorizing Dimensions}
+We now investigate the orchestrating function \texttt{compute\_einsum} further.
+The first step is to categorize each dimension of the two input tensors into one of the four categories \em batch, kept left, contracted \em or \em summed left \em in case of the left hand side tensor, similarly for the right hand side.
+We do this by simply looping over all so called \em column identifiers \em specified by the format string (columnn identifiers are basically just the individual characters in the string) and checking if this identifier is also present in the \texttt{rhs\_string} and \texttt{target\_string}.
+Based on these checks we assign the categories, see table~\ref{tab:column_identifiers}.
+We do similarly for the right hand side.
+We store the indices in appropriate arrays, for example, all the dimension indices which are batch dimensions in the \texttt{lhs\_tensor} will be stored in \texttt{batch\_dims\_lhs}.
 
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
-\maketitle
+\subsubsection{Transposing and Reshaping Operand Tensors}
+Now that we have categorized all the dimensions, we bring both tensors into the canonical form
 
-\let\thefootnote\relax\footnotetext{AEPRO 2025, March 1, Jena, Germany.
-Copyright \copyright 2025 for this paper by its authors.
-Use permitted under Creative Commons License Attribution 4.0 International (CC BY 4.0).}
+\[ \text{lhs\_tensor}[\vec{b}, \vec{k_l}, \vec{c}, \vec{s_l}] \ , \ \text{rhs\_tensor}[\vec{b}, \vec{k_r}, \vec{c}, \vec{s_r}] \quad . \]
 
-\section{Introduction}
+One very important detail is that the order of the \em batch \em and \em contracted \em dimensions after transposing have the same ordering in the operand tensors.
+To illustrate this, say we have the einsum expression \texttt{einsum("ab,ba->", lhs\_matrix, rhs\_matrix)}.
+Naively, there is no need to transpose both matrices, as both of them are in the canonical form (every dimension is a contraction dimension).
+However, the orderings of these contracted dimensions do not match, hence we do need to transpose a matrix, either the left or the right hand side.
 
-\subsection{Background}
+~\\
+To this end, the algorithm reorders the indices-lists for the right hand side such that the resulting order of column identifiers matches the left hand side.
+For this purpose, we employ the function \texttt{column\_identifiers\_like}.
 
-\subsection{Related Work}
-I used this~\cite{hptt2017} C++ library
+~\\
+Furthermore, it is worth mentioning that there are different ways to transpose tensors.
+One way is the "lazy" way, where we just alter meta attributes of the tensor, in particular we alter the \em strides\em , which specify by "how much" we have to multiply an index of a given dimension in order to get the physical index to access the element in the underlying data array.
+This lazy way of doing it is very fast, as there is no need to move large amount of data.
+However, there is one big problem: If a tensor has non-canonical strides, we cannot reshape it.
+This is very bad, because we actually do want to reshape the tensors afterwards.
+Hence, we need to use the other way of transposing tensors: In he "eager" approach we basically move the data such that it is aligned with the new ordering of the dimensions.
+This is obviously much more expensive time and space complexity wise, but there is at least one benefit: The data is well aligned for the next operation (after reshaping): The following reduction operation will benefit from a stride access pattern of 1.
 
-\subsection{Our Contributions}
+\subsubsection{Reducing Operand Tensors}
+Now that we have both operand tensors in the desired shape, we have a fix plan on how to operate on these tensors, which means the following precedure is identical to all possible inputs.
+Only the reshaping and transposing of the result matrix at the very end needs specific treatment.
 
-\subsection{Outline}
+We start by reducing the fourth dimension of both operand tensors.
+This is fairly efficient as we have efficient access patterns as described earlier.
+After that, the two tensors will have the following form:
 
-\section{The Algorithm}
+\[ \text{lhs\_tensor}[\vec{b}, \vec{k_l}, \vec{c}] \ , \ \text{rhs\_tensor}[\vec{b}, \vec{k_r}, \vec{c}] \quad . \]
 
-\subsection{Internal Representation of Mock Labels}
+It is worth mentioning that it might be possible to combine all of these operations so far (transposing, reshaping, reducing fourth dimension) into one big one.
+This will at least save some memory, as now we have three tensors for each operand tensor in memory: The original one which we have to leave untouched, the transposed one, as well as the reduced one.
+I still decided to implement it this way, as this approach is more modular.
+Furthermore, I initially implemented a transpose method myself, but I realized that it was the bottleneck of my implementation.
+Hence, I decided to use an external library for transposing, such that the performance of my implementation is not limited by a poor implemented transposing function.
 
-\subsection{Efficient Preprocessing of Input Data}
-\label{sub:sec:preprocessing}
+Of course we can now free both transposed tensors after the reduction operation, which we also do.
+However, I figured that it might be beneficial to actually keep some memory, preferably from the bigger tensor, in order to later reuse it for the result of the following batch matrix multiplication operation.
 
-The following findings are important to speed up preprocessing of the input data:
+\subsubsection{Batch Matrix Multiplication}
+Now we take care of contracting the appropriate dimensions by performing a batch matrix multiplication.
+Note, though, that the contraction dimensions are the last one in each operand tensor (the third dimension), and hence we actually perform a \em batch matrix matmul transpose \em operation, i.e.
+we treat the rhs operand batch matrices as being transposed.
+I chose this implementation, as this way we have a stride of one when multiplying individual matrices of a batch both in the lhs and rhs matrix, with respect to the inner most loop.
+Consequently, this ensures excellent spatial locality and hence we make perfect use of the cache.
+The result of this operation will be the three dimensiol tensor \texttt{bmm\_result}, with dimension sizes
 
+\[ \text{bmm\_result} = [b, k_l, k_r] \quad . \]
+
+As hinted, we might store the result of this operation in previously allocated memory, all provided that the memory region is big enough.
+Otherwise we do have to allocate new memory.
+The question whether our available memory is sufficient for the new data depends on on the dimension sizes of the operand tensors and their categorized dimensions: We have
+
+\begin{align*}
+|\text{lhs\_tensor}| &= b \cdot k_l \cdot c \cdot s_l \\
+|\text{rhs\_tensor}| &= b \cdot k_r \cdot c \cdot s_r \\
+|\text{bmm\_result}| &= b \cdot k_l \cdot k_r \quad ,
+\end{align*}
+
+and hence the memory will be sufficiently big enough iff
+
+\[ c \cdot s_l \geq k_r \]
+
+in case that the lhs tensor is the bigger one of the two (similar if rhs tensor is bigger).
+
+\subsubsection{Reshaping and Transposing the Result}
+Now, there is only one final step: Bringing \texttt{bmm\_result} in the correct shape specified by the format string.
+To this end, we have to firstly split the current three dimensions of our result tensor into the individual ones.
+Hence, we we look at our inital tensors and their column identifiers again to get the dimension sizes.
+
+For example, so split the batch dimension, we check the the batch dimensions of \texttt{lhs\_tensor} and store their specific dimension sizes in the same order as they are present in the tensor.
+We do this, as our \texttt{bmm\_result} tensor has the same ordering of its data with respect to the batch dimension as \texttt{lhs\_tensor}.
+Note that we specifically transposed \texttt{rhs\_tensor} to match the batch and contraction order of \texttt{lhs\_tensor}.
+
+After doing similar for the remaining two dimensions, we reshape our tensor accordingly.
+Now we have the result tensor as desired, only with potential wrong ordering of the dimensions.
+Hence, we need a final transposition.
+But instead of creating a whole new tensor and doing an expensive transpose call, we are very lazy and just return the current tensor as is, only with accordingly altered strides metadata.
+
+To illustrate this, say we have a matrix a 5x10 matrix $M$, and we want to transpose it. When accessing $M[i][j]$ we have to calculate the physical index of the row-major order stored matrix. Hence:
+
+\[ \text{physical\_index}_M(i, j) = 10 \cdot i + 1 \cdot j \quad . \]
+
+When we now transpose this matrix, we can alter this translation function instead of actually moving the data. In this example:
+
+\[ \text{physical\_index}_{M^T}(i, j) = 1 \cdot i + 10 \cdot j \quad . \]
+
+This means that instead of moving data, we transpose the coefficients of this function.
+These coefficients are stored as \em strides \em in the metadata of a tensor.
+NumPy does so as well, and hence we employ this by doing this final transposition the lazy manner by setting these NumPy metadata attributes.
+
+This implementation choice is not "cheating", as NumPy's einsum implementation does the same lazy transposition.
+Hence, it is only fair to do so myself, as this way we get more meaningful comparisons.
+To prove my point, run the following lines of Python:
+
+\begin{verbatim}
+import numpy as np
+matrix = np.array([[1,2],[3,4]])
+print(matrix.strides)
+transposed = np.einsum("ab,c->ba", matrix, np.array([1]))
+print(transposed.strides)
+\end{verbatim}
+
+You will likely see that the strides got altered simply.
+
+\subsection{Transposition Method}
+Let us now take a closer look at the smaller "puzzle pieces" making up the main \texttt{comput\_einsum} function, starting with the tensor transposition method.
+As already discussed, I chose to use an external library for this purpose, and hence there is not much to analyze, as we just pass the control flow to the library.
+It is important to note though that we allocate memory for another tensor each call, as this is necessary for the library call, as well as ensuring that the original tensor stays constant.
+Hence the constant specifier in the method declaration.
+% TODO: write about how efficient it is, vectoriziation, multi threading etc.
+
+\subsection{Reshaping Method}
+The implementation of this method is very sraight forward.
+As discussed, when calling this method, we ensure that the tensor has canonical strides, as in this case we only have to change the meta attributes \texttt{ndim} and \texttt{shapes}.
+There isn't else much to say, this is a simple and fast method which will run in $O(1)$.
+
+\subsection{Reduction Method}
+This method is not very generic, as it will only reduce the last dimension of the tensor.
+By doing so, I kept the implementation simple yet it is sufficient for our computations.
+At the beginning we have to allocate memory for the resulting tensor.
+This is the first method we look at which has an impact on performance, thus we should be cautious with implementation choices.
+Hence, we write the folowing compiler-optimization-friendly code:
+
+\begin{verbatim}
+#pragma omp parallel for
+for (size_t i = 0; i < target_size; i++) {
+    T sum = 0;
+    size_t start_idx = i * last_dimension;
+    size_t end_idx = start_idx + last_dimension;
+
+    for (size_t k = start_idx; k < end_idx; k++) {
+        sum += this->data[k];
+    }
+
+    target_data[i] = sum;
+}
+\end{verbatim}
+
+Clearly, this code is very efficient.
+There are two minor problems though: First, there is a potential risk of overflow when summing up the values.
+Second, we parallelize the outer most loop, which is a wise choice, but in very rare cases when \texttt{target\_size} equals one, like when reducing a one-dimensional vector, there won't be any parallelization.
+
+To quickly address both issues, let me point out that, first, in the case of an overflow, there is not much that can be done other than throwing an error.
+However, implementing this would create overhead, and hence, it is ultimately the responsibility of the user to ensure correct computations.
+Second, we just accept that we are unlucky if such a case happens.
+There is also no quick fix, as both loops cannot get collapsed into one because of the line \texttt{T sum = 0;}, and also we would need to write template specializations for every possible type \texttt{T}, as else the compiler won't compile omp reduction clause.
+In the end, it comes down to keeping the implementation simple (\em KISS\em ).
+
+% TODO: Explain vectorized versions
+Despite that, I actually tried to optimize this method by using explicit vectorization in specialized template methods.
+Interestingly, they don't seem to have any performance benefits compared to the generic implementation, which is probably due to the fact that modern compilers will vectorize the generic implementation as well.
+The code is pretty much self explanatory, we do employ an expensive \texttt{loadu} instead of a \texttt{load} vector instruction, as the data won't be aligned generally.
+
+\subsection{Batch Matrix Multiplication}
+Let us now investigate the final sub-function \texttt{batch\_matrix\_matmul\_transpose} found in \texttt{include/}.
+It is basically very straight forward, also note that we treat the rhs operand batch matrices as transposed, which allows for small strides in the inner most loop.
+We begin by allocating the memory needed for the output, but before allocating new memory, we check if the available memory we reserved is sufficiently large to hold the new data, and if so use it instead.
+Next, loop over the batch dimension, loop over the rows of lhs tensor, and loop over the rows of rhs tensor (as it is transposed).
+We collapse these three loops into one, and inside every iteration, which specifies one entry in the output tensor, we perform an inner product between the two last dimensions of the operand tensors at the specified index.
+In code it looks like this:
+
+\begin{verbatim}
+#pragma omp parallel for collapse(3)
+for (size_t b = 0; b < batch_size; ++b) {
+    for (size_t i = 0; i < rows; ++i) {
+        for (size_t j = 0; j < cols; ++j) {
+            const T* lhs_row = lhs.data + b * rows * inner_dim + i * inner_dim;
+            const T* rhs_col = rhs.data + b * cols * inner_dim + j * inner_dim;
+            T* result_row = result_data + b * rows * cols + i * cols;
+
+            T sum = T(); // Initialize sum as zero for type T
+            for (size_t k = 0; k < inner_dim; ++k) {
+                sum += lhs_row[k] * rhs_col[k];
+            }
+            result_row[j] = sum;
+        }
+    }
+}
+\end{verbatim}
+
+We face the \textbf{exact} problems we faced at the reduction method, and my responses are the same, we just keep the implementation simple for the sake of demonstration.
+This function could be manually vectorized as well, but the compiler should do a decent job at optimizing this function, and hence we shouldn't face a big performance penalty.
+
+% overflow, data types, combining methods
+\subsection{Parameter Specification}
+As hinted multiple times, my einsum implementation is not very complete but rather specific.
+For example, NumPy's implementation also supports passing only one tensor as an argument, passing scalars as arguments, or calculating traces just to name a few features.
+
+For the sake of simplicity my implementation only accepts two (multi-dimensional) NumPy arrays of the same data type, and only implements transposition, reduction, and contraction.
+It is able to return a scalar value though, and there is also a workaround to only passing in one NumPy array: Just pass in \texttt{np.array([1], dtype)} as a second operand, and specify the only dimension as a reduction one.
+This will have the same effect as only passing in the original tensor, and there is also little overhead, since in my implementation I check whether one operand tensor is a scalar with value \texttt{T(1)}, and if so skip the batch matrix multiplication.
+
+Also, supported data types are limited by those supported by the external transposition library.
+Hence, integer value are actually not supported.
+The only supported data types are \texttt{np.float64} and \texttt{np.complex128}.
+
+\subsection{Possible Improvements}
+My implementation lacks some functionality to be used as a library itself.
+But the implementation could be a good starting point for building one, and because it is written fairly modular, it can be augmented without much effort.
+
+There are several Improvements one could make, but the most obvious ones are:
 \begin{itemize}
-\item Reading many small files concurrently, with multiple threads (compared to a single thread), takes advantage of the internal parallelism of SSDs and thus leads to higher throughput \cite{Zhuang2016}.
-
-\item C-string manipulation functions are often significantly faster than their C++ pendants.
-For example, locating substrings with \texttt{strstr} is around five times faster than using the C++ \texttt{std::string} function \texttt{find}.
-
-\item Hardcoding regular expressions with \emph{while, for, switch} or \emph{if-else} statements results in faster execution times than using standard RegEx libraries, where regular expressions are compiled at runtime into state machines.
-
-\item Changing strings in place, instead of treating them as immutable objects, eliminates allocation and copying overhead.
-
+  \item Augmenting supported data types. Therefore a new implementation of the transposition method is needed.
+  \item More support for parameters, maybe also tensors of different data types, different operations, etc.
+  \item More optimized main method, more heuristics, etc.
+  \item More optimized sub methods, improving parallelization, vectorization by checking for aligned data, etc.
 \end{itemize}
@@
 \section{Experiments}
+\label{sec:experiments}
 
@@
 \section{Conclusions}
+\label{sec:conclusions}
 

### 2024-12-22 Exam Assignments/Code/einspeed/einspeed.py

@@ -2,18 +2,18 @@ import einspeed
 import numpy as np
 import time
 
-dtype = np.complex128
+dtype = np.float32
 epsilon = 1e-2 # we need to be generous here, as there is a big computing error on large tensors
-size = 50
-matrix_lhs = (1+np.random.rand(size, size, size)).astype(dtype)
-matrix_rhs = (1+np.random.rand(size, size, size)).astype(dtype)
+size = 40
+matrix_lhs = (1+np.random.rand(size, 4 * size, size)).astype(dtype)
+matrix_rhs = (1+np.random.rand(size, size, 3 * size)).astype(dtype)
 # matrix_lhs = np.array([[[1,2],[3,4]],[[5,6], [7,8]]], dtype)
 # matrix_rhs = matrix_lhs
 # matrix_rhs = np.array([1], dtype)
 copy = np.copy(matrix_lhs)
 vector = np.array([1, 2, 3], dtype)
 # string = "abc,def->" # einsum implementation is HORRENDOUS at reducing
-string = "abc,def->"
+string = "aik,bkj->bij"
 
 start_time = time.time()
 correct = np.einsum(string, matrix_lhs, matrix_rhs)
@@ -25,4 +25,5 @@ result = einspeed.einsum(string, matrix_lhs, matrix_rhs)
 print(f"result in {time.time() - start_time} seconds")
 print("Correct: ", (np.abs((result-correct)) < epsilon).all() and (matrix_lhs==copy).all())
 print(result.strides)
-print(correct, result)
+# print(correct, result)
+print(np.max(np.abs((result - correct))))

### 2024-12-22 Exam Assignments/Code/einspeed/include/blas.h

@@ -46,8 +46,9 @@ Tensor<T> batch_matrix_matmul_transpose(const Tensor<T>& lhs, const Tensor<T>& r
                 const T* lhs_row = lhs.data + b * rows * inner_dim + i * inner_dim;
                 const T* rhs_col = rhs.data + b * cols * inner_dim + j * inner_dim;
                 T* result_row = result_data + b * rows * cols + i * cols;
-
                 T sum = T(); // Initialize sum as zero for type T
+
+                #pragma omp simd
                 for (size_t k = 0; k < inner_dim; ++k) {
                     sum += lhs_row[k] * rhs_col[k];
                 }

### 2024-12-22 Exam Assignments/Code/einspeed/include/numpy_types.h

@@ -9,6 +9,7 @@ template <typename T>
 constexpr int numpy_type_of() {throw std::runtime_error("Unsupported data type for NumPy array");}
 
 template<> constexpr int numpy_type_of<long>() {return NPY_INT64;}
+template<> constexpr int numpy_type_of<float>() {return NPY_FLOAT32;}
 template<> constexpr int numpy_type_of<double>() {return NPY_FLOAT64;}
 template<> constexpr int numpy_type_of<std::complex<double>>() {return NPY_COMPLEX128;}
 
@@ -20,6 +21,11 @@ struct numpy_to_cpp_type;
 template <>
 struct numpy_to_cpp_type<NPY_INT64> {
     using type = int64_t;
+};
+
+template<>
+struct numpy_to_cpp_type<NPY_FLOAT32> {
+    using type = float;
 };
 
 template <>
@@ -51,6 +57,7 @@ const int NUMPY_TYPES[] = {
 
 constexpr const int SUPPORTED_NUMPY_TYPES[] = {
     // NPY_INT64, // no int64, as hptt does not support it
+    NPY_FLOAT32,
     NPY_FLOAT64,
     NPY_COMPLEX128
 };

### 2024-12-22 Exam Assignments/Code/einspeed/include/tensor.h

@@ -349,7 +349,7 @@ Tensor<float> Tensor<float>::reduce() const {
         __m256 vec_sum = _mm256_setzero_ps(); // Initialize to zero
         
         for (size_t k = start_idx; k < start_idx + last_dimension; k += 8) {
-            // Load 4 values at once
+            // Load 8 values at once
             __m256 vec_vals = _mm256_loadu_ps(&this->data[k]);
             vec_sum = _mm256_add_ps(vec_sum, vec_vals); // Add the values
         }

### 2024-12-28 Exam Assignments/paper/paper.tex
verbatim -> lstlisting

@@
 \usepackage{booktabs}
+\usepackage{pgfplotstable}
+\usepackage{fancyvrb}
+\usepackage{listings}
+\usepackage{etoolbox}
+
+\lstset{
+    breaklines=true       % Enable line breaking
+}
+% Define a custom lstlisting command that avoids page breaks
+\BeforeBeginEnvironment{lstlisting}{\noindent\begin{minipage}{\linewidth}}
+\AfterEndEnvironment{lstlisting}{\end{minipage}}
+
+% TODO: write acronyms like cpu, gpu capitalized
+
+\setlength{\emergencystretch}{3em} % Allow LaTeX to stretch lines slightly before exceeding width
+\tolerance=1000                   % Allow more "badness" in line breaks before exceeding width
+\hbadness=10000                   % Report warnings for badness levels higher than this value
+
+\pretolerance=2000  % Try harder to find good breaks before a paragraph
+\tolerance=3000     % Relax line-breaking rules further
+\emergencystretch=5em % Allow additional stretch space
+\renewcommand{\texttt}[1]{\begingroup\ttfamily\sloppy\hbadness=10000 #1\endgroup}
+\linepenalty=1000 % Increase the penalty for exceeding width
+\sloppy
 
@@
 \item \textbf{Topic and background:} This paper concerns tensor operations, specifically those implemented in the Python library numpy with its \texttt{numpy.einsum} method.
-Tenor operations are very common in todays data driven world, especially with the rise of \em Deep Learning\em .
+Tensor operations are very common in todays data driven world, especially with the rise of \em Deep Learning\em .
 Hence, we should ensure that those important methods are implemented efficiently, as \em time \em and \em space complexity \em are a big concern when operating on increasingly bigger data.
@@
 
+\noindent
 In order to not always hardcode these operations, \em einsum \em can be employed by passing in the two operand tensors, together with a format string, which defines how every dimension will be treated.
@@
 \end{table}
-
+\noindent
 You have to read the table in the following manner.
@@
 
-~\\
 Next, we have to ensure that the tuple \texttt{(std::string, Tensor<T>, Tensor<T>)} is a valid einsum expression.
@@
 
+\noindent
 In case the tuple passed all these checks, the function won't throw any error, and hence we can now proceed with the function \texttt{compute\_einsum} knowing we have a valid einsum expression.
@@
 Now that we have categorized all the dimensions, we bring both tensors into the canonical form
-
 \[ \text{lhs\_tensor}[\vec{b}, \vec{k_l}, \vec{c}, \vec{s_l}] \ , \ \text{rhs\_tensor}[\vec{b}, \vec{k_r}, \vec{c}, \vec{s_r}] \quad . \]
-
 One very important detail is that the order of the \em batch \em and \em contracted \em dimensions after transposing have the same ordering in the operand tensors.
@@
 
-~\\
 To this end, the algorithm reorders the indices-lists for the right hand side such that the resulting order of column identifiers matches the left hand side.
@@
 
-~\\
 Furthermore, it is worth mentioning that there are different ways to transpose tensors.
@@
 This is fairly efficient as we have efficient access patterns as described earlier.
-After that, the two tensors will have the following form:
-
+After that, the two tensors will have the following form
 \[ \text{lhs\_tensor}[\vec{b}, \vec{k_l}, \vec{c}] \ , \ \text{rhs\_tensor}[\vec{b}, \vec{k_r}, \vec{c}] \quad . \]
-
 It is worth mentioning that it might be possible to combine all of these operations so far (transposing, reshaping, reducing fourth dimension) into one big one.
@@
 The result of this operation will be the three dimensiol tensor \texttt{bmm\_result}, with dimension sizes
-
 \[ \text{bmm\_result} = [b, k_l, k_r] \quad . \]
-
 As hinted, we might store the result of this operation in previously allocated memory, all provided that the memory region is big enough.
@@
 The question whether our available memory is sufficient for the new data depends on on the dimension sizes of the operand tensors and their categorized dimensions: We have
-
 \begin{align*}
@@
 \end{align*}
-
 and hence the memory will be sufficiently big enough iff
-
 \[ c \cdot s_l \geq k_r \]
-
 in case that the lhs tensor is the bigger one of the two (similar if rhs tensor is bigger).
@@
 To illustrate this, say we have a matrix a 5x10 matrix $M$, and we want to transpose it. When accessing $M[i][j]$ we have to calculate the physical index of the row-major order stored matrix. Hence:
-
 \[ \text{physical\_index}_M(i, j) = 10 \cdot i + 1 \cdot j \quad . \]
-
 When we now transpose this matrix, we can alter this translation function instead of actually moving the data. In this example:
-
 \[ \text{physical\_index}_{M^T}(i, j) = 1 \cdot i + 10 \cdot j \quad . \]
-
 This means that instead of moving data, we transpose the coefficients of this function.
@@
 To prove my point, run the following lines of Python:
-
-\begin{verbatim}
+\begin{lstlisting}[language=Python]
 import numpy as np
@@
 print(transposed.strides)
-\end{verbatim}
-
+\end{lstlisting}
 You will likely see that the strides got altered simply.
@@
 
-\begin{verbatim}
-#pragma omp parallel for
+\begin{lstlisting}[language=C++]
+#pragma omp parallel for simd aligned(data: 32) aligned(target_data: 32)
 for (size_t i = 0; i < target_size; i++) {
@@
     for (size_t k = start_idx; k < end_idx; k++) {
-        sum += this->data[k];
+        sum += data[k];
     }
@@
 }
-\end{verbatim}
-
+\end{lstlisting}
+\\\\
 Clearly, this code is very efficient.
@@
 
-% TODO: Explain vectorized versions
 Despite that, I actually tried to optimize this method by using explicit vectorization in specialized template methods.
 Interestingly, they don't seem to have any performance benefits compared to the generic implementation, which is probably due to the fact that modern compilers will vectorize the generic implementation as well.
-The code is pretty much self explanatory, we do employ an expensive \texttt{loadu} instead of a \texttt{load} vector instruction, as the data won't be aligned generally.
+The code is pretty much self explanatory, we do employ an efficient \texttt{load} instead of a \texttt{loadu} vector instruction, as the data will be aligned because of the previous transpose operation.
+Here is the code for tensors of type \texttt{float}:
+
+\begin{lstlisting}[language=C++]
+#pragma omp parallel for
+for (size_t i = 0; i < target_size; i++) {
+    float sum = 0;
+    size_t start_idx = i * last_dimension;
+
+    // Use AVX SIMD for vectorized summation
+    __m256 vec_sum = _mm256_setzero_ps(); // Initialize to zero
+
+    for (size_t k = start_idx; k < start_idx + last_dimension; k += 8) {
+        // Load 8 values at once
+        __m256 vec_vals = _mm256_load_ps(&this->data[k]);
+        vec_sum = _mm256_add_ps(vec_sum, vec_vals); // Add the values
+    }
+
+    // Sum the partial results
+    sum += vec_sum[0] + vec_sum[1] + vec_sum[2] + vec_sum[3]
+        +  vec_sum[4] + vec_sum[5] + vec_sum[6] + vec_sum[7];
+
+    target_data[i] = sum;
+}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
-#pragma omp parallel for collapse(3)
+\begin{lstlisting}[language=C++]
+#pragma omp parallel for simd aligned(result_data: 32) aligned(lhs_data: 32) aligned(rhs_data: 32) collapse(3)
 for (size_t b = 0; b < batch_size; ++b) {
@@
         for (size_t j = 0; j < cols; ++j) {
-            const T* lhs_row = lhs.data + b * rows * inner_dim + i * inner_dim;
-            const T* rhs_col = rhs.data + b * cols * inner_dim + j * inner_dim;
+            const T* lhs_row = lhs_data + b * rows * inner_dim + i * inner_dim;
+            const T* rhs_col = rhs_data + b * cols * inner_dim + j * inner_dim;
             T* result_row = result_data + b * rows * cols + i * cols;
+            T sum = T(); // Initialize sum as zero for type T
 
-            T sum = T(); // Initialize sum as zero for type T
+            #pragma omp simd
             for (size_t k = 0; k < inner_dim; ++k) {
@@
 }
-\end{verbatim}
-
+\end{lstlisting}
+\\\\
 We face the \textbf{exact} problems we faced at the reduction method, and my responses are the same, we just keep the implementation simple for the sake of demonstration.
@@
   \item More support for parameters, maybe also tensors of different data types, different operations, etc.
-  \item More optimized main method, more heuristics, etc.
+  \item More optimized main method, more heuristics like only transposing when needed, etc.
   \item More optimized sub methods, improving parallelization, vectorization by checking for aligned data, etc.
+  \item Utilize gpu for parallelization.
 \end{itemize}
@@
 
-Table~\ref{tab:results} shows the running times of the resolution step of the five best placed teams.
+Now that everything is implemented, we test our einsum implementation and compare it to NumPy's.
+To this end, I have written \texttt{einspeed.py}, which compares both implementations on a variety of test cases focussing on different aspects of einsum calculations.
+
+\subsection{Correctness}
+We start by assessing the correctness.
+Our implementation seems to operate very precisely, as for small problem sizes both implementations yield the same output.
+For larger problem sizes, however, there is a growing discrepancy in the entries of the result tensor.
+This discrepancy can grow arbitrary large for increasing tensor sizes, which is why a relative measure is more meaningful.
+We calculate it like this:
+
+\begin{lstlisting}[language=Python]
+maximum_relative_discrepancy = np.max(np.abs((result - correct) / correct))
+\end{lstlisting}
+\\\\
+Note that per constraction no entry of \texttt{correct} is zero.
+It turns out that this relative discrepancy stays below a value of approximately \texttt{1e-4}, which is very commendable.
+One might ask where this error arises, and I my answer to that will be that numerical operations of floating point numbers don't form a mathematical \em field\em .
+Consequently, the order of operation will have an effect on the final output, and I think that my implementation employs a different order than NumPy's.
+
+Ultimately, however, neither implementanion is exactly correct, and my implementation isn't necessarily worse.
+In fact, on some test cases it was closer to a hard coded answer, but this doesn't have to mean anything, as these computations will suffer from numerical errors as well.
+Just to give an example, one can test this by running:
+
+\begin{lstlisting}[language=Python]
+lhs_vector = np.random.rand(100)
+rhs_vector = np.random.rand(100)
+print(np.einsum("i,i->", lhs_vector, rhs_vector))
+print(einspeed.einsum("i,i->", lhs_vector, rhs_vector))
+print(lhs_vector @ rhs_vector)
+\end{lstlisting}
+\\\\
+In conclusion, my implementation seems to operate correctly, and hence we now focus on performance.
+
+\subsection{Performance}
+When running the test cases, is is apparent that NumPy did a really good job at implementing einsum, as it outperforms my multi-threaded implementation on many test cases.
+I think this is mostly due to the fact that my memory management is not the most efficient as I use a lot of memory, and NumPy probably uses a bit more sophisticated algorithms for the computations as well.
+Notably, there seems to be only a constant factor that separates my implementation from NumPy's, and this factor is roughly 20.
+
+These findings emphasize that the developers behind NumPy really know what they are doing.
+Hence, it is no surprise that it is very challenging to implement certain methods more efficient than they have.
+Still, it is very interesting to deeper understand the mechanism behind some implementations, and to test out the limits of computations myself.
+
+Very interestingly, NumPy's einsum function seems to \textbf{really} struggle at big reduction operations, which is very surprising, as they are very simple to implement.
+This is one of the key findings of this paper.
+There seems to be a time complexity issue with its implementation, as with growing operand sizes the running time becomes horrendously bad.
+To give evidence, we execute this einsum expression
+\begin{lstlisting}[language=Python]
+einsum("abc,def->", lhs_tensor, rhs_tensor)
+\end{lstlisting}
+at different operand scales, and compare both implementations. The results are shown in table~\ref{tab:results}.
 
 \begin{table}[htbp]
-  \caption{Comparison of the F-measure and the running times of the resolution step of the five best placed teams.
-The input data for the resolution step consisted of 29{,}787 in JSON formatted e-commerce websites.
-Measurements were taken on a
-laptop running Ubuntu 19.04 with 16 GB of RAM and two Intel Core i5-4310U CPUs.
-The underlying SSD was a 500\,GB 860 EVO mSATA.
-We cleared the page cache, dentries, and inodes before each run to avoid reading the input data from RAM instead of the SSD.}
+  \caption{Comparison of einsum running times.
+Both NumPy's and my einspeed implementation are tested at varying problem sizes.
+The problem size specifies the number of entries in the two input tensors.
+Running times are measured in seconds, the speedup describes the fraction $\frac{numpy\_time}{einspeed\_time}$.
+  Measurements were taken on a PC running Ubuntu 24.04 with 32 GB of RAM and an AMD Ryzen 5 5600G CPU (6 Cores, 12 Threads).}
   \label{tab:results}
 \resizebox{\columnwidth}{!}{
-  \begin{tabular}{lcrr}
-    \toprule
-    Team& Language & F-measure & Running time (s)\\
-    \midrule
-	PictureMe (\textbf{this paper}) &C++& 0.99 & \textbf{0.61}\\
-    DBGroup@UniMoRe &Python& 0.99 & 10.65\\
-    DBGroup@SUSTech &C++& 0.99 & 22.13\\
-    eats\_shoots\_and\_leaves &Python& 0.99 & 28.66\\
-    DBTHU &Python& 0.99& 63.21\\
-  \bottomrule
-\end{tabular}
+  \pgfplotstabletypeset[
+    col sep=comma,
+    every head row/.style={% Styling for the header row
+      before row=\toprule, 
+      after row=\midrule},
+    every last row/.style={% Styling for the last row
+      after row=\bottomrule},
+  ]{src/scaling_study.csv}
 }
@@
 \label{sec:conclusions}
+In this paper, I implemented the einsum function from scratch, in order to evaluate the existing NumPy implementanion.
+Despite the fact that I certainly could have done better, I was pretty satisfied with my work at the end.
+The fact that my implementation can "keep up" with NumPy's contributed to this.
+Yet, I couldn't outperform NumPy generally, only in very specific cases.
+To be honest, that was predictable.
+Therefore, these findings emphasize the quality of of libraries like NumPy, but they also show that we should be cautious about performance when using these functions for special purposes, like big reduction operations.
+Based on my results I advise the reader not to use NumPy's einsum implementation for big reduction operations, as this will result in a serious bottleneck.
+One should hardcode this operation instead, or use a different einsum implementation.
+For the latter case, one might use the results of this paper as a starting point to build a better einsum implementation.
 

### 2024-12-28 Exam Assignments/Code/einspeed/einspeed.py

@@ -1,29 +1,134 @@
 import einspeed
 import numpy as np
 import time
+import csv
 
-dtype = np.float32
-epsilon = 1e-2 # we need to be generous here, as there is a big computing error on large tensors
-size = 40
-matrix_lhs = (1+np.random.rand(size, 4 * size, size)).astype(dtype)
-matrix_rhs = (1+np.random.rand(size, size, 3 * size)).astype(dtype)
-# matrix_lhs = np.array([[[1,2],[3,4]],[[5,6], [7,8]]], dtype)
-# matrix_rhs = matrix_lhs
-# matrix_rhs = np.array([1], dtype)
-copy = np.copy(matrix_lhs)
-vector = np.array([1, 2, 3], dtype)
-# string = "abc,def->" # einsum implementation is HORRENDOUS at reducing
-string = "aik,bkj->bij"
+def benchmark_einsum(format_string, lhs_shape, rhs_shape, dtype = np.float32, test_case = "Default", epsilon = 1e-4):
+    lhs_tensor = (1+np.random.rand(*lhs_shape)).astype(dtype)
+    rhs_tensor = (1+np.random.rand(*rhs_shape)).astype(dtype)
+    
+    lhs_copy = np.copy(lhs_tensor)
+    rhs_copy = np.copy(rhs_tensor)
 
-start_time = time.time()
-correct = np.einsum(string, matrix_lhs, matrix_rhs)
-print(f"correct result in {time.time() - start_time} seconds")
-print(correct.strides)
+    start_time = time.time()
+    correct = np.einsum(format_string, lhs_tensor, rhs_tensor)
+    numpy_time = time.time() - start_time
 
-start_time = time.time()
-result = einspeed.einsum(string, matrix_lhs, matrix_rhs)
-print(f"result in {time.time() - start_time} seconds")
-print("Correct: ", (np.abs((result-correct)) < epsilon).all() and (matrix_lhs==copy).all())
-print(result.strides)
-# print(correct, result)
-print(np.max(np.abs((result - correct))))
+    start_time = time.time()
+    result = einspeed.einsum(format_string, lhs_copy, rhs_copy)
+    einspeed_time = time.time() - start_time
+
+    maximum_relative_discrepancy = np.max(np.abs((result - correct) / correct))
+
+    print(f"Test Case: {test_case} {str(dtype)}")
+    print("Correct: ", (maximum_relative_discrepancy < epsilon) and (lhs_tensor==lhs_copy).all() and (rhs_tensor==rhs_copy).all())
[…]

### 2024-12-28 Exam Assignments/Code/einspeed/include/blas.h

@@ -3,7 +3,6 @@
 
 #include "tensor.h"
 
-
 template <typename T>
 Tensor<T> batch_matrix_matmul_transpose(const Tensor<T>& lhs, const Tensor<T>& rhs, T* available_memory, size_t available_memory_size) {
     // Check dimensions
@@ -35,16 +34,20 @@ Tensor<T> batch_matrix_matmul_transpose(const Tensor<T>& lhs, const Tensor<T>& r
     }    
     else {
         if (available_memory) delete[] available_memory;
-        result_data = new T[result_size];
+        posix_memalign((void**)&result_data, 32, result_size * sizeof(T));
     }
 
+    // data of lhs and rhs used for the omp clause
+    const T* lhs_data = lhs.data;
+    const T* rhs_data = rhs.data;
+
     // Perform batch matrix multiplication with transpose using OpenMP
-    #pragma omp parallel for collapse(3)
+    #pragma omp parallel for simd aligned(result_data: 32) aligned(lhs_data: 32) aligned(rhs_data: 32) collapse(3)
     for (size_t b = 0; b < batch_size; ++b) {
         for (size_t i = 0; i < rows; ++i) {
             for (size_t j = 0; j < cols; ++j) {
-                const T* lhs_row = lhs.data + b * rows * inner_dim + i * inner_dim;
-                const T* rhs_col = rhs.data + b * cols * inner_dim + j * inner_dim;
+                const T* lhs_row = lhs_data + b * rows * inner_dim + i * inner_dim;
+                const T* rhs_col = rhs_data + b * cols * inner_dim + j * inner_dim;
                 T* result_row = result_data + b * rows * cols + i * cols;
                 T sum = T(); // Initialize sum as zero for type T
 
@@ -61,95 +64,165 @@ Tensor<T> batch_matrix_matmul_transpose(const Tensor<T>& lhs, const Tensor<T>& r
 }
 
 
-// template <>
-// Tensor<float> batch_matrix_matmul_transpose<float>(const Tensor<float>& lhs, const Tensor<float>& rhs) {
-//     // Check dimensions
-//     if (lhs.ndim != 3 || rhs.ndim != 3) {
-//         throw std::invalid_argument("Both tensors must have 3 dimensions for batch matrix multiplication.");
-//     }
-//     if (lhs.shape[0] != rhs.shape[0]) {
-//         throw std::invalid_argument("Batch sizes must match for batch matrix multiplication.");
-//     }
-//     if (lhs.shape[2] != rhs.shape[2]) {
[…]

### 2024-12-28 Exam Assignments/Code/einspeed/include/compute_einsum.h

@@ -1,8 +1,9 @@
 #ifndef COMPUTE_EINSUM_H
 #define COMPUTE_EINSUM_H
 
-// uncomment the following line to show debug information
+// uncomment the following lines to show debug information
 // #define DEBUG
+// #define DEBUG_VERBOSE
 
 #include <unordered_map>
 #include <unordered_set>
@@ -54,17 +55,17 @@ PyObject* compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
     //
     // bmm_result[batch, kept left, kept right]
     //
-    // To transform it back to the desired shape specified by target_string, we need to track the column identifiers.
+    // To transform it back to the desired shape specified by target_string, we need to reshape and transpose the tensor based on the column specifiers in target_string.
 
+    // Let's start by parsing the format string
     std::string lhs_string, rhs_string, target_string;
     is_valid_einsum_expression(s, lhs_tensor, rhs_tensor, lhs_string, rhs_string, target_string); // throws error if not
 
-    
     // Vectors to store categorized index types
     std::vector<size_t> batch_dims_lhs, kept_left_dims_lhs, contracted_dims_lhs, summed_left_dims_lhs;
     std::vector<size_t> batch_dims_rhs, kept_right_dims_rhs, contracted_dims_rhs, summed_right_dims_rhs;
 
-    // Split lhs_string and rhs_string by identifying the roles of each index (batch, contracted, etc.)
+    // Identifying the roles of each column identifier (batch, contracted, etc.)
     // Staring with lhs_string
     for (size_t i = 0; i < lhs_string.length(); i++) {
         char lhs_char = lhs_string[i];
@@ -98,7 +99,7 @@ PyObject* compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
     Tensor<T> transposed_lhs = lhs_tensor.transpose(merge_vectors({batch_dims_lhs, kept_left_dims_lhs, contracted_dims_lhs, summed_left_dims_lhs}));
     Tensor<T> transposed_rhs = rhs_tensor.transpose(merge_vectors({batch_dims_rhs, kept_right_dims_rhs, contracted_dims_rhs, summed_right_dims_rhs}));
 
-#ifdef DEBUG
+#ifdef DEBUG_VERBOSE
     std::cout << "Transposed Tensors:" << '\n';
     transposed_lhs.print();
     transposed_rhs.print();
[…]

### 2024-12-28 Exam Assignments/Code/einspeed/include/tensor.h

@@ -158,8 +158,6 @@ public:
         return numpy_array;
     }
 
-    
-
 
     // Reshape method
     Tensor<T>& reshape(const std::vector<size_t>& new_shape) {
@@ -185,17 +183,19 @@ public:
     Tensor<T> reduce() const {
         // reduces over the last dimension
         const size_t last_dimension = this->shape[this->ndim-1];
-        const size_t target_size = this->calculate_size(this->shape, this->ndim) / last_dimension;
-        T* target_data = new T[target_size];
+        const size_t target_size = Tensor<T>::calculate_size(this->shape, this->ndim - 1);
+        T* target_data = nullptr;
+        posix_memalign((void**)&target_data, 32, target_size * sizeof(T));
+        T* data = this->data;
 
-        #pragma omp parallel for
+        #pragma omp parallel for simd aligned(data: 32) aligned(target_data: 32)
         for (size_t i = 0; i < target_size; i++) {
             T sum = 0;
             size_t start_idx = i * last_dimension;
             size_t end_idx = start_idx + last_dimension;
 
             for (size_t k = start_idx; k < end_idx; k++) {
-                sum += this->data[k];
+                sum += data[k];
             }
             
             target_data[i] = sum;
@@ -209,10 +209,6 @@ public:
 
     // Utility Methods
 
-    bool inline is_scalar() const {
-        return ndim == 0 && data;
-    }
-    
     static size_t calculate_size(const size_t* shape, size_t ndim) {
         size_t size = 1;
         for (size_t i = 0; i < ndim; ++i) {
@@ -314,7 +310,8 @@ Tensor<T> Tensor<T>::transpose(const std::vector<size_t>& perm) const {
         new_shape[i] = shape[perm[i]];
     }
 
-    T* transposed = new T[Tensor<T>::calculate_size(shape, ndim)];
+    T* transposed = nullptr;
+    posix_memalign((void**)&transposed, 32, Tensor<T>::calculate_size(shape, ndim) * sizeof(T));
     int* tmp_size = cast_all<size_t, int>(ndim, shape);
     int* tmp_perm = cast_all<size_t, int>(ndim, perm.data());
     // create a plan (shared_ptr)
@@ -327,7 +324,7 @@ Tensor<T> Tensor<T>::transpose(const std::vector<size_t>& perm) const {
     delete[] tmp_size;
     delete[] tmp_perm;
     Tensor<T> result(ndim, new_shape, transposed);
[…]

### 2024-12-28 Exam Assignments/paper/paper.tex
refining paper

@@
 \usepackage{etoolbox}
+
+% \widowpenalty=10000  % Prevents widows (single lines at the top of a page)
+% \clubpenalty=10000   % Prevents orphans (single lines at the bottom of a page)
 
@@
 % \item \textbf{Topic and background:} What topic does the paper deal with? What is the point of departure for your research? Why are you studying this now?
-\item \textbf{Topic and background:} This paper concerns tensor operations, specifically those implemented in the Python library numpy with its \texttt{numpy.einsum} method.
+\item \textbf{Topic and background:} This paper concerns tensor operations, specifically those implemented in the Python library NumPy with its \texttt{numpy.einsum} method.
 Tensor operations are very common in todays data driven world, especially with the rise of \em Deep Learning\em .
@@
 This paper tests all of these operations, as well as their combinations, and compares it to a hardcoded implementation written in C++.
-\item \textbf{Method:} To this end, I implemented a bare einsum implementation which implements these three operatios in C++, where one external library for transposing tensors is employed.
+\item \textbf{Method:} To this end, I implemented a bare einsum implementation which implements these three operations in C++, where one external library for transposing tensors is employed.
 Then, both implementations are compared, and the findings are presented in this paper.
 \item \textbf{Key findings:} The library einsum implementation does a really good job at transposing tensors.
-However, when contraction and especially reduction take up much of the workload, numpy's einsum implementation is remarkably ineffecient, as my C++ implementation executes faster by magnitudes.
+However, when contraction and especially reduction take up much of the workload, NumPy's einsum implementation is remarkably inefficient, as my C++ implementation executes faster by magnitudes.
 % \item \textbf{Conclusions or implications:} What do these findings mean? What broader issues do they speak to?
-\item \textbf{Conclusions or implications:} These findings emphasize that we cannot trust every implementation of a library to be implemented efficiently, even when speaking of popular libraries like numpy.
-In this specfic case of einsum, I advise the reader not to use it for big data operations which rely heavily on contraction or reduction.
+\item \textbf{Conclusions or implications:} These findings emphasize that we cannot trust every implementation of a library to be implemented efficiently, even when speaking of popular libraries like NumPy.
+In this specific case of einsum, I advise the reader not to use it for big data operations which rely heavily on contraction or reduction.
 \end{enumerate}
@@
 
-\let\thefootnote\relax\footnotetext{AEPRO 2025, March 1, Jena, Germany.
+% \let\thefootnote\relax\footnotetext{AEPRO 2025, March 1, Jena, Germany.
 Copyright \copyright 2025 for this paper by its authors.
@@
 For example \texttt{ik,kj->ij} specifies a normal matrix-matrix-multiplication.
-Here is how these \em column identifiers \em of \texttt{lhs\_string} in the eisnum expression \texttt{<lhs\_string>,<rhs\_string>-><target\_string>} are interpreted:
+Here is how these \em column identifiers \em of \texttt{lhs\_string} in the einsum expression \texttt{<lhs\_string>,<rhs\_string>-><target\_string>} are interpreted:
 
@@
 \subsection{My Contributions}
-In order to have a complete einsum implementation, many steps are necessary, like parsing the input string, transposing and reshaping the operand tensors, reducing them, batch matrix multiply them, and reshape and retranspose the resulting matrix again.
+In order to have a complete einsum implementation, many steps are necessary, like parsing the input string, transposing and reshaping the operand tensors, reducing them, batch matrix multiply them, and reshape and transpose the resulting matrix again.
 The details of all of these steps can be found in section \ref{sec:algorithm}.
@@
 The first step is to categorize each dimension of the two input tensors into one of the four categories \em batch, kept left, contracted \em or \em summed left \em in case of the left hand side tensor, similarly for the right hand side.
-We do this by simply looping over all so called \em column identifiers \em specified by the format string (columnn identifiers are basically just the individual characters in the string) and checking if this identifier is also present in the \texttt{rhs\_string} and \texttt{target\_string}.
+We do this by simply looping over all so called \em column identifiers \em specified by the format string (column identifiers are basically just the individual characters in the string) and checking if this identifier is also present in the \texttt{rhs\_string} and \texttt{target\_string}.
 Based on these checks we assign the categories, see table~\ref{tab:column_identifiers}.
@@
 Now that we have categorized all the dimensions, we bring both tensors into the canonical form
-\[ \text{lhs\_tensor}[\vec{b}, \vec{k_l}, \vec{c}, \vec{s_l}] \ , \ \text{rhs\_tensor}[\vec{b}, \vec{k_r}, \vec{c}, \vec{s_r}] \quad . \]
+% \[ \text{lhs\_tensor}[\vec{b}, \vec{k_l}, \vec{c}, \vec{s_l}] \ , \ \text{rhs\_tensor}[\vec{b}, \vec{k_r}, \vec{c}, \vec{s_r}] \quad . \]
+\[ \text{lhs\_tensor}[{b}, {k_l}, {c}, {s_l}] \ , \ \text{rhs\_tensor}[{b}, {k_r}, {c}, {s_r}] \quad . \]
 One very important detail is that the order of the \em batch \em and \em contracted \em dimensions after transposing have the same ordering in the operand tensors.
@@
 \subsubsection{Reducing Operand Tensors}
-Now that we have both operand tensors in the desired shape, we have a fix plan on how to operate on these tensors, which means the following precedure is identical to all possible inputs.
+Now that we have both operand tensors in the desired shape, we have a fix plan on how to operate on these tensors, which means the following procedure is identical to all possible inputs.
 Only the reshaping and transposing of the result matrix at the very end needs specific treatment.
@@
 After that, the two tensors will have the following form
-\[ \text{lhs\_tensor}[\vec{b}, \vec{k_l}, \vec{c}] \ , \ \text{rhs\_tensor}[\vec{b}, \vec{k_r}, \vec{c}] \quad . \]
+% \[ \text{lhs\_tensor}[\vec{b}, \vec{k_l}, \vec{c}] \ , \ \text{rhs\_tensor}[\vec{b}, \vec{k_r}, \vec{c}] \quad . \]
+\[ \text{lhs\_tensor}[{b}, {k_l}, {c}] \ , \ \text{rhs\_tensor}[{b}, {k_r}, {c}] \quad . \]
 It is worth mentioning that it might be possible to combine all of these operations so far (transposing, reshaping, reducing fourth dimension) into one big one.
@@
 Consequently, this ensures excellent spatial locality and hence we make perfect use of the cache.
-The result of this operation will be the three dimensiol tensor \texttt{bmm\_result}, with dimension sizes
+The result of this operation will be the three dimensional tensor \texttt{bmm\_result}, with dimension sizes
 \[ \text{bmm\_result} = [b, k_l, k_r] \quad . \]
@@
 To this end, we have to firstly split the current three dimensions of our result tensor into the individual ones.
-Hence, we we look at our inital tensors and their column identifiers again to get the dimension sizes.
+Hence, we we look at our initial tensors and their column identifiers again to get the dimension sizes.
 
-For example, so split the batch dimension, we check the the batch dimensions of \texttt{lhs\_tensor} and store their specific dimension sizes in the same order as they are present in the tensor.
+For example, to split the batch dimension, we check the the batch dimensions of \texttt{lhs\_tensor} and store their specific dimension sizes in the same order as they are present in the tensor.
 We do this, as our \texttt{bmm\_result} tensor has the same ordering of its data with respect to the batch dimension as \texttt{lhs\_tensor}.
@@
 \end{lstlisting}
+\\\\
 You will likely see that the strides got altered simply.
@@
 \subsection{Transposition Method}
-Let us now take a closer look at the smaller "puzzle pieces" making up the main \texttt{comput\_einsum} function, starting with the tensor transposition method.
+Let us now take a closer look at the smaller "puzzle pieces" making up the main \texttt{compute\_einsum} function, starting with the tensor transposition method.
 As already discussed, I chose to use an external library for this purpose, and hence there is not much to analyze, as we just pass the control flow to the library.
@@
 Hence the constant specifier in the method declaration.
-% TODO: write about how efficient it is, vectoriziation, multi threading etc.
+% TODO: write about how efficient it is, vectorization, multi threading etc.
 
 \subsection{Reshaping Method}
-The implementation of this method is very sraight forward.
+The implementation of this method is fairly simple.
 As discussed, when calling this method, we ensure that the tensor has canonical strides, as in this case we only have to change the meta attributes \texttt{ndim} and \texttt{shapes}.
@@
 This is the first method we look at which has an impact on performance, thus we should be cautious with implementation choices.
-Hence, we write the folowing compiler-optimization-friendly code:
+Hence, we write the following compiler-optimization-friendly code:
 
@@
 \end{lstlisting}
-
+\raggedbottom
 \subsection{Batch Matrix Multiplication}
@@
 Hence, integer value are actually not supported.
-The only supported data types are \texttt{np.float64} and \texttt{np.complex128}.
+The only supported data types are \texttt{np.float32}, \texttt{np.float64}, \texttt{np.complex64}, and \texttt{np.complex128}.
 
@@
   \item More optimized sub methods, improving parallelization, vectorization by checking for aligned data, etc.
-  \item Utilize gpu for parallelization.
+  \item Utilize GPU for parallelization.
 \end{itemize}
@@
 \\\\
-Note that per constraction no entry of \texttt{correct} is zero.
+Note that per construction no entry of \texttt{correct} is zero.
 It turns out that this relative discrepancy stays below a value of approximately \texttt{1e-4}, which is very commendable.
@@
 
-Ultimately, however, neither implementanion is exactly correct, and my implementation isn't necessarily worse.
+Ultimately, however, neither implementation is exactly correct, and my implementation isn't necessarily worse.
 In fact, on some test cases it was closer to a hard coded answer, but this doesn't have to mean anything, as these computations will suffer from numerical errors as well.
@@
 \label{sec:conclusions}
-In this paper, I implemented the einsum function from scratch, in order to evaluate the existing NumPy implementanion.
+In this paper, I implemented the einsum function from scratch, in order to evaluate the existing NumPy implementation.
 Despite the fact that I certainly could have done better, I was pretty satisfied with my work at the end.

### 2024-12-28 Exam Assignments/Code/einspeed/include/numpy_types.h

@@ -11,6 +11,7 @@ constexpr int numpy_type_of() {throw std::runtime_error("Unsupported data type f
 template<> constexpr int numpy_type_of<long>() {return NPY_INT64;}
 template<> constexpr int numpy_type_of<float>() {return NPY_FLOAT32;}
 template<> constexpr int numpy_type_of<double>() {return NPY_FLOAT64;}
+template<> constexpr int numpy_type_of<std::complex<float>>() {return NPY_COMPLEX64;}
 template<> constexpr int numpy_type_of<std::complex<double>>() {return NPY_COMPLEX128;}
 
 
@@ -31,6 +32,11 @@ struct numpy_to_cpp_type<NPY_FLOAT32> {
 template <>
 struct numpy_to_cpp_type<NPY_FLOAT64> {
     using type = double;
+};
+
+template<>
+struct numpy_to_cpph_type<NPY_COMPLEX64> {
+    using type = std::complex<float>;
 };
 
 template <>
@@ -59,6 +65,7 @@ constexpr const int SUPPORTED_NUMPY_TYPES[] = {
     // NPY_INT64, // no int64, as hptt does not support it
     NPY_FLOAT32,
     NPY_FLOAT64,
+    NPY_COMPLEX64,
     NPY_COMPLEX128
 };
 

### 2024-12-28 Exam Assignments/Code/einspeed/include/numpy_types.h
tensor unit tests

@@ -35,7 +35,7 @@ struct numpy_to_cpp_type<NPY_FLOAT64> {
 };
 
 template<>
-struct numpy_to_cpph_type<NPY_COMPLEX64> {
+struct numpy_to_cpp_type<NPY_COMPLEX64> {
     using type = std::complex<float>;
 };
 

### 2024-12-28 Exam Assignments/Code/einspeed/include/tensor.h

@@ -24,7 +24,7 @@ public:
     T* data = nullptr;        // Pointer to the actual data
                               // IMPORTANT: Tensor class does NOT free the data!
 
-    Tensor(size_t ndim, size_t* shape, T* data)
+    Tensor(size_t ndim, const size_t* shape, T* data)
         : ndim(ndim), data(data) {
             this->shape = new size_t[ndim];
             std::memcpy(this->shape, shape, ndim * sizeof(size_t));
@@ -235,6 +235,10 @@ public:
 
         // Print data in a formatted way
         if (ndim) Tensor<T>::print_data_recursive(data, shape, ndim, 0);
+    }
+
+    std::vector<size_t> get_shape() const {
+        return std::vector<size_t>(shape, shape + ndim);
     }
 
 private:

### 2024-12-28 Exam Assignments/Code/einspeed/tests/catch_tests_tensor.cpp

@@ -1,0 +1,118 @@
+#include "catch.hpp"
+#include "tensor.h"
+
+// Helper function to create a Tensor object with random data for testing
+template <typename T>
+Tensor<T> create_tensor(const std::vector<size_t>& shape) {
+    size_t size = 1;
+    for (auto s : shape) size *= s;
+    T* data = nullptr;
+    posix_memalign((void**)&data, 32, size * sizeof(T));
+
+    // Fill tensor with some test data (e.g., sequential integers)
+    for (size_t i = 0; i < size; ++i) {
+        data[i] = static_cast<T>(i + 1);
+    }
+
+    return Tensor<T>(shape.size(), shape.data(), data);
+}
+
+TEST_CASE("Tensor Transpose", "[transpose]") {
+    SECTION("Transpose 2D Tensor") {
+        std::vector<size_t> shape = {2, 3};
+        Tensor<float> tensor = create_tensor<float>(shape);
+        std::vector<size_t> perm = {1, 0};  // Swap rows and columns
+        Tensor<float> transposed = tensor.transpose(perm);
+
+        REQUIRE(transposed.get_shape() == std::vector<size_t>{3, 2});
+        REQUIRE(transposed.data[0] == tensor.data[0]);
+        REQUIRE(transposed.data[1] == tensor.data[3]);
+        REQUIRE(transposed.data[2] == tensor.data[1]);
+        REQUIRE(transposed.data[3] == tensor.data[4]);
+        REQUIRE(transposed.data[4] == tensor.data[2]);
+        REQUIRE(transposed.data[5] == tensor.data[5]);
+    }
+
+    SECTION("Transpose 3D Tensor") {
+        std::vector<size_t> shape = {2, 3, 4};
+        Tensor<float> tensor = create_tensor<float>(shape);
+        std::vector<size_t> perm = {2, 0, 1};  // Move third dim to first place
+        Tensor<float> transposed = tensor.transpose(perm);
+
+        REQUIRE(transposed.get_shape() == std::vector<size_t>{4, 2, 3});
+    }
+}
+
+TEST_CASE("Tensor Reshape", "[reshape]") {
+    SECTION("Reshape 1D to 2D Tensor") {
+        std::vector<size_t> shape = {6};
+        Tensor<float> tensor = create_tensor<float>(shape);
+        std::vector<size_t> new_shape = {2, 3};  // Reshape into a 2x3 matrix
[…]

### 2024-12-30 Exam Assignments/Chapters/10/10.tex
exam assignments week 10

\documentclass[../../main.tex]{subfiles}

\begin{document}
\section{Debugging and Profiling}
\subsection{Donald Knuth on Premature Optimization}
Donald Knuth states: "We should forget about small efficiencies, say about 97\% of the time: premature optimization is the root of all evil."
This quote is a reminder that we should not focus on optimizing code before we have a clear understanding of the problem we are trying to solve. This is because premature optimization can lead to code that is difficult to read, maintain, and debug. It can also lead to code that is less flexible and harder to change in the future. Instead, we should focus on writing clear, readable, and maintainable code that solves the problem at hand. Once we have a clear understanding of the problem and have written code that works, we can then focus on optimizing it if necessary. I agree with this quote because I have seen firsthand how premature optimization can lead to code that is difficult to work with. By focusing on writing clear, readable, and maintainable code first, we can avoid many of the pitfalls of premature optimization and create code that is easier to work with in the long run. Optimization should only be done later when necessary and should be based on data and profiling, not on assumptions or premature guesses.

\subsection{Debugging With cout Is a Valid Alternative}
Let us analyze slide 7, which states debuggers are not always preferred for debugging by some programmers, including famous ones like Linus Torvalds etc. On the slide there is a link to a webpage, which features a discussion about debuggers. The main statement is that debuggers have their place, but they are not always the best tool for the job, especially when dealing with scalable systems. Also, the author mentions that debuggers do find mistakes in the code and help to make quick fixes, but in order to find the "real" fixes, that make the program flow more organized and well structured, one has to deeply understand the code one is working on and see the bigger picture.

I agree with the statement that debuggers should not be the preferred tool for debugging generally. Let me phrase it like this: If there aren't many alternatives to using a debugger, the real mistake (metaphorically the \emph{bug}) was made beforehand. Bad structuring and not writing tests or checks for parameters will lead to complex control flows. Hence, not using debuggers will enforce the programmer to take preventive measures, such as writing cleaner code, adding proper tests, and ensuring good code structure. This approach can lead to more maintainable and understandable code in the long run. Consequently, the resulting code will be more scalable and easier to debug, which ultimately leads to a more efficient development process. This is why I prefer this approach over using debuggers.

\end{document}

### 2024-12-30 Exam Assignments/Code/einspeed/einspeed.py

@@ -65,21 +65,21 @@ def run_test_cases():
         # There will be a big discrepancy for np.float32, but einspeed's result is actually CLOSER to the correct result!
 
         # Reduction
-        # benchmark_einsum("abc,def->", [50, 42, 55], [63, 69, 127], dtype, "Multidimensional Reduction")
+        benchmark_einsum("abc,def->", [50, 42, 55], [63, 69, 127], dtype, "Multidimensional Reduction")
 
-        # # Contraction
-        # benchmark_einsum("i,i->", [600000000], [600000000], dtype, "Vector Dot Product")
-        # benchmark_einsum("ij,ij->", [9000, 80000], [9000, 80000], dtype, "Element-wise Contraction")
-        # benchmark_einsum("ij,ji->", [9000, 80000], [80000, 9000], dtype, "Element-wise Contraction with Transpose")
-        # benchmark_einsum("ijk,ijk->", [1000, 2000, 300], [1000, 2000, 300], dtype, "High-Dimensional Contraction")
+        # Contraction
+        benchmark_einsum("i,i->", [600000000], [600000000], dtype, "Vector Dot Product")
+        benchmark_einsum("ij,ij->", [9000, 80000], [9000, 80000], dtype, "Element-wise Contraction")
+        benchmark_einsum("ij,ji->", [9000, 80000], [80000, 9000], dtype, "Element-wise Contraction with Transpose")
+        benchmark_einsum("ijk,ijk->", [1000, 2000, 300], [1000, 2000, 300], dtype, "High-Dimensional Contraction")
 
-        # # Outer product
-        # benchmark_einsum("i,j->ij", [100000], [20000], dtype, "Outer Product")
-        # benchmark_einsum("ij,k->ijk", [300, 2000], [500], dtype, "Outer Product 3D")
+        # Outer product
+        benchmark_einsum("i,j->ij", [100000], [20000], dtype, "Outer Product")
+        benchmark_einsum("ij,k->ijk", [300, 2000], [500], dtype, "Outer Product 3D")
         
         # Transpose-like operations
         benchmark_einsum("ij,jk->ik", [300, 400], [400, 500], dtype, "Matrix Multiplication")
-        benchmark_einsum("ijk,jl->ikjl", [100, 300, 300], [300, 400], dtype, "Tensor Contraction")
[…]

### 2024-12-30 Exam Assignments/Code/einspeed/include/blas.h

@@ -117,8 +117,8 @@ Tensor<float> batch_matrix_matmul_transpose<float>(const Tensor<float>& lhs, con
                 size_t k = 0;
                 for (; k + 7 < inner_dim; k += 8) { // Process 8 elements at a time
                     // Load 8 floats from lhs_row and rhs_col
-                    __m256 vec_lhs = _mm256_load_ps(lhs_row + k);
-                    __m256 vec_rhs = _mm256_load_ps(rhs_col + k);
+                    __m256 vec_lhs = _mm256_loadu_ps(lhs_row + k);
+                    __m256 vec_rhs = _mm256_loadu_ps(rhs_col + k);
 
                     // Multiply and accumulate
                     __m256 vec_product = _mm256_mul_ps(vec_lhs, vec_rhs);
@@ -198,8 +198,8 @@ Tensor<double> batch_matrix_matmul_transpose<double>(const Tensor<double>& lhs, 
                 size_t k = 0;
                 for (; k + 3 < inner_dim; k += 4) { // Process 4 elements at a time
                     // Load 4 floats from lhs_row and rhs_col
-                    __m256d vec_lhs = _mm256_load_pd(lhs_row + k);
-                    __m256d vec_rhs = _mm256_load_pd(rhs_col + k);
+                    __m256d vec_lhs = _mm256_loadu_pd(lhs_row + k);
+                    __m256d vec_rhs = _mm256_loadu_pd(rhs_col + k);
 
                     // Multiply and accumulate
                     __m256d vec_product = _mm256_mul_pd(vec_lhs, vec_rhs);

### 2024-12-30 Exam Assignments/Code/einspeed/include/compute_einsum (copy).h

@@ -1,0 +1,321 @@
+#ifndef COMPUTE_EINSUM_H
+#define COMPUTE_EINSUM_H
+
+// uncomment the following lines to show debug information
+// #define DEBUG
+// #define DEBUG_VERBOSE
+
+#include <unordered_map>
+#include <unordered_set>
+#include <stdexcept>
+#include <string>
+
+#include "tensor.h"
+#include "blas.h"
+#include "func.h"
+
+
+// **************************************************
+// *** The core functionality is implemented HERE ***
+// **************************************************
+
+
+// Some declarations (implementation can be found below)
+std::vector<size_t> column_identifiers_like(const std::vector<size_t>& target_column_identifiers, const std::string& target_identifier_string, const std::string& origin_identifier_string);
+template <typename T> bool is_valid_einsum_expression(const char* const s, const Tensor<T>& lhs_tensor, const Tensor<T>& rhs_tensor, std::string& lhs_string, std::string& rhs_string, std::string& target_string);
+
+
+template <typename T>
+PyObject* compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const Tensor<T>& rhs_tensor) {
+
+    // *************************
+    // *** Function Overview ***
+    // *************************
+
+    // We categorize each column identifier c in string lhs_string into one of the following four categories:
+    //
+    // present in   |  rhs_string  | target_string |
+    // ---------------------------------------------
+    // batch        |      yes     |      yes      |
+    // kept left    |      no      |      yes      |
+    // contracted   |      yes     |      no       |
+    // summed left  |      no      |      no       |
+
+    // We do the same for rhs_string. Then, we transpose and reshape lhs_tensor and rhs_tensor in the following manner:
+    //
+    // lhs_tensor[batch, kept left, contracted, summed left]
+    // rhs_tensor[batch, kept right, contracted, summed right]
+    //
+    // IMPORTANT: the column identifiers of batch and contracted of both tensors are obviously the same.
+    //            We have to ensure that they also have the same dimensions and ordering!
+
+    // Now, we are ready to perform an axis summation on lhs_tensor over the last dimension (= summed left). The same for rhs_tensor.
[…]

### 2024-12-30 Exam Assignments/Code/einspeed/include/compute_einsum.h

@@ -2,28 +2,26 @@
 #define COMPUTE_EINSUM_H
 
 // uncomment the following lines to show debug information
-// #define DEBUG
+#define DEBUG
 // #define DEBUG_VERBOSE
 
 #include <unordered_map>
 #include <unordered_set>
 #include <stdexcept>
 #include <string>
+#include <omp.h> // Include OpenMP for timing
 
 #include "tensor.h"
 #include "blas.h"
 #include "func.h"
 
-
 // **************************************************
 // *** The core functionality is implemented HERE ***
 // **************************************************
 
-
 // Some declarations (implementation can be found below)
 std::vector<size_t> column_identifiers_like(const std::vector<size_t>& target_column_identifiers, const std::string& target_identifier_string, const std::string& origin_identifier_string);
 template <typename T> bool is_valid_einsum_expression(const char* const s, const Tensor<T>& lhs_tensor, const Tensor<T>& rhs_tensor, std::string& lhs_string, std::string& rhs_string, std::string& target_string);
-
 
 template <typename T>
 PyObject* compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const Tensor<T>& rhs_tensor) {
@@ -96,8 +94,23 @@ PyObject* compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const
     // Reshaping the lhs and rhs tensors as described:
     // lhs_tensor[batch, kept left, contracted, summed left]
     // rhs_tensor[batch, kept right, contracted, summed right]
+#ifdef DEBUG
+    double start_time_transpose_lhs = omp_get_wtime();
+#endif
     Tensor<T> transposed_lhs = lhs_tensor.transpose(merge_vectors({batch_dims_lhs, kept_left_dims_lhs, contracted_dims_lhs, summed_left_dims_lhs}));
+#ifdef DEBUG
+    double end_time_transpose_lhs = omp_get_wtime();
+    std::cout << "Time taken to transpose LHS: " << (end_time_transpose_lhs - start_time_transpose_lhs) << " seconds\n";
+#endif
+
+#ifdef DEBUG
+    double start_time_transpose_rhs = omp_get_wtime();
+#endif
     Tensor<T> transposed_rhs = rhs_tensor.transpose(merge_vectors({batch_dims_rhs, kept_right_dims_rhs, contracted_dims_rhs, summed_right_dims_rhs}));
+#ifdef DEBUG
[…]

### 2025-01-07 Exam Assignments/Chapters/11/11.tex
week 11

\documentclass[../../main.tex]{subfiles}

\begin{document}
\section{Designing SSD-Friendly Applications}
\subsection{Adopt Compact Data Structures}
In this lecture we analyzed how exactly SSDs function, and deduced from that knowledge how we should write our programs in order to fully employ its capacity. On slide 13, we conclude the first paradigms for SSD friendly code:

~\\
First, we should keep in mind the page size of the SSD, which typically is 4KB. Hence, small read and write operation below this mark will have the same cost of a 4KB one, and in case of a write operation we may not fully utilized the space as the page we are writing to flash itself is not fully utilized. Thus, we should combine smaller IO operations in two fewer bigger ones, provided that these operations are sequential, which directly leads us to the second conclusion:

~\\
We should prefer compact data structures over scattered ones. This is due to the fact that we utilized the pages better and the controller of the SSD can better optimize this. Practically speaking, it is advisable to use fewer files (maybe even only one) to store our data, and in this file we should group related data into sequential sections.

~\\
Following these advises will result in more efficient IO operations and will also reduce the wear of the SSD and hence contribute to its longevity.

~\\
\subsection{Designing SSD-friendly Applications for Better Application Performance and Higher IO Efficiency}
This paper introduces the reader to the core functioning of SSDs and presents different optimization techniques to improve performance or longevity of the SSD. One optimization presented in the paper directly follows on from the previous discussion. The other one I chose to present analyzes the page distribution inside the SSD and its effect on performance.

\begin{description}
    \item[Seperate hot data from cold data] We have already established that we should group related data. But additionally, or rather more importantly, we should group hot data sequentially (i.e. data accessed very often) as well as cold data. The optimum for all data would be a smooth transition between these two extrema. The reason for this is very similar as before, as when accessing data from the SSD we also access unwanted data due the teh page size. Optimally, we also require this additional data, otherwise this data section is accessed redundantly. To maximize the chance of the first scenario, we organize our data in the proposed manner.
    \item[Avoid full SSD usage] When we assume that the used pages distributed equally among the entire SSD, and the SSD is occupied by a factor of $A$ ($A = 0.5 \implies$ half the pages are in use), then for each block on average we have $\frac{no\_occupied\_pages\_in\_block}{no\_pages\_in\_block} = A$. Thus, we see that the garbage collector will have to do way less data movement when $A$ is low. In fact, $\frac{1}{1-A}$ blocks will need to be considered when trying to free a block on average, and also note that for each of these blocks, $A \cdot no\_pages\_in\_block$ pages will need to be copied. Thus, we face a double penalty for bigger values of $A$ ($\frac{A \cdot no\_pages\_in\_block}{1-A}$ will need to be moved; in the paper $P$ is used for $no\_pages\_in\_block$).
\end{description}
\end{document}

### 2025-01-14 Exam Assignments/Chapters/12/12.tex
week 12

\documentclass[../../main.tex]{subfiles}

\begin{document}
\section{Advanced Topics}
\subsection{Cancellation Points in OpenMP}
On slide 6, a new concept called \em cancellation points \em were introduced. They are a new concept since OpenMP 4.0, and they allow the programmer to dynamically stop the execution of a parallel region based on a condition. This is useful when the parallel region is running for a long time and the programmer wants to stop it if a certain condition is met, like one solution (out of possible many or even infinite) is found. The syntax for cancellation points is as follows:

~\\
Use \texttt{\#pragma omp cancel for} to signal cancellation to all necessary threads. The \texttt{for} can be replaced with \texttt{parallel} or \texttt{sections} depending on the context. Now that we have signaled the cancellation, we also have to check for it in all the threads, which is done with \texttt{\#pragma omp cancellation point for}. If a cancellation is signaled, all threads will stop executing the corresponding parallel region and continue with the code after the parallel region.

~\\
The following example demonstrates the use of cancellation points in OpenMP:

\begin{figure}[h]
    \begin{lstlisting}
#pragma omp parallel // start parallel region
{
#pragma omp for schedule(dynamic)
    for (int i = 0; i < biggest_possible_number; ++i) {
        if (is_solution(i)) { // find some solution, not necessary the smallest
            final_solution = i;
#pragma omp cancel for // signal cancellation, because we found a solution
    }
#pragma omp cancellation point for // check for cancellations signalled from other threads
    // cancel for loop if cancellations signalled
    }
} // end parallel region
    \end{lstlisting}
    \caption{Example of using cancellation points in OpenMP (slide 6)}
\end{figure}

~\\
Now, the reader may ask why we need additional directives for this behavior, couldn't we just use a global variable and check it in the parallel region? Well, we definitely could do that, but the cancellation points are more efficient and easier to use. The OpenMP runtime system can check the cancellation condition at the end of the loop iterations, and if the condition is met, it will stop the execution of the parallel region. This is more efficient than checking the condition in every iteration of the loop, and it is also easier to use since the programmer doesn't have to write the code for checking the condition. Furthermore, imagine the scenario of a for loop with \em dynamic \em scheduling. If we used a global variable for the cancellation condition, we would only cancel the current batch of iterations the thread is working on, and potentially there are many more batches of iterations left which are not yet cancelled. Needless to say, this is not very efficient.

~\\
\subsection{CUDA}
CUDA is a parallel computing platform and application programming interface (API) model created by NVIDIA. It allows software developers to use a CUDA-enabled graphics processing unit (GPU) for general-purpose processing. The CUDA platform is used in many scientific applications, where high performance is required. It is also used in many machine learning applications, where large amounts of data need to be processed quickly. The CUDA platform provides a set of libraries and tools for developing parallel applications, and it is supported by many programming languages, including C, C++, and Fortran. Thus, let us lake a look at the basic structure of a CUDA program shown in figure~\ref{fig:cuda}.

~\\
The CUDA code in figure~\ref{fig:cuda} demonstrates the basic structure of a CUDA program that performs element-wise addition of two arrays. It begins with a kernel function, \texttt{add}, which defines the computation executed on the GPU. The kernel uses a unique thread index, calculated from \texttt{blockIdx.x} and \texttt{threadIdx.x}, to identify the array element each thread should process. In the \texttt{main} function, memory is allocated on both the host (CPU) and the device (GPU), and data is initialized and transferred to the GPU. The kernel is launched using a grid of blocks, with each block containing multiple threads, enabling parallel processing of the arrays. After the kernel execution, the results are copied back from the device to the host for verification. Finally, all allocated memory is freed. This structure illustrates key concepts of CUDA programming, including thread management, memory allocation, data transfer, and kernel execution.

\begin{figure}[h]
\vspace{-5em}
\begin{lstlisting}[language=C, basicstyle=\ttfamily\small]
// Kernel function to add elements of two arrays
__global__ void add(int *a, int *b, int *c, int n) {
    int index = blockIdx.x * blockDim.x + threadIdx.x;
    if (index < n) {
        c[index] = a[index] + b[index];
    }
}

int main() {
    int n = 10240; // Array size
    int size = n * sizeof(int);

    // Allocate memory on the host (CPU)
    int *h_a = (int *)malloc(size);
    int *h_b = (int *)malloc(size);
    int *h_c = (int *)malloc(size);

    // Initialize host arrays
    for (int i = 0; i < n; i++) {
        h_a[i] = i;
        h_b[i] = i * 2;
    }

    // Allocate memory on the device (GPU)
    int *d_a, *d_b, *d_c;
    cudaMalloc((void **)&d_a, size);
    cudaMalloc((void **)&d_b, size);
    cudaMalloc((void **)&d_c, size);

    // Copy data from host to device
    cudaMemcpy(d_a, h_a, size, cudaMemcpyHostToDevice);
    cudaMemcpy(d_b, h_b, size, cudaMemcpyHostToDevice);

    // Define the number of threads per block and blocks per grid
    int threadsPerBlock = 256;
    int blocksPerGrid = (n + threadsPerBlock - 1) / threadsPerBlock;

    // Launch kernel
    add<<<blocksPerGrid, threadsPerBlock>>>(d_a, d_b, d_c, n);

    // Check for errors in kernel launch
    cudaError_t err = cudaGetLastError();
    if (err != cudaSuccess) {
        printf("CUDA Error: %s\n", cudaGetErrorString(err));
        return -1;
    }

    // Copy result back to host
    cudaMemcpy(h_c, d_c, size, cudaMemcpyDeviceToHost);

    for (int i = 0; i < n; i++) {
        printf("%d + %d = %d\n", h_a[i], h_b[i], h_c[i]);
    }
    // Free memory
    ...

    return 0;
}
\end{lstlisting}
    \caption{Basic CUDA program}
    \label{fig:cuda}
\end{figure}

~\\
\subsection{MPI}
MPI stands for Message Passing Interface, and it is a standard for parallel programming. It is used for distributed memory systems, where each processor has its own memory. The processors communicate with each other by sending messages. The MPI library is available in many programming languages like C, C++, and Fortran, and is used in many scientific applications, where high performance is required. To this end, MPI provides functions for sending and receiving messages, and for synchronizing the processors.

~\\
One of the fundamental features of MPI is its ability to send and receive messages between processors. The \texttt{MPI\_Send} function is used to send data from one processor to another, while \texttt{MPI\_Recv} is used to receive data. These functions ensure point-to-point communication, which is critical for many parallel algorithms. For example, \texttt{MPI\_Send} requires parameters such as the data buffer, count of elements, data type, destination rank, tag, and communicator. Similarly, \texttt{MPI\_Recv} specifies the source rank, among other parameters, to successfully retrieve messages. Figure~\ref{fig:mpi} demonstrates how one might use these functions.

\begin{figure}[h]
    \begin{lstlisting}
#include <iostream>
#include <assert.h>
#include <mpi.h>

#define MPI_Assert(error) assert(error == MPI_SUCCESS)

int main(int argc, char **argv) {
    MPI_Init(&argc, &argv);

    int l_rank, l_comm_size, l_data = 0;

    MPI_Comm_rank(MPI_COMM_WORLD, &l_rank);
    MPI_Comm_size(MPI_COMM_WORLD, &l_comm_size);

    MPI_Status status;

    // program logic
    if (l_rank) {
        MPI_Assert(MPI_Recv(&l_data, 1, MPI_INT, l_rank-1, l_rank-1, MPI_COMM_WORLD, &status));
        std::cout << "process " << l_rank << " received data " << l_data << " from process " << status.MPI_SOURCE << ".\n";
    }
    l_data += l_rank;
    std::cout << "process " << l_rank << " sends data " << l_data << " to process " << (l_rank+1) % l_comm_size << ".\n";
    MPI_Assert(MPI_Send(&l_data, 1, MPI_INT, (l_rank+1) % l_comm_size, l_rank, MPI_COMM_WORLD));
    if (l_rank == 0) {
        MPI_Assert(MPI_Recv(&l_data, 1, MPI_INT, l_comm_size-1, l_comm_size-1, MPI_COMM_WORLD, &status));
        std::cout << "process " << l_rank << " received data " << l_data << " from process " << status.MPI_SOURCE << ".\n";
    }
    MPI_Finalize();
}
    \end{lstlisting}
    \caption{Basic MPI program}
    \label{fig:mpi}
\end{figure}

~\\
In this little program every process sends one message to the next one based on their rank. The data that is send is a single integer which acts as an accumulator. The program starts with process 0, which sends the data to process 1. Process 1 receives the data, adds its rank to it, and sends it to process 2. This continues until the last process, which sends the data back to process 0, which receives it and prints the final result.

~\\
Beyond basic point-to-point communication, MPI also provides more advanced collective communication methods. One such function is \texttt{MPI\_Gather}, which allows a root processor to collect data from multiple processors and combine it into a single array. This is particularly useful for operations like gathering results from parallel computations into a central location for further analysis or visualization.

\end{document}

### 2025-03-04 Exam Assignments/Chapters/02/02.tex
Week 09
Refining all chapters

@@
     In the same way, false sharing occurs in computing when multiple threads modify different parts of the same cache line.
-Even though each thread may be working on separate variables, they're forced to constantly invalidate and reload the cache line they "share," leading to substantial inefficiencies.
+Even though each thread may be working on separate variables, they're forced to constantly invalidate and reload the cache line they "share", leading to substantial inefficiencies.
 
     ~\\
-    Here is an illustration of the sceario:
+    Here is an illustration of the scenario:
 

### 2025-03-04 Exam Assignments/Chapters/03/03.tex
Refining all chapters

@@
 ~\\
-In this code, a team of thread get created my the \em \#pragma omp parallel \em clause.
+In this code, a team of threads gets created by the \em \#pragma omp parallel \em clause.
 Their task is to print their respective thread numbers.
@@
 While still not defining the order of the tasks, they yet seem to execute in order.
-Interestingly, this is not quite true, as the main reason the seem to do so is because the \em terminate \em in order.
+Interestingly, this is not quite true, as the main reason they seem to do so is because their \em termination \em is in order.
 This is simply due to the fact that the work load for a thread with a higher thread number is substantially higher than for another one with a lower number.
@@
 
+~\\
 \subsubsection{Game of Life}
@@
 It calculates a random number quite efficiently.
-Another benfit is that we do not need multiple instances of random number generators, and we also don't run the risk of misusing those by calling their methods in different threads.
+Another benefit is that we do not need multiple instances of random number generators, and we also don't run the risk of misusing those by calling their methods in different threads.
 The code is shown in the \hyperref[fig:calculate_pi_v2]{figure} below.

### 2025-03-04 Exam Assignments/Chapters/04/04.tex
Refining all chapters

@@
 ~\\
-And here are the performance gains compared to std::sort():
+And here are the performance gains compared to std::sort(). Note the different coloring:
 
@@
 Let us now analyze the second two pictures, showing the running times at different array sizes.
-The graphs are as expected, the running times of the parallel algorithms is better, even more so for very large array sizes, as in those instances we can fully employ the parallel power of the cpu cores without much overhead relative to the running time, thus decreasing the running time significantly.
+The graphs are as expected, the running times of the parallel algorithms is better, even more so for very large array sizes, as in those instances we can fully employ the parallel power of the CPU cores without much overhead relative to the running time, thus decreasing the running time significantly.
 This is why the speedup gets better with increasing array size initially.
@@
 By declaring variables as volatile in C and C++, we make the compiler read and write the variable from and to memory every time we access it.
-Cache coherency protocols then ensure that every local cache of the cpu cores has the same value of the shared variable.
+Cache coherency protocols then ensure that every local cache of the CPU cores has the same value of the shared variable.
 So everything should be good then, right?

### 2025-03-04 Exam Assignments/Chapters/08/08.tex
Refining all chapters

@@
     \subsection{Analyzing slide 14}
-    Slide 14 explains how we can utilize vector instrinsics to transpose 8x8 matrices. Here is the code:
+    Slide 14 explains how we can utilize vector intrinsics to transpose 8x8 matrices. Here is the code:
     \begin{lstlisting}
@@
 Hence, after a final permutation we should be done calculating a column.
-To this end, we use \texttt{\_mm256\_permute2f128\_ps} and specify that we want to use the lower bits of both $a$ and $e$ by setting the mask to \texttt{CC}:
+To this end, we use \texttt{\_mm256\_permute2f128\_ps} and specify that we want to use the lower bits of both $a$ and $e$ by setting the mask to \texttt{0x20}:
 \begin{lstlisting}
@@
 If one writes dependent code, i.e.
-the next line of code may only be executed after the current one because of a dependency, the latency of the operation will be the main factor for performance, since we want our result as quickly as possible to procede with the execution.
+the next line of code may only be executed after the current one because of a dependency, the latency of the operation will be the main factor for performance, since we want our result as quickly as possible to proceed with the execution.
 
@@
 %  \\ ($no\_clock\_cycles \approx \frac{no\_operations}{throughput\_of\_operation}$ (Note that the number of operations does not directly correlate with the number of instructions in code, as certain code instructions need more cpu instructions)).
-In this case, the latency has a neglectable influence on the performance, especially if the number of operations grows bigger.
+In this case, the latency has a negligible influence on the performance, especially if the number of operations grows bigger.
 
@@
 The "key" idea is to simply load multiple consecutive keys into one vector register, and to compare them all against the search key.
-If one comparison is successful, we can calculate the index and return tha value at that index.
+If one comparison is successful, we can calculate the index and return the value at that index.
 Otherwise, we keep probing with the next set of keys.
@@
 
-    \item[Vectorized Fingerprinting] As mentioned, vectorized linear probing require integer keys.
+    \item[Vectorized Fingerprinting] As mentioned, vectorized linear probing requires integer keys.
 So what could we do if wanted to use strings instead?
-The "key" idea is to calculate "keys" for the keys themself (quite many keys around here).
+The "key" idea is to calculate "keys" for the keys themselves (quite many keys around here).
 One can think of these keys as hash values (though not necessarily the same as the real hashes) of the keys.

### 2025-03-04 Exam Assignments/Chapters/07/07.tex
Refining all chapters

@@
 \begin{description}
-    \item[aligned] ~\\ This clause tells the compiler that every element of an array $a$ is $n$-bit aligned if one uses \texttt{\#pragma opm simd aligned(a, n)}.
+    \item[aligned] ~\\ This clause tells the compiler that every element of an array $a$ is $n$-byte aligned if one uses \texttt{\#pragma opm simd aligned(a, n)}.
 The compiler employs this information to effectively vectorize the code section.
@@
 \subsubsection{AVX512 data types}
+\begin{figure}[h]
 \resizebox{\textwidth}{!}{ % Scale to the width of the text
@@
 }
+\caption{Illustration of AVX512 data types. \texttt{\_\_m512i} is not depicted precisely.}
+\end{figure}
 
@@
 \subsubsection{Intel MMX}
-In the paper \href{https://dl.acm.org/doi/pdf/10.1145/242857.242865}{Intel MMX for Multimedia PCs} the authors introduced the new MMX technology, which they integrated into their new cpus.
-It allows the cpu to employ vector instructions to increase throuphput and hence optimize running time of common computing tasks, such as rendering pixels for example.
+In the paper \href{https://dl.acm.org/doi/pdf/10.1145/242857.242865}{Intel MMX for Multimedia PCs} the authors introduced the new MMX technology, which they integrated into their new CPUs.
+It allows the CPU to employ vector instructions to increase throughput and hence optimize running time of common computing tasks, such as rendering pixels for example.
 They illustrated the use cases in several examples, two of them being:

### 2025-03-04 Exam Assignments/Chapters/09/09.tex
Refining all chapters

@@
 
-Now, we differentiate between two kinds of operations: Usual cpu instructions and memory operations.
+Now, we differentiate between two kinds of operations: Usual CPU instructions and memory operations.
 The reason for this is that memory operations are relatively much more time expensive, and the recent hardware trends make this gap grow bigger.
-Hence, we would like to know whether our program is bottlenecked by these expensive memory operations, which we very much would like to avoid, as we want to utilize our cpu with useful instructions instead.
+Hence, we would like to know whether our program is bottlenecked by these expensive memory operations, which we very much would like to avoid, as we want to utilize our CPU with useful instructions instead.
 
@@
 
-If our program is not memory-bound, it still executes not instantly, as it is bound by the execution speed of the cpu.
+If our program is not memory-bound, it still executes not instantly, as it is bound by the execution speed of the CPU.
 We have a \em compute-bound\em .
 In a sense, every program or operation experiences a compute-bound.
-Hence, it is more useful when we quantify this bound, by analyzing the cpu frequency and the execution speed of individual operations, so that we can estimate the program's execution time.
+Hence, it is more useful when we quantify this bound, by analyzing the CPU frequency and the execution speed of individual operations, so that we can estimate the program's execution time.
 
@@
 
-This is, because a lot of data will read from and written to memory, which, as explained, is very time consuming.
+This is, because a lot of data will be read from and written to memory, which, as explained, is very time consuming.
 Luckily, computer engineers introduced caches to this end, which store some data of main memory, and provide it to the program with faster access times.
-The only problem will be if the caches or, more precisely, their hash bucket are full.
+The only problem will be if the caches, or, more precisely, their hash bucket are full.
 Because caches are usually not \em fully-associative\em , which means that if we have a cache line in memory, we can not write it to any cache line in the cache as we like.
@@
     \item[Temporal locality] describes the concept that recently accessed memory has a higher probability to be accessed again.
-    \item[Spatial locality] on the other hand is used to refer to the concept that memory, which is spatially close to recently accessed memory, has a higher probability of being accessed themself.
+    \item[Spatial locality] on the other hand is used to refer to the concept that memory, which is spatially close to recently accessed memory, has a higher probability of being accessed itself.
 \end{description}
 
+\noindent
 But why is all that useful?
@@
 Hence, when running the left algorithm, $a$ has good spatial locality, whereas $b$'s is horrendous, as it leads to a cache miss on every access.
-The idea is now to find a tradeoff between these two.
+The idea now is to find a tradeoff between these two.
 Let us analyze the access pattern of the two algorithms:

### 2025-03-04 Exam Assignments/Chapters/01/calculate_pi.cpp
restructuring

@@ -1,38 +1,0 @@
-#include <iostream>
-#include <omp.h>
-#include <random>
-
-using namespace std;
-
-int main() {
-    
-    int n = 100000000; // number of points to generate
-    int counter = 0; // counter for points lying in the first quadrant of a unit circle
-    auto start_time = omp_get_wtime(); // omp_get_wtime() is an OpenMP library routine
-
-    // compute n points and test if they lie within the first quadrant of a unit circle
-    #pragma omp parallel
-    {
-        default_random_engine re{(size_t) omp_get_thread_num()};
-        uniform_real_distribution<double> zero_to_one{0.0, 1.0};
-
-        int local_counter = 0;
-        int local_n = (n / omp_get_num_threads()) + ((n % omp_get_num_threads() > omp_get_thread_num()) ? 1 : 0);
-        for (int i = 0; i < local_n; ++i) {
-            auto x = zero_to_one(re); // generate random number between 0.0 and 1.0
-            auto y = zero_to_one(re); // generate random number between 0.0 and 1.0
-            if (x * x + y * y <= 1.0) { // if the point lies in the first quadrant of a unit circle
-                ++local_counter;
-            }
-        }
-        #pragma omp atomic
-        counter += local_counter;
-    }
-
-    auto run_time = omp_get_wtime() - start_time;
-    auto pi = 4 * (double(counter) / n);
-
-    cout << "pi: " << pi << endl;
-    cout << "run_time: " << run_time << " s" << endl;
-    cout << "n: " << n << endl;
-}

### 2025-03-04 Exam Assignments/Chapters/03/calculate_pi_v2.cpp

@@ -1,42 +1,0 @@
-#include <iostream>
-#include <omp.h>
-#include <random>
-
-using namespace std;
-
-double rnd(unsigned int *seed)
-{
-    *seed = (1140671485 * (*seed) + 12820163) % (1 << 24);
-    return ((double)(*seed)) / (1 << 24);
-}
-
-int main()
-{
-    int n = 100000000;                 // number of points to generate
-    int counter = 0;                   // counter for points lying in the first quadrant of a unit circle
-    auto start_time = omp_get_wtime(); // omp_get_wtime() is an OpenMP library routine
-
-    // compute n points and test if they lie within the first quadrant of a unit circle
-    #pragma omp parallel reduction(+:counter)
-    {
-        unsigned int seed = omp_get_thread_num();
-        size_t local_n = (n / omp_get_num_threads()) + ((n % omp_get_num_threads() > omp_get_thread_num()) ? 1 : 0);
-
-        for (int i = 0; i < local_n; ++i)
-        {
-            auto x = rnd(&seed); // generate random number between 0.0 and 1.0
-            auto y = rnd(&seed); // generate random number between 0.0 and 1.0
-            if (x * x + y * y <= 1.0)
-            { // if the point lies in the first quadrant of a unit circle
-                ++counter;
-            }
-        }
-    }
-
-    auto run_time = omp_get_wtime() - start_time;
-    auto pi = 4 * (double(counter) / n);
-
-    cout << "pi: " << pi << endl;
-    cout << "run_time: " << run_time << " s" << endl;
-    cout << "n: " << n << endl;
-}

### 2025-03-04 Exam Assignments/Chapters/04/min_max_quicksort.cpp

@@ -1,167 +1,0 @@
-#include <iostream>
-#include <vector>
-#include <limits>
-#include <omp.h>
-#include <parallel/algorithm>
-#include <fstream>
-
-// compute the average of two integers without overflow
-inline int64_t average(int64_t a, int64_t b) {
-    return (a & b) + ((a ^ b) >> 1);
-}
-
-// partitioning function for quicksort
-inline int64_t
-partition(int64_t *arr, int64_t left, int64_t right, int64_t pivot, int64_t &smallest, int64_t &biggest) {
-    int64_t *left_ptr = &arr[left];
-    int64_t *right_ptr = &arr[right];
-    while (left_ptr < right_ptr) {
-        smallest = (*left_ptr < smallest) ? *left_ptr : smallest;
-        biggest = (*left_ptr > biggest) ? *left_ptr : biggest;
-        if (*left_ptr > pivot) {
-            --right_ptr;
-            std::swap(*left_ptr, *right_ptr);
-        } else {
-            ++left_ptr;
-        }
-    }
-    return left_ptr - arr;
-}
-
-inline void insertion_sort(int64_t *arr, int64_t left, int64_t right) {
-    for (int64_t i = left + 1; i <= right; i++) {
-        int64_t key = arr[i];
-        int64_t j = i - 1;
-        while (j >= left && arr[j] > key) {
-            arr[j + 1] = arr[j];
-            j = j - 1;
-        }
-        arr[j + 1] = key;
-    }
-}
-
-// the core recursive quicksort function
-void qs_core(int64_t *arr, int64_t left, int64_t right, const int64_t pivot) {
-    if (right - left < 32) {
-        insertion_sort(arr, left, right);
-        return;
-    }
-
-    int64_t smallest = std::numeric_limits<int64_t>::max();
-    int64_t biggest = std::numeric_limits<int64_t>::min();
-    int64_t bound = partition(arr, left, right + 1, pivot, smallest, biggest);
-
-    if (smallest == biggest)
-        return;
-
-#pragma omp task final(bound - left < 10000)
-    qs_core(arr, left, bound - 1, average(smallest, pivot));
-    qs_core(arr, bound, right, average(pivot, biggest));
-}
-
-// wrapper for the quicksort function
-void min_max_quicksort(int64_t *arr, int64_t n, int num_threads = omp_get_max_threads()) {
-#pragma omp parallel num_threads(num_threads)
-#pragma omp single nowait
[…]

### 2025-03-04 Exam Assignments/Chapters/04/min_max_quicksort_array_sizes.cpp

@@ -1,165 +1,0 @@
-#include <iostream>
-#include <vector>
-#include <limits>
-#include <omp.h>
-#include <parallel/algorithm>
-#include <fstream>
-
-// compute the average of two integers without overflow
-inline int64_t average(int64_t a, int64_t b) {
-    return (a & b) + ((a ^ b) >> 1);
-}
-
-// partitioning function for quicksort
-inline int64_t
-partition(int64_t *arr, int64_t left, int64_t right, int64_t pivot, int64_t &smallest, int64_t &biggest) {
-    int64_t *left_ptr = &arr[left];
-    int64_t *right_ptr = &arr[right];
-    while (left_ptr < right_ptr) {
-        smallest = (*left_ptr < smallest) ? *left_ptr : smallest;
-        biggest = (*left_ptr > biggest) ? *left_ptr : biggest;
-        if (*left_ptr > pivot) {
-            --right_ptr;
-            std::swap(*left_ptr, *right_ptr);
-        } else {
-            ++left_ptr;
-        }
-    }
-    return left_ptr - arr;
-}
-
-inline void insertion_sort(int64_t *arr, int64_t left, int64_t right) {
-    for (int64_t i = left + 1; i <= right; i++) {
-        int64_t key = arr[i];
-        int64_t j = i - 1;
-        while (j >= left && arr[j] > key) {
-            arr[j + 1] = arr[j];
-            j = j - 1;
-        }
-        arr[j + 1] = key;
-    }
-}
-
-// the core recursive quicksort function
-void qs_core(int64_t *arr, int64_t left, int64_t right, const int64_t pivot) {
-    if (right - left < 32) {
-        insertion_sort(arr, left, right);
-        return;
-    }
-
-    int64_t smallest = std::numeric_limits<int64_t>::max();
-    int64_t biggest = std::numeric_limits<int64_t>::min();
-    int64_t bound = partition(arr, left, right + 1, pivot, smallest, biggest);
-
-    if (smallest == biggest)
-        return;
-
-#pragma omp task final(bound - left < 10000)
-    qs_core(arr, left, bound - 1, average(smallest, pivot));
-    qs_core(arr, bound, right, average(pivot, biggest));
-}
-
-// wrapper for the quicksort function
-void min_max_quicksort(int64_t *arr, int64_t n, int num_threads = omp_get_max_threads()) {
-#pragma omp parallel num_threads(num_threads)
-#pragma omp single nowait
[…]

### 2025-03-04 Exam Assignments/Chapters/04/plot.py

@@ -1,31 +1,0 @@
-import pandas as pd
-import matplotlib.pyplot as plt
-
-# Read the CSV file into a pandas DataFrame
-# df = pd.read_csv('/home/jonas/Documents/University/5. Semester/Algorithm Engineering/Exam Assignments/Chapters/04/sorting_algorithms_running_times.csv')
-df = pd.read_csv('/home/jonas/Documents/University/5. Semester/Algorithm Engineering/Exam Assignments/Chapters/04/speedup_array.csv')
-
-print(df.columns)
-# Plotting the data
-plt.figure(figsize=(10, 6))
-
-# Loop through each column (except for the first one 'threads') and plot the data
-for column in df.columns[1:]:
-    plt.plot(df['Array Size'], df[column], label=f' {column}', marker='o')
-
-# Add titles and labels
-plt.xscale('log')
-plt.title('Speedup over std::sort at varying array sizes', fontsize=14)
-# plt.title('Speedup over std::sort (Time vs. Array Size)', fontsize=14)
-# plt.title('Performance of Sorting Algorithms (Time vs. Threads)', fontsize=14)
-plt.xlabel('Array Size (in MB)', fontsize=12)
-# plt.xlabel('Threads', fontsize=12)
-# plt.ylabel('Time (seconds)', fontsize=12)
-plt.ylabel('Speedup', fontsize=12)
-
-# Show grid and legend
-plt.grid(True)
-plt.legend()
-
-# Show the plot
-plt.show()

### 2025-03-04 Exam Assignments/Code/05/Catch2/catch_tests_fizzbuzz.cpp

@@ -1,50 +1,0 @@
-#include "catch.hpp"
-#include "fizzbuzz.h"
-
-TEST_CASE("Test positives", "[classic]")
-{
-   SECTION("Test all up to 10") {
-      REQUIRE(fizzbuzz(1) == "1"); 
-      REQUIRE(fizzbuzz(2) == "2");
-      REQUIRE(fizzbuzz(3) == "fizz");
-      REQUIRE(fizzbuzz(4) == "4");
-      REQUIRE(fizzbuzz(5) == "buzz");
-      REQUIRE(fizzbuzz(6) == "fizz");
-      REQUIRE(fizzbuzz(7) == "7");
-      REQUIRE(fizzbuzz(8) == "8");
-      REQUIRE(fizzbuzz(9) == "fizz");
-      REQUIRE(fizzbuzz(10) == "buzz");
-   }
-
-   SECTION("Test all multiples of 3 only up to 100") {
-      for (int i = 3; i <= 100; i+=3) {
-         if (i % 5) REQUIRE(fizzbuzz(i) == "fizz");
-      }
-   }
-
-   SECTION("Test all multiples of 5 only up to 100") {
-      for (int i = 5; i <= 100; i += 5) {
-         if (i % 3) REQUIRE(fizzbuzz(i) == "buzz");
-      }
-   }
-
-   SECTION("Test all multiples of 3 and 5 up to 100") {
-      for (int i = 15; i <= 100; i += 15) {
-         REQUIRE(fizzbuzz(i) == "fizzbuzz");
-      }
-   }
-}
-
-TEST_CASE("Test negatives", "[classic]")
-{
-   REQUIRE(fizzbuzz(-1) == "-1");
-   REQUIRE(fizzbuzz(-2) == "-2");
-   REQUIRE(fizzbuzz(-3) == "fizz");
-   REQUIRE(fizzbuzz(-4) == "-4");
-   REQUIRE(fizzbuzz(-5) == "buzz");
-   REQUIRE(fizzbuzz(-6) == "fizz");
-   REQUIRE(fizzbuzz(-7) == "-7");
-   REQUIRE(fizzbuzz(-8) == "-8");
-   REQUIRE(fizzbuzz(-9) == "fizz");
-   REQUIRE(fizzbuzz(-10) == "buzz");
-}

### 2025-03-04 Exam Assignments/Code/05/Catch2/fizzbuzz/include/fizzbuzz.h

@@ -1,3 +1,0 @@
-#include <string>
-
-std::string fizzbuzz(int const);

### 2025-03-04 Exam Assignments/Code/05/Catch2/fizzbuzz/src/fizzbuzz.cpp

@@ -1,14 +1,0 @@
-#include <string>
-
-std::string fizzbuzz(int const number)
-{
-   if (number != 0)
-   {
-      auto m3 = number % 3;
-      auto m5 = number % 5;
-      if (!m5 && !m3) { return "fizzbuzz"; }
-      else if (!m5) { return "buzz"; }
-      else if (!m3) { return "fizz"; }
-   }
-   return std::to_string(number);
-}

### 2025-03-04 Exam Assignments/Code/05/quantpng/CMakeFiles/quantpng.dir/compiler_depend.ts

@@ -1,2 +1,0 @@
-# CMAKE generated file: DO NOT EDIT!
-# Timestamp file for compiler generated dependencies management for quantpng.

### 2025-03-04 answers/chapters/01/01.tex
restructuring

--- 2024-11-10 Exam Assignments/Chapters/01/01.tex
@@
 
-~\\
+\begin{figure}[h]
     \includegraphics[width=\textwidth]{page_15.png}
+    \caption{Slide 15 from lecture 1}
+\end{figure}
 
 ~\\
-Let us proof the satement
+Let us prove the statement
 \[
@@
 
+\noindent
 Even better is the approximation by parabolas, which pass through the end points of the interval as well as the mid point. This is called the \em Simpson Rule\em.
@@
 \]
-
+\noindent
 where \(I_{(t, t+h)}(f)\) tries to approximate \(\int_{t}^{t+h} f(x) dx\).
@@
 \subsubsection{Memory Hierarchy}
-The inverse correlation between speed and volume of different memory types war presented in the book.
+The inverse correlation between speed and volume of different memory types was presented in the book.
 This lead to the memory hierarchy, where we try exploit the benefits of both worlds be having big, slow memory, and a smaller amount of fast memory, which we the computer tries to employ as much as possible.

### 2025-03-04 answers/chapters/11/11.tex
restructuring

--- 2025-01-07 Exam Assignments/Chapters/11/11.tex
@@
 First, we should keep in mind the page size of the SSD, which typically is 4KB.
-Hence, small read and write operation below this mark will have the same cost of a 4KB one, and in case of a write operation we may not fully utilized the space as the page we are writing to flash itself is not fully utilized.
+Hence, small read and write operation below this mark will have the same cost of a 4KB one, and in case of a write operation we may not fully utilize the space as the page we are writing to flash itself is not fully utilized.
 Thus, we should combine smaller IO operations in two fewer bigger ones, provided that these operations are sequential, which directly leads us to the second conclusion:
@@
 
-~\\
+\bigskip
 \subsection{Designing SSD-friendly Applications for Better Application Performance and Higher IO Efficiency}
-This paper introduces the reader to the core functioning of SSDs and presents different optimization techniques to improve performance or longevity of the SSD.
+\href{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=7552037}{This paper} introduces the reader to the core functioning of SSDs and presents different optimization techniques to improve performance or longevity of the SSD.
 One optimization presented in the paper directly follows on from the previous discussion.
@@
 The optimum for all data would be a smooth transition between these two extrema.
-The reason for this is very similar as before, as when accessing data from the SSD we also access unwanted data due the teh page size.
+The reason for this is very similar as before, as when accessing data from the SSD we also access unwanted data due the page size.
 Optimally, we also require this additional data, otherwise this data section is accessed redundantly.

### 2025-03-04 paper/paper.tex
restructuring

--- 2024-12-28 Exam Assignments/paper/paper.tex
@@
 We face the \textbf{exact} problems we faced at the reduction method, and my responses are the same, we just keep the implementation simple for the sake of demonstration.
-This function could be manually vectorized as well, but the compiler should do a decent job at optimizing this function, and hence we shouldn't face a big performance penalty.
+This function can be manually vectorized as well, however the compiler should do a decent job at optimizing this function, and hence we shouldn't face a big performance penalty.
+For \texttt{float} and \texttt{double} vectorized code can be found in specialized template methods below.
 

### 2025-03-04 answers/code/02/pi_numerical_integration.cpp

@@ -1,0 +1,23 @@
+#include <iomanip>
+#include <iostream>
+#include <omp.h>
+
+using namespace std;
+
+int main() {
+    int num_steps = 100000000; // number of rectangles
+    double width = 1.0 / double(num_steps); // width of a rectangle
+    double sum = 0.0; // for summing up all heights of rectangles
+
+    double start_time = omp_get_wtime(); // wall clock time in seconds
+    #pragma omp parallel for reduction(+:sum)
+    for (int i = 0; i < num_steps; i++) {
+        double x = (i + 0.5) * width; // midpoint
+        sum = sum + (1.0 / (1.0 + x * x)); // add new height of a rectangle
+    }
+    double pi = sum * 4 * width; // compute pi
+    double run_time = omp_get_wtime() - start_time;
+
+    cout << "pi with " << num_steps << " steps is " << setprecision(17)
+        << pi << " in " << setprecision(6) << run_time << " seconds\n";
+}

### 2025-03-04 answers/chapters/06/06.tex
reworking chapter 6

\documentclass[../../main.tex]{subfiles}

\begin{document}
    \section{Auto Vectorization}
    \subsection{Characteristics of SSE, AVX(2) and AVX-512}

    \begin{description}
        \item[SSE (Streaming SIMD Extensions):] ~\\ Introduced in 1999 by Intel in its Pentium III series\footnotemark,  SSE supports 128-bit vector registers, enabling SIMD operations on 4 single-precision floating-point or 2 double-precision values in parallel. Limited in width and functionality compared to later extensions. \\
\footnotetext{Source: \href{https://en.wikipedia.org/wiki/Streaming\_SIMD\_Extensions}{https://en.wikipedia.org/wiki/Streaming\_SIMD\_Extensions}}
        \item[AVX/AVX2 (Advanced Vector Extensions):] ~\\ AVX were proposed by Intel in March 2008 and they extended the SIMD width from 128 to 256 bits, doubling parallelism compared to SSE\footnotemark. AVX employs 16 vector registers and supports various additional instructions, like for instance so called \em three-operand SIMD instructions\em, which store the result of the operation in a third independent register. AVX2 expand most integer instructions to 256 bits and introduced yet again new features like gathering vector elements from non-contiguous memory locations.  \\
\footnotetext{Source: \href{https://en.wikipedia.org/wiki/Advanced\_Vector\_Extensions}{https://en.wikipedia.org/wiki/Advanced\_Vector\_Extensions}}
        \item[AVX-512:] ~\\ Widens SIMD to 512 bits, allowing operations on 16 single-precision floats or 8 double-precision values simultaneously. It increases the number of registers to 32, and adds 8 new mask registers for conditional operations\footnotemark. New instructions and features were added, like including 4 operand operations, and introducing explicit rounding control, etc.\\
\footnotetext{Source: \href{https://en.wikipedia.org/wiki/AVX-512}{https://en.wikipedia.org/wiki/AVX-512}}
    \end{description}

    % \bigskip
    \subsection{Impact of Memory Aliasing on Performance}
    Memory aliasing occurs when two or more pointers reference overlapping memory regions, making it difficult for the compiler or CPU to optimize memory accesses.
    To illustrate this, say we have the following code:

    \begin{lstlisting}
void add(float *a, float *b, int n) {
    for (int i = 0; i < n; i++) {
        a[i] = a[i] + b[i]; 
    }
}
    \end{lstlisting}

    \noindent
    At first glance, it looks well written such that the compiler has no difficulty vectorizing it. But this is a fallacy, as the programmer has made the hidden assumption that $\texttt{a}$ and $\texttt{b}$ don't overlap. For the argument's sake, let's assume that $\texttt{a = b + 1}$. This would mean that when updating $\texttt{a[i]}$, we also update $\texttt{b[i+1]}$, which is accessed in the following iteration. Thus, we have a dependency of every iteration to the previous one, and hence vectorizing it would change the the programs behavior and hence the compiler won't do it. The solution of course is to mark $\texttt{a}$ and $\texttt{b}$ as $\texttt{\_\_restrict\_\_}$ if they indeed point to not-overlapping memory regions.

    \bigskip
    \subsection{Advantages of Unit Stride Memory Access}
    Unit stride (stride-1) memory access sequentially accesses adjacent memory locations, maximizing cache utilization and resulting in higher bandwidth utilization. Furthermore, unit stride access patterns facilitate the utilization of efficient vector instructions employed by the compiler. Larger strides on the other hand, like stride-8, result in less efficient cache usage and increased memory latency. Vectorization is also harder and less efficient. Thus, it is generally a good advise trying to employ unit strides when accessing the memory whenever possible.

    \bigskip
    \subsection{When to Prefer Structure of Arrays}
    Structure of Arrays (SoA) is preferred when working with SIMD/vectorized operations or when the workload requires processing individual fields of a dataset independently. This is simply due to the fact of strided and homogenous memory access like discussed previously. Especially for big data with these mentioned operations one should consider SoA over AoS (Array of Structures), which, on the other hand, may be used in less compute intensive tasks to improve readability and maintainability of the code, or in cases where the access patterns would align with such a layout.

\end{document}

### 2025-03-05 project/.legacy/einspeed copy.cpp
Writing Readme.txt for einspeed project

@@ -1,56 +1,0 @@
-#define NPY_NO_DEPRECATED_API NPY_1_7_API_VERSION
-#include <Python.h>
-#include <numpy/arrayobject.h>
-#include "parse_tensor.h"
-#include "tensor.h"
-
-#include <omp.h>
-#include <iostream>
-
-template <typename T>
-static PyObject* einsum(PyObject* self, PyObject* args) {
-// *** <Parsing Arguments> ***
-    PyObject *t1, *t2;
-    const char *s;
-    // Parse the input argument as a generic PyObject
-    if (!PyArg_ParseTuple(args, "sOO", &s, &t1, &t2)) {
-        return NULL;
-    }
-    // TODO: Check for nullptr
-    auto tensor1 = parse_tensor<T>(t1);
-    auto tensor2 = parse_tensor<T>(t2);
-// *** </Parsing Arguments>
-
-
-    std::cout << tensor1->data[3] << "\n";
-
-    Py_RETURN_NONE;
-}
-
-
-static PyObject* einsumd(PyObject* self, PyObject* args) {
-    return einsum<double>(self, args);
-}
-
-static PyObject* einsum(PyObject* self, PyObject* args) {
-    return einsum<float>(self, args);
-}
-
-static PyMethodDef MyMethods[] = {
-    {"einsumd", einsumd, METH_VARARGS, "einsum two double tensors"},
-    {"einsum", einsum, METH_VARARGS, "einsum two float tensors"},
-    {NULL, NULL, 0, NULL}
-};
-
-static struct PyModuleDef einspeed = {
-    PyModuleDef_HEAD_INIT,
-    "einspeed",
-    NULL,
-    -1,
-    MyMethods
-};
-
-PyMODINIT_FUNC PyInit_einspeed(void) {
-    import_array(); // Initialize NumPy C API
-    return PyModule_Create(&einspeed);
-}

### 2025-03-05 project/.legacy/func.cpp

@@ -1,9 +1,0 @@
-// #include "func.h"
-
-// template <typename FROM, typename TO>
-// TO* cast_all(size_t _n, FROM *_a) {
-//     TO *_r = new TO[_n];
-//     for (size_t i = 0; i < _n; i++)
-//         _r[i] = static_cast<TO>(_a[i]);
-//     return _r;
-// }

### 2025-03-05 project/.legacy/setup.py

@@ -1,34 +1,0 @@
-# from setuptools import setup, Extension
-# import numpy as np
-
-# module = Extension(
-#     "mymodule",
-#     sources=["mymodule.c"],
-#     include_dirs=[np.get_include()]
-# )
-
-# setup(
-#     name="mymodule",
-#     version="1.0",
-#     ext_modules=[module]
-# )
-
-
-from setuptools import setup, Extension
-import numpy as np
-
-module = Extension(
-    "mymodule",
-    sources=["mymodule.cpp"],
-    include_dirs=[np.get_include()],
-    extra_compile_args=["-O3", "-fopenmp"],  # Optimization and OpenMP support
-    extra_link_args=["-fopenmp"],            # Link OpenMP libraries
-)
-
-setup(
-    name="mymodule",
-    version="1.0",
-    ext_modules=[module],
-)
-
-

### 2025-03-05 project/.legacy/tensor.h

@@ -1,288 +1,0 @@
-#ifndef TENSOR_H
-#define TENSOR_H
-
-#include <Python.h>
-#include <numpy/arrayobject.h>
-// #include <cstddef>
-#include <vector>
-#include <algorithm>
-#include "numpy_types.h"
-// #include <stdexcept>
-#include <cstring> // For memcpy
-
-#include <iostream>
-
-template <typename T>
-class Tensor {
-public:
-    size_t ndim = 0;          // Number of dimensions
-    std::vector<size_t> shape = nullptr;  // Pointer to array defining the shape
-    T* data = nullptr;        // Pointer to the actual data
-
-    Tensor(size_t ndim, std::vector<size_t> shape, T* data)
-        : ndim(ndim), shape(shape), data(data) {};
-    Tensor(size_t ndim, size_t *shape, T *data) 
-        : ndim(ndim), shape(shape, shape + ndim), data(data) {}
-    Tensor() {}
-    
-    Tensor(Tensor<T> &&other)
-        : ndim(other.ndim), shape(other.shape), data(other.data) {
-        // Nullify the moved-from object to avoid double deletion
-        other.ndim = 0;
-        other.shape = nullptr;
-        other.data = nullptr;
-    }
-
-    Tensor<T>& operator=(Tensor<T> &&other) {
-        if (this != &other) {  // Prevent self-assignment
-            // Clean up any existing resources (if necessary)
-            delete[] shape;
-            // Don't delete data as it belongs to Python or is managed elsewhere.
-
-            // Move the resources
-            ndim = other.ndim;
-            shape = other.shape;
-            data = other.data;
-
-            // Nullify the moved-from object
-            other.ndim = 0;
-            other.shape = nullptr;
-            other.data = nullptr;
-        }
-        return *this;
-    }
-
-
-    ~Tensor() {
-        std::cout << "DESTROYING TENSOR:\n";
-        // this->print();
-        if (this->shape) delete[] this->shape; // It is okay to delete this, since when casting to PyObject, we will create a copy (which we will later also delete though)
-        // if (this->data) delete[] this->data; // Don't delete data as it belongs to Python
-    }
-
-    PyObject* to_PyObject() const {
-        // Initialize NumPy (if not already done)
-        if (!PyArray_API) {
-            import_array(); // Necessary to initialize NumPy C API
-        }
-
-        // Create a NumPy array
-        npy_intp* np_shape = new npy_intp[ndim];
-        for (size_t i = 0; i < ndim; ++i) {
[…]

### 2025-03-05 project/include/blas.h

@@ -3,6 +3,8 @@
 
 #include "tensor.h"
 
+// Implementation of batch matrix matrix multiplication, where the second batch matrices are treated as being transposed.
+
 template <typename T>
 Tensor<T> batch_matrix_matmul_transpose(const Tensor<T>& lhs, const Tensor<T>& rhs, T* available_memory, size_t available_memory_size) {
     // Check dimensions

### 2025-03-05 project/include/func.h

@@ -7,6 +7,8 @@
 #include <cctype>
 #include <locale>
 
+// In this file some basic subroutines are implemented.
+
 template <typename FROM, typename TO>
 TO* cast_all(size_t _n, const FROM* _a) {
     TO *_r = new TO[_n];

### 2025-03-05 project/include/numpy_types.h

@@ -5,6 +5,8 @@
 #include <numpy/arrayobject.h>
 #include <complex>
 
+// This file specifies the conversion from NumPy types to native C++ types.
+
 template <typename T>
 constexpr int numpy_type_of() {throw std::runtime_error("Unsupported data type for NumPy array");}
 

### 2025-03-05 project/include/parse_tensor.h

@@ -6,6 +6,8 @@
 #include "tensor.h"
 #include "func.h"
 
+// Some parsing methods of Python objects and NumPy arrays are implemented here.
+
 const PyArrayObject* PyObject_to_PyArrayObject(const PyObject * const _t) {
     // Ensure the input is a NumPy array
     if (!PyArray_Check(_t)) {

### 2025-03-05 project/include/tensor.h

@@ -16,6 +16,9 @@
 #include "func.h"
 #include "hptt.h"
 
+// This class implements the main tensor data structure used by this einsum implementation.
+// Some of the important methods are transpose and reduce.
+
 template <typename T>
 class Tensor {
 public:

### 2025-03-05 project/include/unroll_loop.h

@@ -3,7 +3,9 @@
 
 #include <utility>
 
+// This is used in scr/einspeed.cpp to loop over the supported data types.
 // Source: https://stackoverflow.com/questions/37602057/why-isnt-a-for-loop-a-compile-time-expression
+
 template<std::size_t N>
 struct num { static const constexpr auto value = N; };
 

### 2025-03-05 project/src/einspeed.cpp

@@ -7,6 +7,8 @@
 #include "numpy_types.h"
 #include "unroll_loop.h"
 
+// Entry point for Python to use the einsum function found below.
+
 static PyObject* einsum(PyObject* const self, PyObject* const args) {
     PyObject *lhs_o, *rhs_o;
     const char *s;

### 2025-03-05 project/.legacy/catch_tests_fizzbuzz.cpp
Writing Readme.txt for einspeed project

@@ -1,55 +1,0 @@
-#include "catch.hpp"
-#include "tensor.h"
-
-TEST_CASE("moin", "[classic]") {
-   SECTION("LOL") {
-      REQUIRE(true);
-   }
-}
-// TEST_CASE("Test positives", "[classic]")
-// {
-//    SECTION("Test all up to 10") {
-//       REQUIRE(fizzbuzz(1) == "1"); 
-//       REQUIRE(fizzbuzz(2) == "2");
-//       REQUIRE(fizzbuzz(3) == "fizz");
-//       REQUIRE(fizzbuzz(4) == "4");
-//       REQUIRE(fizzbuzz(5) == "buzz");
-//       REQUIRE(fizzbuzz(6) == "fizz");
-//       REQUIRE(fizzbuzz(7) == "7");
-//       REQUIRE(fizzbuzz(8) == "8");
-//       REQUIRE(fizzbuzz(9) == "fizz");
-//       REQUIRE(fizzbuzz(10) == "buzz");
-//    }
-
-//    SECTION("Test all multiples of 3 only up to 100") {
-//       for (int i = 3; i <= 100; i+=3) {
-//          if (i % 5) REQUIRE(fizzbuzz(i) == "fizz");
-//       }
-//    }
-
-//    SECTION("Test all multiples of 5 only up to 100") {
-//       for (int i = 5; i <= 100; i += 5) {
-//          if (i % 3) REQUIRE(fizzbuzz(i) == "buzz");
-//       }
-//    }
-
-//    SECTION("Test all multiples of 3 and 5 up to 100") {
-//       for (int i = 15; i <= 100; i += 15) {
-//          REQUIRE(fizzbuzz(i) == "fizzbuzz");
-//       }
-//    }
-// }
-
-// TEST_CASE("Test negatives", "[classic]")
-// {
-//    REQUIRE(fizzbuzz(-1) == "-1");
-//    REQUIRE(fizzbuzz(-2) == "-2");
-//    REQUIRE(fizzbuzz(-3) == "fizz");
-//    REQUIRE(fizzbuzz(-4) == "-4");
-//    REQUIRE(fizzbuzz(-5) == "buzz");
-//    REQUIRE(fizzbuzz(-6) == "fizz");
-//    REQUIRE(fizzbuzz(-7) == "-7");
-//    REQUIRE(fizzbuzz(-8) == "-8");
[…]

### 2025-03-05 project/.legacy/parse_tensor.cpp

@@ -1,41 +1,0 @@
-// #include "parse_tensor.h"
-// #include "func.h"
-
-// // Mapping C++ types to NumPy type constants
-// template <typename T>
-// constexpr int _numpy_type();
-// template <>
-// constexpr int _numpy_type<double>() { return NPY_DOUBLE; }
-// template <>
-// constexpr int _numpy_type<float>() { return NPY_FLOAT; }
-// template <>
-// constexpr int _numpy_type<int>() { return NPY_INT; }
-// template <>
-// constexpr int _numpy_type<long>() { return NPY_LONG; }
-
-
-// template <typename T>
-// Tensor<T>* parse_tensor(PyObject *_t) {
-//     // Ensure the input is a NumPy array
-//     if (!PyArray_Check(_t)) {
-//         PyErr_SetString(PyExc_TypeError, "Expected a NumPy array");
-//         return nullptr;
-//     }
-
-//     // Cast to PyArrayObject and check type
-//     PyArrayObject* _a = reinterpret_cast<PyArrayObject*>(_t);
-//     if (PyArray_TYPE(_a) != _numpy_type<T>()) {
-//         // TODO: customize Error message
-//         PyErr_SetString(PyExc_TypeError, "Expected a NumPy array of type double");
-//         return NULL;
-//     }
-
-//     // Get number of dimensions and shape
-//     size_t ndim = (size_t) PyArray_NDIM(_a);
-//     // npy_intp* shape = PyArray_DIMS(array);  // Array of dimension sizes
-
-//     // Get a pointer to the data
-//     // double* data = static_cast<double*>(PyArray_DATA(array));
-
-//     return new Tensor<T>(PyArray_NDIM(_a), cast_all<npy_intp, size_t>(ndim, PyArray_DIMS(_a)), static_cast<T*>(PyArray_DATA(_a)));
-// }

### 2025-03-05 project/.legacy/tensor.cpp

@@ -1,11 +1,0 @@
-// #include "tensor.h"
-
-// template <typename T>
-// Tensor<T>::Tensor(size_t ndim, size_t* shape, T* data)
-//     : ndim(ndim), shape(shape), data(data) {}
-
-// template <typename T>
-// Tensor<T>::~Tensor() {
-//     delete[] shape;
-//     // delete[] data; Don't delete data as it belongs to Python
-// }

### 2025-03-05 project/.legacy/tests/catch_main.cpp

@@ -1,3 +1,0 @@
-#define CATCH_CONFIG_MAIN
-#define CATCH_CONFIG_FAST_COMPILE // for faster recompilation
-#include "catch.hpp"

### 2025-03-05 project/.legacy/tests/catch_tests_blas.cpp

@@ -1,63 +1,0 @@
-#define CATCH_CONFIG_MAIN
-#include "catch.hpp"
-#include "tensor.h"
-#include "blas.h"
-
-TEST_CASE("Batch Matrix Multiplication", "[batch_matmul]") {
-    REQUIRE(true);
-    // SECTION("Basic batch multiplication with 2x2 matrices") {
-    //     // Create two tensors with batch size 2 and matrices of size 2x2
-    //     size_t shape[3] = {2, 2, 2}; // Batch size 2, 2x2 matrices
-    //     double data_lhs[] = {1, 2, 3, 4,   // First matrix in batch
-    //                          5, 6, 7, 8};  // Second matrix in batch
-
-    //     double data_rhs[] = {9, 10, 11, 12,  // First matrix in batch
-    //                          13, 14, 15, 16}; // Second matrix in batch
-
-    //     Tensor<double> lhs(3, shape, data_lhs);
-    //     Tensor<double> rhs(3, shape, data_rhs);
-
-    //     // Expected output: Batch of multiplied matrices
-    //     double expected_data[] = {
-    //         31, 34,  71, 78,  // First batch result
-    //         149, 162, 221, 240 // Second batch result
-    //     };
-    //     Tensor<double> expected_result(3, shape, expected_data);
-
-    //     // Perform batch matrix multiplication
-    //     auto result = batch_matmul(lhs, rhs);
-
-    //     // Compare the resulting tensor with the expected tensor
-    //     REQUIRE(result.ndim == expected_result.ndim);
-    //     REQUIRE(std::equal(result.shape, result.shape + 3, expected_result.shape));
-    //     REQUIRE(std::equal(result.data, result.data + 8, expected_result.data));
-    // }
-
-    // SECTION("Invalid batch dimensions should throw an exception") {
-    //     // Batch size mismatch between tensors
-    //     size_t shape_lhs[3] = {2, 2, 2};
-    //     size_t shape_rhs[3] = {3, 2, 2}; // Batch size mismatch
-    //     double data[] = {1, 2, 3, 4, 5, 6, 7, 8};
-
-    //     Tensor<double> lhs(3, shape_lhs, data);
-    //     Tensor<double> rhs(3, shape_rhs, data);
-
-    //     // Expect an exception due to mismatched batch dimensions
[…]

### 2025-03-05 project/include/compute_einsum (copy).h

@@ -1,321 +1,0 @@
-#ifndef COMPUTE_EINSUM_H
-#define COMPUTE_EINSUM_H
-
-// uncomment the following lines to show debug information
-// #define DEBUG
-// #define DEBUG_VERBOSE
-
-#include <unordered_map>
-#include <unordered_set>
-#include <stdexcept>
-#include <string>
-
-#include "tensor.h"
-#include "blas.h"
-#include "func.h"
-
-
-// **************************************************
-// *** The core functionality is implemented HERE ***
-// **************************************************
-
-
-// Some declarations (implementation can be found below)
-std::vector<size_t> column_identifiers_like(const std::vector<size_t>& target_column_identifiers, const std::string& target_identifier_string, const std::string& origin_identifier_string);
-template <typename T> bool is_valid_einsum_expression(const char* const s, const Tensor<T>& lhs_tensor, const Tensor<T>& rhs_tensor, std::string& lhs_string, std::string& rhs_string, std::string& target_string);
-
-
-template <typename T>
-PyObject* compute_einsum(const char* const s, const Tensor<T>& lhs_tensor, const Tensor<T>& rhs_tensor) {
-
-    // *************************
-    // *** Function Overview ***
-    // *************************
-
-    // We categorize each column identifier c in string lhs_string into one of the following four categories:
-    //
-    // present in   |  rhs_string  | target_string |
-    // ---------------------------------------------
-    // batch        |      yes     |      yes      |
-    // kept left    |      no      |      yes      |
-    // contracted   |      yes     |      no       |
-    // summed left  |      no      |      no       |
-
-    // We do the same for rhs_string. Then, we transpose and reshape lhs_tensor and rhs_tensor in the following manner:
-    //
-    // lhs_tensor[batch, kept left, contracted, summed left]
-    // rhs_tensor[batch, kept right, contracted, summed right]
-    //
-    // IMPORTANT: the column identifiers of batch and contracted of both tensors are obviously the same.
-    //            We have to ensure that they also have the same dimensions and ordering!
-
-    // Now, we are ready to perform an axis summation on lhs_tensor over the last dimension (= summed left). The same for rhs_tensor.
[…]

### 2025-03-05 project/einspeed.py
Adding unit test for batch matrix matrix multiplication

@@ -3,6 +3,8 @@ import numpy as np
 import time
 import csv
 
+# Top level python program to test the implementation.
+
 def benchmark_einsum(format_string, lhs_shape, rhs_shape, dtype = np.float32, test_case = "Default", epsilon = 1e-4):
     lhs_tensor = (1+np.random.rand(*lhs_shape)).astype(dtype)
     rhs_tensor = (1+np.random.rand(*rhs_shape)).astype(dtype)

### 2025-03-05 project/tests/catch_tests_tensor_blas.cpp

@@ -1,0 +1,147 @@
+#include "catch.hpp"
+#include "blas.h"
+
+// Testing methods from the Tensor class and the batch matrix matrix multiplication method in blas.h.
+
+// Helper function to create a Tensor object with random data for testing
+template <typename T>
+Tensor<T> create_tensor(const std::vector<size_t>& shape) {
+    size_t size = 1;
+    for (auto s : shape) size *= s;
+    T* data = nullptr;
+    posix_memalign((void**)&data, 32, size * sizeof(T));
+
+    // Fill tensor with some test data (e.g., sequential integers)
+    for (size_t i = 0; i < size; ++i) {
+        data[i] = static_cast<T>(i + 1);
+    }
+
+    return Tensor<T>(shape.size(), shape.data(), data);
+}
+
+TEST_CASE("Tensor Transpose", "[transpose]") {
+    SECTION("Transpose 2D Tensor") {
+        std::vector<size_t> shape = {2, 3};
+        Tensor<float> tensor = create_tensor<float>(shape);
+        std::vector<size_t> perm = {1, 0};  // Swap rows and columns
+        Tensor<float> transposed = tensor.transpose(perm);
+
+        REQUIRE(transposed.get_shape() == std::vector<size_t>{3, 2});
+        REQUIRE(transposed.data[0] == tensor.data[0]);
+        REQUIRE(transposed.data[1] == tensor.data[3]);
+        REQUIRE(transposed.data[2] == tensor.data[1]);
+        REQUIRE(transposed.data[3] == tensor.data[4]);
+        REQUIRE(transposed.data[4] == tensor.data[2]);
+        REQUIRE(transposed.data[5] == tensor.data[5]);
+    }
+
+    SECTION("Transpose 3D Tensor") {
+        std::vector<size_t> shape = {2, 3, 4};
+        Tensor<float> tensor = create_tensor<float>(shape);
+        std::vector<size_t> perm = {2, 0, 1};  // Move third dim to first place
+        Tensor<float> transposed = tensor.transpose(perm);
+
+        REQUIRE(transposed.get_shape() == std::vector<size_t>{4, 2, 3});
+    }
+}
+
+TEST_CASE("Tensor Reshape", "[reshape]") {
+    SECTION("Reshape 1D to 2D Tensor") {
+        std::vector<size_t> shape = {6};
+        Tensor<float> tensor = create_tensor<float>(shape);
[…]

### 2025-03-05 project/tests/catch_tests_tensor.cpp
Adding unit test for batch matrix matrix multiplication

@@ -1,118 +1,0 @@
-#include "catch.hpp"
-#include "tensor.h"
-
-// Helper function to create a Tensor object with random data for testing
-template <typename T>
-Tensor<T> create_tensor(const std::vector<size_t>& shape) {
-    size_t size = 1;
-    for (auto s : shape) size *= s;
-    T* data = nullptr;
-    posix_memalign((void**)&data, 32, size * sizeof(T));
-
-    // Fill tensor with some test data (e.g., sequential integers)
-    for (size_t i = 0; i < size; ++i) {
-        data[i] = static_cast<T>(i + 1);
-    }
-
-    return Tensor<T>(shape.size(), shape.data(), data);
-}
-
-TEST_CASE("Tensor Transpose", "[transpose]") {
-    SECTION("Transpose 2D Tensor") {
-        std::vector<size_t> shape = {2, 3};
-        Tensor<float> tensor = create_tensor<float>(shape);
-        std::vector<size_t> perm = {1, 0};  // Swap rows and columns
-        Tensor<float> transposed = tensor.transpose(perm);
-
-        REQUIRE(transposed.get_shape() == std::vector<size_t>{3, 2});
-        REQUIRE(transposed.data[0] == tensor.data[0]);
-        REQUIRE(transposed.data[1] == tensor.data[3]);
-        REQUIRE(transposed.data[2] == tensor.data[1]);
-        REQUIRE(transposed.data[3] == tensor.data[4]);
-        REQUIRE(transposed.data[4] == tensor.data[2]);
-        REQUIRE(transposed.data[5] == tensor.data[5]);
-    }
-
-    SECTION("Transpose 3D Tensor") {
-        std::vector<size_t> shape = {2, 3, 4};
-        Tensor<float> tensor = create_tensor<float>(shape);
-        std::vector<size_t> perm = {2, 0, 1};  // Move third dim to first place
-        Tensor<float> transposed = tensor.transpose(perm);
-
-        REQUIRE(transposed.get_shape() == std::vector<size_t>{4, 2, 3});
-    }
-}
-
-TEST_CASE("Tensor Reshape", "[reshape]") {
-    SECTION("Reshape 1D to 2D Tensor") {
-        std::vector<size_t> shape = {6};
-        Tensor<float> tensor = create_tensor<float>(shape);
-        std::vector<size_t> new_shape = {2, 3};  // Reshape into a 2x3 matrix
[…]

### 2025-03-05 project/include/compute_einsum.h
adjusting python file and running scaling study again

@@ -2,7 +2,7 @@
 #define COMPUTE_EINSUM_H
 
 // uncomment the following lines to show debug information
-#define DEBUG
+// #define DEBUG
 // #define DEBUG_VERBOSE
 
 #include <unordered_map>

### 2025-03-05 project/einspeed.py
adjusting python file and running scaling study again

@@ -5,6 +5,7 @@ import csv
 
 # Top level python program to test the implementation.
 
+# This helper method creates randomized tensors of the specified shape and compares the running time of my einsum implementation with NumPy's.
 def benchmark_einsum(format_string, lhs_shape, rhs_shape, dtype = np.float32, test_case = "Default", epsilon = 1e-4):
     lhs_tensor = (1+np.random.rand(*lhs_shape)).astype(dtype)
     rhs_tensor = (1+np.random.rand(*rhs_shape)).astype(dtype)
@@ -23,6 +24,7 @@ def benchmark_einsum(format_string, lhs_shape, rhs_shape, dtype = np.float32, te
     maximum_relative_discrepancy = np.max(np.abs((result - correct) / correct))
 
     print(f"Test Case: {test_case} {str(dtype)}")
+    # There will be an error, and it might grow big due to numerical issues. It should just be numerical issues though.
     print("Correct: ", (maximum_relative_discrepancy < epsilon) and (lhs_tensor==lhs_copy).all() and (rhs_tensor==rhs_copy).all())
     print(f"Maximum relative discrepancy: {maximum_relative_discrepancy}")
     print(f"NumPy: {numpy_time} s\neinspeed: {einspeed_time} s")
@@ -32,6 +34,7 @@ def benchmark_einsum(format_string, lhs_shape, rhs_shape, dtype = np.float32, te
     return (numpy_time, einspeed_time, maximum_relative_discrepancy)
 
 
+# Simple utility method to create .csv files.
 def create_csv(file_name, *args):
     """
     Creates a CSV file from column data provided as tuples.
@@ -61,6 +64,7 @@ def create_csv(file_name, *args):
             writer.writerow(row)
 
 
+# Testing various einsum expression for correctness and speed.
 def run_test_cases():
     DTYPES = [np.float32, np.float64, np.complex128]
     for dtype in DTYPES:
@@ -70,41 +74,28 @@ def run_test_cases():
         benchmark_einsum("abc,def->", [50, 42, 55], [63, 69, 127], dtype, "Multidimensional Reduction")
 
         # Contraction
-        benchmark_einsum("i,i->", [600000000], [600000000], dtype, "Vector Dot Product")
-        benchmark_einsum("ij,ij->", [9000, 80000], [9000, 80000], dtype, "Element-wise Contraction")
[…]

### 2025-03-05 paper/paper.tex
Little improvements in the paper

@@
 To this end, I implemented \texttt{is\_valid\_einsum\_expression} function found in \texttt{include/compute\_einsum.h}.
-It checks the following five criteria:
+It checks the following four criteria:
 
@@
 Of course we can now free both transposed tensors after the reduction operation, which we also do.
-However, I figured that it might be beneficial to actually keep some memory, preferably from the bigger tensor, in order to later reuse it for the result of the following batch matrix multiplication operation.
+However, I realized that it might be beneficial to actually keep some memory, preferably from the bigger tensor, in order to later reuse it for the result of the following batch matrix multiplication operation.
 
@@
 \subsection{Batch Matrix Multiplication}
-Let us now investigate the final sub-function \texttt{batch\_matrix\_matmul\_transpose} found in \texttt{include/}.
+Let us now investigate the final sub-function \texttt{batch\_matrix\_matmul\_transpose} found in \texttt{blas.h}.
 It is basically very straight forward, also note that we treat the rhs operand batch matrices as transposed, which allows for small strides in the inner most loop.
